### 2017-05-15 Thresholding Bandit/ThresholdingBandit.Rmd
add gitignore, update Thresholding Bandit

--- 
title: "The Thresholding Bandit Problem"
author: "Tim Radtke" 
date: "4/23/2017"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \newtheorem{theorem}{Theorem}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    #latex_engine: xelatex
fontsize: 12pt
urlcolor: blue
---

The *Thresholding Bandit Problem* described in Locatelli et al. (2016) can be considered a variant of a variant. It can be framed into the wider literature of *Pure Exploration* multi-armed bandit problems. In particular, it shares characteristics with the *Top-$m$* problem. This problem is concerned with finding the $m$ best arms as described by the means of their corresponding distributions. This in turn is similar to the *combinatorial bandit* problem, which also aims at finding the $m$ best arms. However, it is able to pull several arms at once: Think of an online shop that shows five recommended products on a product detail page. These five recommended products might each be represented by an arm, and we look for the products with the largest mean conversion rate. The thresholding bandit problem we discuss here, however, is concerned with pulling a single arm at a time. And so a more appropriate situation is that of a website presenting a banner. Again, think of an online shop trying to promote a certain category. The content team came up with a number of different designs for the banner, and it's not clear how many click-throughs they will gather.

The idea now is that we would like to classify the banners into two distinct groups: A group with a mean conversion rate $\mu_i$ below threshold $\tau$, and a group with mean conversion rate $\mu_i$ above threshold $\tau$. This might be the optimization problem when we are concerned with not falling below a certain minimum level click-through rate with the banners we're choosing.

If it turns out that in general we can find relatively quickly whether arms are above or below a threshold, then this kind of test can be used to safeguard against very bad versions in a test. Before we move to a Top-$m$ test, we might want to run a thresholding version and then continue only with the arms classified into the group above the threshold. This might be justified when adjusting parts of the checkout process of an online shop, where the conversion rate or the average order value should not drop below a threshold.

## Setup

In any case, the problem boils down to the following. As standard in multi-armed bandit problems, we have $K$ arms $\mathcal{A} = \{1, \dots K\}$. We can pull arm $k$ at time $t$ to collect feedback in form of the random variable $X_{k,t}$ which is distributed accoring to the arm's distribution $\nu_k$. In general, we are concerned with estimating for each arm $i$ the mean $\mu_i$ of its distribution $\nu_i$. In contrast to most bandit problems, however, we do not directly compare arms with each other by comparing their means. In the case of pure exploration bandits, it is for example necessary to compare the arms because one aims to find the best one. In the case of the thresholding bandit, however, we compare each arm's mean $\mu_i$ individually against the threshold $\tau \in \mathbb{R}$ which is known and fixed before the experiment. One might compare this to a pure exploration bandit in which the mean of the best arm is known upfront. We will see later that this fact leads to advantages in the design of algorithms which are not as readily available in other bandit settings.

In Locatelli (2016), the arms' distributions are assumed to follow a $R$-sub-Gaussian distribution. Under this assumption, the authors are able to proof a general lower bound as well as an algorithm with a matching upper bound. Both hinge upon the *complexity* of a given problem. Since bandit problems can differ in the number of the arms $K$ and the means of each arm, some pose easier identification problems than others. Indeed, in the thresholding problem, the difficulty of the problem naturally depends on how close (the means of the) distributions are to the fixed threshold. Intuitively, the closer a mean $\mu_i$ is to the threshold $\tau$, the more difficult it will be to distinguish the two successfully based on samples from the distribution $\nu_i$. For the case of the sub-Gaussian distributions, a natural measure of this distance is given by simply $\Delta_i = |\tau - \mu_i| + \epsilon$, where $\epsilon$ defines a $2\epsilon$ interval around the threshold $\tau$ in which we allow ourselves to make a mistake at no cost.

If we extend this idea to the set of arms $i \in \{1,...,K\}$ of a given bandit problem, a natural measure of the overall diffulty of the problem, or the problem's complexity, might be given by $H = \sum_{i=1}^{K} \Delta_i^{-2}$. In particular, it shows that just adding additional arms leads to an increase in complexity (since we need to sample more of them), just as smaller $\Delta_i$ will increase the complexity as arms become harder to classify. (WHERE DOES THIS COMPLEXITY ORIGINALLY COME FROM? WHAT IS ITs MOTIVATION? ITS NOT THAAAT NATURAL)

We also see that this definition may lose relevancy as we move away from $R$-sub-Gaussian distributions. (ADD QUICK EXAMPLE OF WHAT CHANGES; REFER TO LATER PART AND DESCRIBE THIS AS THE MOTIVATION OF THE PAPER)

Another characteristic uniquely identifying bandit problems is the regret measure that we aim to optimize. As explained during the introduction, multi-armed bandits traditionally optimize cumulative regret. The notion of regret considered in the thresholding bandit, however, is more akin to the regret in pure exploration bandits or the error probability in sequential hypothesis testing. Indeed, one can most easily describe the regret in the thresholding case as the probability of making a misclassification of any arm. Or, more rigorously, let us define the set of all arms with mean larger than the threshold, $\mathcal{S}_\tau = \{i \in \mathcal{A}: \mu_i \geq \tau\}$. Consequently, $\mathcal{S}_\tau^C = \{i \in \mathcal{A}: \mu_i < \tau\}$. At the end of the sampling rule/procedure (!!we have not mentioned yet the fact that we're in the fixed budget case!!), we demand that any algorithm returns the two corresponding *estimated* sets: $\hat{\mathcal{S}}_{\tau} = \{i \in \mathcal{A}: \hat{\mu}_i \geq \tau\}$, as well as $\hat{\mathcal{S}}_{\tau}^C = \{i \in \mathcal{A}: \hat{\mu}_i < \tau\}$. We simply demand an answer to the question which arms have means above the threshold $/tau$, and which arms lie below  $\tau$. Let $T$ be the overall number of samples the algorithm is allowed to pull. We define the loss as $\mathcal{L}(T) = \mathbbm{1}_{\{\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset\}} \cdot \mathbbm{1}_{\{\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset\}}$. That is, we compare the return set with the true set and collect a loss of 1 if any arm with a true mean outside the $2\epsilon$ interval around $\tau$ was misclassified. It is important to realize that arms with means inside the $2 \epsilon$ interval may be misclassified without it affecting the loss. These arms are considered close enough to the threshold that we do not care whether they lie above or below.

ADD A GRAPHIC THAT VISUALIZES THE ESTIMATED AND THE REQUIRED RETURN SETS

From there, it is natural to define the expected regret/loss as $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \land (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset))$. This, of course, is the mentioned probability of making a mistake.

ADD REMARK ON WHAT THIS IMPLIES WITH REGARD TO THE MEANS

Lastly, it is to mention that we are considering the fixed-budget version of the thresholding bandit problem. We thus fix the number of samples that the algorithm will pull upfront at $T$. The task of the algorithm is then to find a sampling rule or strategy which distributes the budget $T$ among the $K$ arms in such a way that the expected regret, or the error probability, is minimized. In other terms, the algorithm aims to use the $T$ samples in a way that lets him return the two sets $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}_{\tau}^C$ with as much confidence as possible. This again shows the contrast to the fixed confidence setting in which one fixes the confidence level upfront and lets the algorithm draw as many samples as necessary to reach the required confidence.

At this point, we have defined the fixed budget thresholding bandit setup for any algorithm in the case of sub-Gaussian distributions. We thus proceed to describe the algorithm proposed in Locatelli et al. (2016).

## The APT Algorithm

In the following we describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016), which is a simple yet effective algorithm and displays some very favorable characteristics when compared with alternative algorithms.

From the general setup, we already know how the algorithm will end. After it has drawn $T$ samples, the algorithm makes its decision of which arms to include in the return sets solely on basis of the empirical means $\hat{\mu}_i(T)$ estimated until time $T$. Furthermore, the threshold and interval around the threshold were fixed upfront and serve as argument to the algorithm: $(\tau, \epsilon)$. For a given arm $i$ and a given time $t$, we let $T_i(t)$ represent the number of samples drawn from arm $i$ until time $t$ (including). Consequently, the empirical mean of arm $i$ at time $t$ is given by $\mu_i = \frac{1}{T_i(t)} \sum_{s=1}^{T_i(t)} X_{i,s}$ where $X_{i,s}$ is the feedback sampled from distribution $\nu_i$ when the arm is pulled for the $s$-th time.

To be specific, the APT algorithm initiates by pulling each arm once. Then, for each $t > K$, it calculates for every arm the values $T_i(t)$ and $\mu_i(t)$. Based on the empirical mean, the algorithm computes the estimated gap between the mean and the threshold as $\hat{\Delta}_i(t) = \hat{\Delta}_i^{\tau, \epsilon}(t) = |\mu_i(t) - \tau| + \epsilon$. The decision of which arm is sampled next is then based on the value $B_i(t+1) = \sqrt{T_i(t)} \hat{\Delta}_i(t)$: the algorithm pulls arm $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$ minimizing this decision quantity. After $T$ pulls overall, the arm returns the sets described in the general setup, $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}^C_{\tau}$.

**Remark**: Obviously, the APT algorithm tends to pull arms more often with corresponding mean close to the threshold and interval (small $\hat{\Delta}_i(t))$. This makes sure that more evidence is gathered on arms which appear to be inherently difficult to classify based on the samples collected until time $T_i(t)$, as these arms stifle the confidence the most. That is, until we have collected enough evidence to make a confident classification even if the mean is truly close to the threshold. In that case, $\sqrt{T_i(t)}$ will raise $B_i(t+1)$. Consequently, other arms farther away from the threshold will be sampled again. In this way, the algorithm tries to distribute the budget $T$ over all arms so as to define a sampling rule optimizing the confidence of the final decision. Of course, that's just Glivenko-Cantelli and the CLT making sure $B_i(t+1)$ is not degenerate.

## Discussion

One of the APT algorithm's main features is its simplicity. The sampling rule is very straightforward, and, as highlighted in Locatelli et al. (2016), it does not require any knowledge about the type of sub-Gaussianity $R$, or the complexity $H$ upfront. This makes the algorithm immediately applicable in practice while it still enjoys its upper bound on the expected regret. This is in particular worth highlighting when comparing APT to algorithms used in the pure exploration bandit setting. There, many algorithms require knowledge about the corresponding complexity upfront. The problem however being, that to know the complexity, the experimenter has to know all means up to a permutation of the arms (for example, see Audibert et al., 2010). This, of course, is not given in practice and thus alternative algorithms try to estimate the complexity online which is not a desirable solution either as especially in the beginning plug-in estimates based on the estimated means may lead to wrong estimates of the complexity and consequently budget allocation based on mistaken beliefs (compare Garivier and Kaufmann, 2016).

Furthermore, Locatelli et al. (2016) were able to present an upper bound for the APT algorithm matching the lower bound in their Theorem 1. Proofs of both bounds are repeated below.

Where the APT algorithm shows room for improvement is when it comes to settings with distributions that are not sub-Gaussian. This is mentioned in Locetelli et al. (2016). They suggest ...