### 2017-05-25 Complexity of Best Arm Identification/kl_complexity.Rmd
Add proof of lower bound on FB complexity

---
title: "On the Complexity of Best Arm Identification"
author: "Tim Radtke"
date: "5/20/2017"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{lemma}{Lemma}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    #latex_engine: xelatex
fontsize: 12pt
urlcolor: blue
---

In the previous chapter, we have discussed the thresholding bandit problem as presented in Locatelli et al. (2016). Doing so, we came across the complexity term $H = \sum_{i=1}^{K} \Delta_i^{-2}$. Furthermore, we have noted the fact that an algorithm and bounds based on *only* the gap between the respective mean and the threshold are useful in the case of sub-Gaussian problems. Furthermore, even for the case of Bernoulli arms, which are sub-Gaussian since their feedback is bounded on $[0,1]$, the gap $\Delta$ becomes less and less appropriate for situations in which both $\mu$ of the Bernoulli, and $\tau$ are very close to 0 or 1.

Even though the complexity $H$ enjoys popularity in the literature of the best m arm identification problem, $m \geq 1$, Kaufmann et al. (2016) show explicitly that $H$ is mainly useful for problems of Gaussian distributions. Indeed, they go so far as to define new measures of complexity that are based on information theoretic metrics. Using those they then display $H$ as an approximation to their measures in the case of Gaussian distributions. In turn, this makes apparent that especially Bernoulli distributions with small means have to be treated differently.

For the current chapter, we will adopt the notation of Kaufmann et al. (2016) so that it will be easier to compare the arguments.

Instead of the thresholding bandit problem, we now consider the problem of identifying the best $m$ out of $K$ arms, that is, the Top-$m$ problem. While we previously looked at a fixed budget formulation, we now mainly focus on a fixed confidence problem. That is, we hold the confidence $1-\delta$ fixed at which a strategy returns the correct $m$ arms while trying to find them in a way that minimizes the sample complexity $\mathbb{E}_{\nu}[\tau]$, that is, the expected number of overall draws of the arms. $\tau$ is also called the stopping time of the algorithm. Algorithms are supposed to find $\mathcal{S}_m^*$, the set of the $m$ arms with the largest means. If we let $(\mu_1, \dots, \mu_K)$ be the vector of the means of the arms, then let $(\mu_{[1]}, \dots, \mu_{[K]})$ be the tuple in which the means are arranged in decreasing order. For the rest of this chapter, we consider only settings with arms for which $\mu_{[m]} > \mu_{[m+1]}$; that is, the set of the top $m$ arms, $\mathcal{S}_m^*$, is unique. After the sampling phase, we require any algorithm to return the set $\hat{S}_m \subset \{1,\dots,K\}$ of $m$ arms that the algorithm believes to have the largest means. In the case of fixed confidence, we fix $\delta$ at the beginning, such that the algorithm only stops sampling when $\mathbb{P}_{\nu}(\hat{S}_m = \mathcal{S}_m^*) \geq 1-\delta$. We try to find algorithms who fulfill this as quickly as possible; that is, they minimize the expected stopping time (or number of draws), $\mathbb{E}_{\nu}[\tau]$.

To arrive at the new measure of complexity proposed by Kaufmann et al. (2016), it will actually make sense to first follow their arguments to a general lower bound on the sample complexity $\mathbb{E}_{\nu}[\tau]$ in the case of fixed confidence problems. This bound is given in their Theorem 4 (for common confidence levels of $\delta \leq 0.15$).

\begin{theorem}[Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem4}
Let $\nu \in \mathcal{M}_m$, where $\mathcal{M}_m$ is the set of uniquely identifiable bandit models described above. Assume that $\mathcal{P}$ satisfies the continuity conditions in Assumption 3. Any $\delta$-PAC algorithm on $\mathcal{M}_m$ satisfies, for $\delta \leq 0.15$:

\begin{equation*}
\mathbb{E}_{\nu}[\tau] \geq \Big[ \sum_{a \in \mathcal{S}^*_m} \frac{1}{KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}^*_m} \frac{1}{KL(\nu_a, \nu_{[m]})} \Big] \der(1-\delta, \delta),
\end{equation*}

where $\der(x,y) := x \log(x/y) + (1-x) \log((1-x)/(1-y))$ is the binary relative entropy, with the convention that $\der(0,0) = \der(1,1) = 0$.
\end{theorem}

To proof this statement, we assume as before that the arms are ordered by their means decreasing in size. For a given $\delta$-PAC algorithm $\mathcal{A} = ((A_t), \tau, \hat{S}_m)$, and under the continuity conditions, there exists for every arm $a \in \{1, \dots, K\}$ an alternative model $\nu' = (\nu_1, \dots, \nu_{a-1}, \nu_a', \nu_{a+1}, \dots, \nu_K)$. In this alternative model, only arm $a$ is modified in accordance with the continuity conditions. Modified in such a way that measured by the Kullback-Leibler divergence, the arm $\nu_{m+1}$ is now closer to $\nu_a$ than $\nu_a'$ is to $\nu_a$, if $\mu_a > \mu_{m+1}$ (consequently, $\mu_a' < \mu_{m+1}$). On the other hand, if $\mu_a < \mu_{m}$, then $\nu_a'$ will be moved away from $\nu_a$ just enough so that $\nu_a$ is closer to $\nu_m$ than to $\nu_a'$ (consequently, $\mu_a' > \mu_m$). These changes in distribution make the original set of optimal arms, $\{1,\dots,m\}$ no longer optimal under the alternative model $\nu'$. Given that the algorithm returns $\hat{S}_m$, we know that for the event $\mathcal{E} = (\hat{S}_m = \{1, \dots, m\})$ (which is element of the by the stopping time $\tau$ induced $\sigma$-Algebra $\mathcal{F}_{\tau}$), that $\mathbb{P}_{\nu}(\mathcal{E}) \geq 1-\delta$ and $\mathbb{P}_{\nu'}(\mathcal{E}) \leq \delta$ (because fixed confidence $\delta$!).

To now give the lower bound on the expected number of samples, $\mathbb{E}_{\nu}[\tau]$, we employ the general lemma given by the authors for such changes in distribution in bandit models.

\begin{lemma}[Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma1}
Let $\nu$ and $\nu'$ be two bandit models with $K$ arms such that for all $a$, the distributions $\nu_a$ and $\nu_a'$ are mutually absolutely continuous. For any almost-surely finite stopping time $\sigma$ with respect to $(\mathcal{F}_t)$,

\begin{equation*}
\sum_{a=1}^{K} \mathbb{E}_{\nu} [N_a(\sigma)] \KL(\nu_a, \nu_a') \geq \sup_{\mathcal{E} \in \mathcal{F}_{\sigma}} \der (\mathbb{P}_{\nu}(\mathcal{E}), \mathbb{P}_{\nu'}(\mathcal{E})),
\end{equation*}

where $\der(x,y)$ is the binary relative entropy as defined above.
\end{lemma}

We can apply this lemma directly on the expected number of draws for each arm, $\mathbb{E}_{\nu}[N_a]$. For each arm $a$, we thus have $\KL(\nu_a, \nu_a') \mathbb{E}_{\nu}[N_a] \geq \der (1-\delta, \delta)$. Using the definition of the alternative model above, we get for every small but fixed $\alpha > 0$, for every arm $a \in \{1, \dots, K\}$ and each arm $b \in \{m+1, \dots, K\}$:

\begin{equation*}
\mathbb{E}_{\nu}[N_a] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_a, \nu_{m+1}) + \alpha}
\end{equation*}

and

\begin{equation*}
\mathbb{E}_{\nu}[N_b] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_b, \nu_{m}) + \alpha}.
\end{equation*}

We can then let $\alpha \to 0$, sum the bounds on the individual arms to get the bound on the overall number of draws:

\begin{equation*}
\mathbb{E}_{\nu}[\tau] = \sum_{a=1}^K \mathbb{E}_{\nu}[N_a] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})} \Big) \der(1-\delta, \delta)
\end{equation*}

Given that we will be naturally interested in either very high or very low confidence of our algorithm, that is, $P_\nu(\mathcal{E})$ is either close to 0 or 1, it is possible to use the following approximation to the binary relative entropy:

\begin{equation*}
\forall x \in [0,1], \quad \der(x,1-x) = \der(1-x,x) \geq \log \frac{1}{2.4x}
\end{equation*}

And so the final lower bound can be written as

\begin{equation*}
\mathbb{E}_{\nu}[\tau] \geq \Big( \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})} \Big) \log \big(\frac{1}{2.4\delta} \big).
\end{equation*}

```{r, echo = FALSE}
x <- seq(0,1,by = 0.001)
bre <- function(x) x * log(x/(1-x)) + (1-x) * log((1-x)/x)
plot(x,bre(x), type = "l")
lines(x, log(1/(2.4*x)), col = "red")
abline(v = 0.15, lty = 2)
```

Given this general lower bound for the sampling complexity, or expected number of draws, it makes sense to define the complexity of a Top-$m$ bandit problem using the Kullback-Leibler divergence. After all, besides the number of arms and the experimenter-chosen confidence, it is just how close arms are with respect to each other in terms of Kullback-Leibler divergence that influences the sampling complexity. And in general, a problem requiring more samples is of course considered to be more complex (at least with respect to the samples, which is what needs to be optimized).

Thus, following Kaufmann et al. (2016), define the complexity $\mathcal{K}_C(\nu)$ for the fixed confidence setting as

\begin{equation}
\mathcal{K}_C(\nu) = \inf_{A \, \text{PAC}} \lim_{\delta \to 0} \sup \frac{\mathbb{E}_{\nu}[\tau_{\delta}]}{\log \frac{1}{\delta}}.
\end{equation}

So if we were to plug in the lower bound on the sampling complexity, we would arrive at a complexity measure that basically depends only on the Kullback-Leibler measuring for each arm how far it is away from the border between the two groups of arms. To see this, consider the lower bound on the complexity measure given in Kaufmann et al. (2016):

\begin{equation*}
\mathcal{K}_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a, \nu_{[m]})}
\end{equation*}

Given this description, we can already note that an intuitive extension of this complexity notion to the thresholding bandit problem might involve replacing $\mu_{[m]}$ and $\mu_{[m+1]}$ by the actual threshold $\tau$ (notation in Locatelli et al.) that divides the group above the threshold from the group below the threshold. In that case, the previously used $\alpha$ might be very similar to the $\epsilon$ used in Locatelli et al. (2016).

To get there, however, at least two important steps remain. First, we need to define a complexity for the fixed budget problem. Second, we need to create algorithms that can match the lower bound.

Kaufmann et al. are able to provide a bound on their complexity measure in the fixed budget case, however only in the case of **two-armed** bandit models. As demanded in their paper, a fixed-budget algorithm is called consistent if the failure probability $p_t(\nu) := \mathbb{P}_{\nu}(\hat{S}_m \neq S^*_m)$ tends to zero as the number of samples goes to infinity. Also, define the complexity in the fixed-budget setting as follows:

\begin{equation}
\mathcal{K}_B(\nu) = \inf_{A \, \text{consistent}} \big(\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu)\big)^{-1}.
\end{equation}

Given this definition, one can interpret the bound given in Theorem 12 of Kaufmann et al. directly as a lower bound on the complexity $\mathcal{K}_B(\nu)$ (note the inverse).

\begin{theorem}[Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem12}
Let $\nu = (\nu_1, \nu_2)$ be a two-armed bandit model such that $\mu_1 > \mu_2$. In the fixed-budget setting, any consistent algorithm satisfies:

\begin{equation*}
\lim_{t \to \infty} \sup - \frac{1}{t} \log p_t(\nu) \leq \inf_{(\nu_1', \nu_2') \in \mathcal{M}: \mu_1' < \mu_2'} \frac{\KL(\nu_1', \nu_1) + \KL(\nu_2', \nu_2)}{2}
\end{equation*}
\end{theorem}

For a proof, consider first the two different bandit models, $\nu$ and $\nu'$, where we fix without loss of generality that $a^* = 1$ is the best arm in $\nu$, while $a^* = 2$ in $\nu'$. Note that in this case we don't necessarily try to fool the algorithm by a change of distribution. Indeed, we want the algorithm $\mathcal{A}$ to be consistent for both problems. Since we are in the fixed-budget setting, the stopping time is fixed at $\tau = t$. Even though the correct decision differs between the two problems, we focus on the event $A = (\hat{S}_1 = 1)$. Given the formulation in Kaufmann et al. (2016), we have $A \in \mathcal{F}_t = \mathcal{F}_{\tau}$. Now, we can apply Lemma 1 as before, as it in essense gives a measure of how different two bandit models are. Here, we want to know how different $\nu$ is from $\nu'$, and so, given we have only two arms,

\begin{equation*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2) \geq \der(\mathbb{P}_{\nu'}(A),\mathbb{P}_{\nu}(A)))
\end{equation*}

The expectations were chosen with respect to $\nu'$ in order to allow the for the following: We have $p_t(\nu) = 1 - \mathbb{P}_{\nu}(A)$ and $p_t(\nu') = \mathbb{P}_{\nu'}(A)$ (since $A$ is wrong in $\nu'$). But since algorithm $\mathcal{A}$ is consistent for both $\nu$ and $\nu'$, we know that there should exist some number of samples such that the probability of event $A$ is bigger in setting $\nu$ (where it is the correct decision), than in setting $\nu'$ (where this would be a wrong decision and failure). Thus, for every $\epsilon > 0$, there exists $t_0(\epsilon)$ such that for all $t \geq t_0(\epsilon)$, $\mathbb{\nu'}(A) \leq \epsilon \leq \mathbb{P}_\nu(A)$. Consequently, if we plug in the bound of $\epsilon$ for $\mathbb{P}_{\nu'}(A)$, then

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2) & \geq \der(\epsilon,1-p_t(\nu))) \\
& = \epsilon \log \frac{\epsilon}{1-p_t(\nu)} + (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} \\
& \geq (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} + \epsilon \log \epsilon.
\end{align*}

Let $\epsilon$ go to zero so that

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) + \mathbb{E}_{\nu'}[N_2(t)]\KL(\nu_2', \nu_2)
& \geq \log \frac{1}{p_t(\nu)} = -\log p_t(\nu).
\end{align*}

Dividing both sides by $t$ and taking the limsup, we arrive at

\begin{align*}
\lim_{t \to 0} \sup - \frac{1}{t} \log p_t(\nu)
& \leq \lim_{t \to 0} \sup \frac{1}{t} \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)] \KL(\nu_a', \nu_a) \\
& \leq \max_{a=1,2} \KL(\nu_a', \nu_a),
\end{align*}

where the last line follows from the fact that the $\frac{1}{t} \sum_{a=1}^2 \mathbb{E}_{\nu'}[N_a(t)]$ part creates a weighted average of the two divergences $\KL(\nu_a', \nu_a)$. As noted in Kaufmann et al. (2016), one arrives at the final result by choosing $\mu_1'$, $\mu_2'$ so that $| \KL(\nu_1', \nu_1) - \KL(\nu_2', \nu_2) |$ is as small as possible.