### 2017-08-06 Thesis/APT_UB_revised/Untitled.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={APT Upper Bound},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{APT Upper Bound}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}

\begin{document}
\maketitle

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{3}
\tableofcontents
}
\subsection{Upper Bound on Loss of
APT}\label{upper-bound-on-loss-of-apt}

\begin{theorem}[Locatelli et al., 2016] \label{theorem:LocatelliTheorem4}
Let $K \geq 0$, $T \geq 2K$, and consider a problem $\mathbb{B}$. Assume that
all arms $\nu_k$ of the problem are R-sub-Gaussian with means $\mu_k$. Let $\tau
\in \mathbb{R}, \epsilon \geq 0$.

Algorithm APT's expected loss is upper bounded on this problem as 
\begin{equation*} \mathbb{E}(\mathcal{L}(T)) \leq \exp
(-\frac{1}{64R^2}\frac{T}{H} + 2 \log (T) + 1)K)) \end{equation*}

where we remind that $H = \sum_i (|\mu_i - \tau | + \epsilon)^{-2}$ and where
$\mathbb{E}$ is the expectation according to the samples of the problem.
\end{theorem}

\emph{Proof of Theorem \ref{theorem:LocatelliTheorem4}:}

In order to bound the expected loss of the APT algorithm, we want to
find an event on which the algorithm would correctly seperate arms from
the threshold, that is, on the event a seperation condition holds. We
then want to show that with large probability, this \emph{favorable}
event does indeed hold. Consequently, the expected loss (probability of
an error) will correspond to the probability that the favorable event
does not hold after all as this leads the seperation condition to fail,
and the algorithm to classify at least one arm incorrectly.

While this is the general idea, we have to jump between seperation
condition, favorable event, and large probability.

\subsubsection{Separation Condition}\label{separation-condition}

As noted in Zhong et al. (2017), a simple requirement to correctly
classify arms at the final round is the following. For every arm
\(i \in \{1, ..., K\}\) at round \(T\),

\[
\hat{\mu}_{i,T_i(n)} \geq \mu_i - \frac{\mu_i - \tau + \epsilon}{2} \quad \text{if} \quad \mu_i \geq \tau
\] or

\[
\hat{\mu}_{i,T_i(n)} \leq \mu_i + \frac{\tau - \mu_i + \epsilon}{2} \quad \text{if} \quad \mu_i < \tau
\] Or simply in terms of the gap between the arm and the threshold as
used as index by the algorithm,

\[
| \hat{\mu}_{i, T_i(n)} - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
\]

Intuitively, the conditions make sense as they ensure that the current
mean estimate is closer to the true mean than to the threshold.
Consequently, there is no possible arrangement left in which the
condition would hold, but the arm be classified incorrectly. Assuming
\(\mu_i < \tau\) and \(\epsilon = 0\), an incorrect classification would
require \(\mu_i < \tau \leq \hat{\mu_i}\). But then
\(| \hat{\mu}_{i, T_i(n)} - \mu_i | \geq |\mu_i - \tau|\), violating the
separation condition.

Thus, we need to show that in the very final round, this condition holds
for every arm \(i \in \{1,...,K\}\) with large probability. Or more
precisely, we need to derive a bound on the probability with which this
does not hold.

\subsubsection{The Favorable Event}\label{the-favorable-event}

The separation condition gives a good idea of what event we need to hold
in order to make a correct classification. The smaller the deviation of
the empirical mean from the population mean, the more difficult it is
for the classification to be wrong. Moreover, if the mean and empirical
mean are identical, then the classification is correct independent of
the threshold. And furthermore, the closer the mean and the threshold
are, the better the estimate must be for a correct classification.

Having to bound the absolute deviation of the sample mean from the
population mean is a very thankful operation. Define the favorable event
\(\xi\) as

\[
\xi = \Big\{\forall i \in \mathbb{A}, \forall s \in \{1,...,T\} : |\frac{1}{s} \sum_{t=1}^{s}X_{i,t} - \mu_i| \leq \sqrt{\frac{T \delta^2}{H s}} \Big\}.
\]

We want the probability that this favorable event occurs to be large,
that is, the probability that the deviation is small, such that the
separation condition holds with large probability. Indeed, we bound the
deviation for \emph{all} arms \(i \leq K\) and for \emph{every} round
\(s \leq T\). Using some Chernoff Hoeffding bounds, we are able to write

\[
\mathbb{P}(\xi) \geq 1 - 2(\log(T) +1)K \exp(-\frac{T\delta^2}{2R^2H})
\]

\subsection{\texorpdfstring{Need to Bound
\(T_i\)}{Need to Bound T\_i}}\label{need-to-bound-t_i}

We now have that with large probability, the favorable event holds. On
the favorable event, we have for every arm \(i\):

\[
| \hat{\mu}_i(t) - \mu_i| \leq \sqrt{\frac{T\delta^2}{HT_i(t)}}
\]

The deviation of arm \(i\) at time \(t\) depends on how large \(H\),
\(T\), and \(T_i(t)\) are. While \(t\), \(H\), \(T\) are deterministic,
the number of times arm \(i\) has been pulled at round \(t\),
\(T_i(t)\), is a random variable. In order to know how large the
probability of the favorable event is (and thus the error probability),
we need to know what realistic values for \(T_i(t)\) will be. Put
differently: To ensure a deviation smaller than
\(\sqrt{\frac{T\delta^2}{HT_i(t)}}\) for arm \(i\), we need to ensure
that arm \(i\) has been pulled at least \(T_i(t)\) times. As usual in
multi-armed bandits, and different from uniform sampling, the number of
pulls is a random variable that also depends on how often other arms are
pulled. In general, we do not want each arm to be pulled an equal
amount, but each arm should be pulled some times, and not too many
times. This especially holds for our exploration setting.

Thus, in order to check that the deviation of arm \(i\) really is small
enough for the separation condition to hold on the favorable event, we
need to bound the number of pulls of every arm from below.

\subsection{\texorpdfstring{Characteristics of a Frequently Pulled Arm
(Lower Bound on
\(T_k(T)\))}{Characteristics of a Frequently Pulled Arm (Lower Bound on T\_k(T))}}\label{characteristics-of-a-frequently-pulled-arm-lower-bound-on-t_kt}

Given that we do not follow a uniform sampling strategy, we know that at
time \(T\), some arms will have been pulled more frequently than others.
In particular, some arms will have been pulled more than \(T/K\) times.
How often an arm is pulled can be related to the individual arm's
contribution to the overall complexity measure. Let \(h_i\) be a measure
of arm \(i\)'s individual problem complexity and write for the overall
problem complexity

\[
H = \sum_{i = 1}^{K} h_i.
\]

Then arm \(i\) contributes a share of \(h_i/H\) to the overall problem
complexity. We can now show that there exists an arm that has been
pulled at least as often as the share of his complexity contribution.

Given the initialization of the APT algorithm, every arm is pulled once
initially. Consequently, there is a budget of \(T-K\) rounds left after
the initialization. As shown in Locatelli et al. (2016), there exists an
arm \(k\) that has been pulled after the initialization (\(\geq 2\)
pulls) and for which at the final round \(T\) holds:
\(T_k(T) - 1 \geq \frac{(T-K)h_k}{H}\). This can be shown by a quick
contradiction proof. Assume the opposite, that is

\[
\forall i \in \{1,...,K\}: T_i(T) - 1 < \frac{(T-K)h_i}{H}
\]

Consequently, at time \(T\), we would have pulled across all arms
\(\sum_{i=1}^KT_i(T)-1\) times. If the previous assumption holds, we
would not have used the entire budget:

\[
\sum_{i=1}^K T_i(T)-1 < \sum_{i=1}^K \frac{(T-K)h_i}{H} = (T-K)\sum_{i=1}^K \frac{h_i}{H} = T-K
\]

This is a contradiction, since as an requirement to the algorithm we
have that \(\sum_{i=1}^K T_i(T)-1 = (T-K)\), that is, we use the entire
budget of samples.

Another such requirement has been that \(T>2K\). Plugging this in gives
\[
T_k(T) - 1 \geq \frac{(T-K)h_k}{H} \leq \frac{(T-1/2T)h_k}{H} = \frac{Th_i}{2H}
\]

Now we know how often the arm has been pulled at the very end. That does
not mean that \(k\) has been pulled at time \(T\). Thus we consider the
time \(t \leq T\) at which arm \(k\) is pulled for the last time. We get
(!!!!!SHOOOOWWW THIIIISSSS!!!!!!!!!)

\[
T_k(t) \geq T_k(T) - 1 \geq \frac{Th_i}{2H}.
\]

This gives us a lower bound on how often arms are pulled that are pulled
frequently relative to their contribution to the problem complexity. Now
we need a lower bound on \(T_i(T)\) for every other arm \(i\).

\subsection{\texorpdfstring{Lower Bound on
\(T_i(T)\)}{Lower Bound on T\_i(T)}}\label{lower-bound-on-t_it}

Consider again as before the round \(t\) in which arm \(k\) is pulled
for the very last time. Since the APT algorithm minimizes at every round
\(s\) the index \(B_j(s)\), and pulls the arm with the smallest index,
we know that arm \(k\) must have had the smallest index at round \(t\).
We have:

\[
B_k(t) \leq B_i(t) \Leftrightarrow \sqrt{T_k(t)}\hat{\Delta}_k(t) \leq \sqrt{T_i(t)} \hat{\Delta}_i(t)
\]

One can already expect that this can be used to derive a lower bound on
\(T_i(t)\) on the RHS, and this is exactly what we do.

On the LHS, we will plug in the lower bound for \(T_k(t)\) derived
before. Additionally bounding \(\Delta_k(t)\) from below, and
\(\hat{\Delta}_i(t)\) from above using the favorable event will give the
desired lower bound.

Thus assume we are on the favorable event \(\xi\) with large
probability. we consequently have

\[
| \hat{\mu}_i(t) - \mu_i| \leq \sqrt{\frac{T\delta^2}{HT_i(t)}}.
\]

As shown in Locatelli et al. (2016), one can use the reverse triangle
inequality to get:

\[
|\hat{\mu}_i(t) - \mu_i| = |(\hat{\mu}_i(t) - \tau) - (\mu_i - \tau)| \geq || \hat{\mu}_i(t) - \tau | - |\mu_i - \tau|| \geq |(|\hat{\mu}_i(t) - \tau| + \epsilon) - (|\mu_i - \tau| + \epsilon)| \geq |\hat{\Delta}_i(t) - \Delta_i|
\]

If we combine this with the favorable event, we get

\[
\Delta_k - \sqrt{\frac{T\delta^2}{HT_k(t)}} \leq \hat{\Delta}_k(t) \leq \Delta_k + \sqrt{\frac{T\delta^2}{HT_k(t)}}
\]

and similar for \(\hat{\Delta}_i\).

We can now lower bound the LHS of the indices as

\[
\Big(\Delta_k - \sqrt{\frac{T\delta^2}{HT_k(t)}}\Big) \sqrt{T_k(t)} \leq B_k(t)
\]

\[
\Big(\Delta_k - \sqrt{\frac{2}{h_k}}\delta \Big) \sqrt{\frac{Th_k}{2H}} \leq B_k(t)
\]

At this point, it is useful to have a well chosen complexity measure. If
we set \(h_i = \frac{1}{\Delta_i^2}\) as in Locatelli et al. (2016), we
get

\[
\Big(\Delta_k - \sqrt{2}\delta \Delta \Big) \sqrt{\frac{T}{2H\Delta_k^2}} \leq B_k(t)
\]

which can be now be reduced to a form that is not specific to \(k\):

\[
\Big( \frac{1}{\sqrt{2}} - \delta \Big) \sqrt{\frac{T}{H}} \leq B_k(t).
\]

Now we need to get rid of \(\hat{\Delta}_i\) in order to bound
\(T_i(t)\). Thus, we bound the RHS again using the favorable event much
as we did for the LHS:

\[
B_i(t) = \hat{\Delta}_i \sqrt{T_i(t)} \leq \Big(\Delta_i + \sqrt{\frac{T\delta^2}{HT_i(t)}}\Big) \sqrt{T_i(t)} \leq \Delta_i \sqrt{T_i(t)} + \delta \sqrt{\frac{T}{H}}
\]

As in Locatelli et al. (2016) we use the fact that by definition
\(\Delta\) and \(\hat{\Delta}\) are positive in order to square both
sides and write

\[
(1 - 2\sqrt{2}\delta)^2 \frac{T}{2H\Delta^2_i} \leq T_i(t)
\]

We thus have derived a lower bound on the amount of pulls for any arm.
That means that we can show that no arm is pulled an insufficient amount
of times with large probability, and that consequently the favorable
event hold and the estimated mean is sufficienctly separated from the
threshold.

\subsection{Check Separation
Condition}\label{check-separation-condition}

\end{document}