### 2017-08-25 Thesis/Literature.Rmd

---
title: "Literature"
output: pdf_document
---

Hermann Chernoff (1958). Sequential Design of Experiments. The Annals of Mathematical Statistics.

Thompson, W. R. (1933). On the likelihood that one unknown probability exceeds another in view of the evidence of two samples. Biometrika 25 285–294.
Thompson, W. R. (1935). On the Theory of Apportionment. Amer. J. Math. 57 450–456.
MR1507085

Gittins, J. C. (1979). Bandit processes and dynamic allocation indices (with discussion).
J. R. Stat. Soc. Ser. B Stat. Methodol. 41 148–177. MR0547241

Burnetas, A. N. and Katehakis, M. N. (1996). Optimal adaptive policies for sequential allocation problems. Adv. in Appl. Math. 17 122–142. MR1390571

Lai, T. L. and Robbins, H. (1985). Asymptotically efficient adaptive allocation rules.
Adv. in Appl. Math. 6 4–22. MR0776826

Robbins, H. (1952). Some aspects of the sequential design of experiments. Bull. Amer.
Math. Soc. (N.S.) 58 527–535. MR0050246

Capp ́e, O., Garivier, A., Maillard, O.-A., Munos, R. and Stoltz, G. (2013). Supple-
ment to “Kullback–Leibler upper confidence bounds for optimal sequential allocation.”
DOI:10.1214/13-AOS1119SUPP.

Capp ́e, O., Garivier, A., Maillard, O.-A., Munos, R. and Stoltz, G. (2013). Kullback–Leibler upper confidence bounds for optimal sequential allocation.

Emilie Kaufmann, Olivier Cappé, and Aurélien Garivier (2012). On Bayesian Upper Confidence Bounds for Bandit Problems. Proceedings of the 15th International Conference on Artificial Intelligence and Statistics.

Eric Brochu, Vlad M. Cora and Nando de Freitas (2010). A Tutorial on Bayesian Optimization of Expensive Cost Functions, with Application to Active User Modeling and Hierarchical Reinforcement Learning
arXiv:1012.2599v1

Jean-Yves Audibert, Sébastien Bubeck, Rémi Munos (2010). Best Arm Identification in Multi-Armed Bandits.

Russo (2016)

Garivier and Kaufmann (2016)

(Kaufmann and Kalyanakrishnan, 2013;

Kaufmann et al. 2016)

Locatelli et al. (2016)