### 2017-08-28 Thesis/Literature.Rmd

---
title: "Literature"
output: pdf_document
---

Hermann Chernoff (1958). Sequential Design of Experiments. The Annals of Mathematical Statistics.

Thompson, W. R. (1933). On the likelihood that one unknown probability exceeds another in view of the evidence of two samples. Biometrika 25 285–294.
Thompson, W. R. (1935). On the Theory of Apportionment. Amer. J. Math. 57 450–456.
MR1507085

Gittins, J. C. (1979). Bandit processes and dynamic allocation indices (with discussion).
J. R. Stat. Soc. Ser. B Stat. Methodol. 41 148–177. MR0547241

Burnetas, A. N. and Katehakis, M. N. (1996). Optimal adaptive policies for sequential allocation problems. Adv. in Appl. Math. 17 122–142. MR1390571

Lai, T. L. and Robbins, H. (1985). Asymptotically efficient adaptive allocation rules.
Adv. in Appl. Math. 6 4–22. MR0776826

Robbins, H. (1952). Some aspects of the sequential design of experiments. Bull. Amer.
Math. Soc. (N.S.) 58 527–535. MR0050246

Capp ́e, O., Garivier, A., Maillard, O.-A., Munos, R. and Stoltz, G. (2013). Supple-
ment to “Kullback–Leibler upper confidence bounds for optimal sequential allocation.”
DOI:10.1214/13-AOS1119SUPP.

Capp ́e, O., Garivier, A., Maillard, O.-A., Munos, R. and Stoltz, G. (2013). Kullback–Leibler upper confidence bounds for optimal sequential allocation.

Emilie Kaufmann, Olivier Cappé, and Aurélien Garivier (2012). On Bayesian Upper Confidence Bounds for Bandit Problems. Proceedings of the 15th International Conference on Artificial Intelligence and Statistics.

Eric Brochu, Vlad M. Cora and Nando de Freitas (2010). A Tutorial on Bayesian Optimization of Expensive Cost Functions, with Application to Active User Modeling and Hierarchical Reinforcement Learning
arXiv:1012.2599v1

Jean-Yves Audibert, Sébastien Bubeck, Rémi Munos (2010). Best Arm Identification in Multi-Armed Bandits.

Bubeck, S. and Cesa-Bianchi, N. (2012). Regret analysis of stochastic and nonstochastic multi-armed bandit problems. Foundations and Trends in Machine Learning 5 1–122.

Russo (2016)

Garivier and Kaufmann (2016)

(Kaufmann and Kalyanakrishnan, 2013;

Kaufmann et al. 2016)

Locatelli et al. (2016)

Zhong et al. (2017)

Bubeck, Sébastien, Cesa-Bianchi, Nicolo, and Lugosi, Gábor. Bandits with heavy tail. Information Theory, IEEE Transactions on, 59(11):7711-7717, 2013a.

Mukherjee et al. (2017)

Victor Gabillon, Mohammad Ghavamzadeh, Alessandro Lazaric, and Sébastien Bubeck. Multi-bandit best arm identification. In *Advances in Neural Information Processing Systems, pages 2222-2230, 2011.

Shouyuan Chen, Tian Lin, Irwin King, Michael R Lyu, and Wei Chen. Combinatorial pure exploration of multi-armed bandits. In *Advances in Neural Information Processing Systems, pages 379-387, 2014.

"Thompson Sampling: An Asymptotically Optimal Finite Time Analysis" by Kaufmann, Korda and Munos (2012)

Lihong Li, Wei Chu, John Langford, Robert E. Schapire. A Contextual-Bandit Approach to Personalized News Article Recommendation. WWW 2010, April 26–30, 2010, Raleigh, North Carolina, USA. ACM 978-1-60558-799-8/10/04.