### 2017-09-03 Thesis/Corollary.Rmd
Add corollary for lower bound

---
title: "Corollary"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 12pt
urlcolor: blue
---

\begin{corollary}
Let $\nu = \nu_1$ be a one-armed thresholding bandit model where $\tau$ is the threshold fixed upfront. Let $\nu$ be from the univariate canonical exponential family of distributions such that it can be parameterized by its mean, $\mu_1$, and the Kullback-Leibler divergence between distributions $\nu_1$ and $\nu_1'$ can be written as $\KL(\nu_1', \nu_1) = \KL(\mu_1', \mu_1)$. In the fixed budget setting, any consistent algorithm satisfies:

$$
\lim_{t \rightarrow \infty}- \frac{1}{t} \log p_t(\nu) \leq \KL(\tau, \mu_1)
$$
\end{corollary}

**Proof**: The proof follows closely the proof of Theorem