### 2017-09-03 Thesis/CorollaryProof.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Corollary},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Corollary}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[lined,boxed]{algorithm2e}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}
\newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%

\begin{document}
\maketitle

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\begin{corollary}
Let $\nu = \nu_1$ be a one-armed thresholding bandit model where $\tau$ is the threshold fixed upfront. Let $\nu$ be from the univariate canonical exponential family of distributions such that it can be parameterized by its mean, $\mu_1$, and the Kullback-Leibler divergence between distributions $\nu_1$ and $\nu_1'$ can be written as $\KL(\nu_1', \nu_1) = \KL(\mu_1', \mu_1)$. In the fixed budget setting, any consistent algorithm satisfies:

$$
- \frac{1}{t} \log p_t(\nu) \leq \KL(\tau, \mu_1)
$$
\end{corollary}

\textbf{Proof}: The proof follows closely the proof of Theorem 12 in
Kaufmann et al. (2016) and is again an application of their
transportation Lemma, Lemma 2.

Consider the one-armed bandit problem, \(K = 1\) with the canonical
exponential family of distributions distribution \(\nu_1\) parameterized
by its mean \(\mu_1\). Furthermore, given we are in the thresholding
bandit problem, we have a threshold \(\tau\) against which the arm is
classified. Assume w.l.o.g. that for \(\nu_1\), \(\mu_1 \geq \tau\).
Thus, \(\mathcal{S}_{\tau}^* = \{1\}\). After \(T\) rounds, a consistent
algorithm returns with error probability \(p_T(\nu)\) the guess
\(\hat{\mathcal{S}}_{\tau}\). Thus, we have for the event
\(A = (\hat{\mathcal{S}}_\tau = \{1\})\) the probability
\(\mathbb{P}_{\nu}(A)\). On the other hand, define the alternative model
\(\nu'\) with mean \(\mu_1' < \tau\) such that the correct
classification from \(\nu\) is now wrong:
\(\mathcal{S}_{\tau}^* {'} = (\mathcal{S}_{\tau}^*)^C = \{\emptyset\}\).
Consequently, since the algorithm is still consistent on this problem,
we have that \(\mathbb{P}_{\nu'}(A) = p_t(\nu')\). Thus, we expect that
as we let \(t\) increase, \(\mathbb{P}_{\nu'}(A)\) will decrease while
\(\mathbb{P}_{\nu}(A)\) will increase.

Indeed, given we consider only consistent algorithms, there should exist
a number of samples for which the probability of the correct decision in
problem \(\nu\) is larger than the error probability in problem
\(\nu'\), that is: for every \(\epsilon > 0\), there exists
\(t_0(\epsilon)\) such that for all \(t \geq t_0(\epsilon)\) we have

\begin{equation*}
\mathbb{P}_{\nu'}(A) \leq \epsilon \leq \mathbb{P}_{\nu}(A). \label{eq:BoundForCorollary}
\end{equation*}

Now, consider again Lemma 2 in Kaufmann et al. (2016). If we let the
algorithm be such that \(T = t\), we can apply the Lemma to the stopping
time \(\sigma = t\) a.s. and the event \(A\) to get

\[
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)).
\]

We can lower bound the RHS by using the bound on
\(\mathbb{P}_{\nu'}(A)\) given in \eqref{eq:BoundForCorollary} and the
monotonicity of the binary relative entropy. We get:

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) & = t \KL(\nu_1', \nu_1) \\
& \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)) \\
& \geq d(\epsilon, 1- p_t(\nu)) \\
& = \epsilon \log\Big(\frac{\epsilon}{1-p_t(\nu)}\Big) + (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) \\
& \geq (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) + \epsilon \log(\epsilon) \\
\end{align*}

Letting \(\epsilon \rightarrow 0\), we get

\[
t \KL(\nu_1', \nu_1) \geq \log \Big(\frac{1}{p_t(\nu)}\Big) = -\log p_t(\nu).
\] To make the bound as tight as possible, we choose the alternative
model that minimizes \(\KL(\nu_1', \nu_1)\) while still ensuring that
\(\mathcal{S}_{\tau}^* {'} = \{\emptyset\}\). In our case of the
thresholding bandit with univariate exponential family distributions,
this is given by choosing \(\mu_1' = \tau\). By pluggin in, we get the
result:

\[
t \KL(\tau, \mu_1) \geq -\log p_t(\nu).
\]

\begin{corollary}
Let $\nu = (\nu_1, ..., \nu_K)$ be a $K$-armed thresholding bandit problem in which $\tau$ is the threshold fixed upfront. For $i \in \{1,...,K\}$, let $\nu_i$ be from the univariate canonical exponential family of distributions such that it can be parameterized by its mean, $\mu_i$, and the Kullback-Leibler divergence between distributions $\nu_i$ and $\nu_i'$ can be written as $\KL(\nu_i, \nu_i') = \KL(\mu_i, \mu_i')$. In the fixed budget setting, every consistent algorithm satisfies:

$$
\lim_{t \rightarrow \infty} -\frac{1}{t} \log p_t(\nu) \leq \max_{a \in \{1,...,K\}} \KL(\tau, \mu_a)
$$
\end{corollary}

\textbf{Proof}: The proof is very similar to the proof of the corollary
for \(K=1\). However, the alternative model needs more motivation than
in the case of a single arm. But first, consider again that we are in
the thresholding bandit problem with \(K\) arms of univariate canonical
exponential family distributions that are parameterized by their means
\(\mu_1, ..., \mu_K\) for the original problem
\(\nu = (\nu_1, ..., \nu_K)\). W.l.o.g, assume that all arms are above
the threshold, that is, \(\mu_1 \geq \tau, ..., \mu_K \geq \tau\). Then
\(\mathcal{S}_{\tau}^* = \{1, ... , K\}\). The correct classification
returned after \(t\) rounds is thus the event
\(A = (\hat{\mathcal{S}}_{\tau} = \{1, ..., K\})\). We can now consider
Lemma 2. The RHS is maximized when the alternative model is chosen such
that \(d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A))\) is maximized.
Given the event \(A\) as defined before, this is the case when
\(\mathbb{P}_{\nu'}(A)\) is as small as possible. For consistent
algorithms, this is the case when \(\nu'\) is chosen such that all
\(\mu_1' < \tau, ..., \mu_K' < \tau\). Then even classifying a single
arm correctly under \(\nu\) would lead to a mistake in \(\nu'\).
Consequently, \(A\) is the most unlikely classification for a consistent
algorithm for the problem \(\nu'\). We thus expect for any consistent
algorithm that \(\mathbb{P}_{\nu}(A)\) becomes large and
\(\mathbb{P}_{\nu'}(A)\) becomes small as the number of observations
\(t\) increases. Again, for every \(\epsilon > 0\), there exists
\(t_0(\epsilon)\) such that for all \(t \geq t_0(\epsilon)\) we have
\(\mathbb{P}_{\nu'}(A) \leq \epsilon \leq \mathbb{P}_\nu(A)\). This
allows us to now use Lemma 1 as previously:

\begin{align*}
\sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\nu_i', \nu_i) & \geq \sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\tau, \nu_i) \\
& \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)) \\
& \geq d(\epsilon, 1-p_t(\nu)) \\
& = \epsilon \log \Big(\frac{\epsilon}{1-p_t(\nu)}\Big) + (1-\epsilon) \log \Big( \frac{1-\epsilon}{p_t(\nu)}\Big) \\
& \geq (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) + \epsilon \log(\epsilon)
\end{align*}

We let \(\epsilon \rightarrow 0\), divide by \(t\) and take the limsup
on both sides as in Kaufmann et al. (2016):

\begin{align*}
\lim_{t \rightarrow \infty} \sup - \frac{1}{t} \log p_t(\nu) & \leq \lim_{t \rightarrow \infty} \sup \frac{1}{t} \sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\tau, \nu_i) \\
& \leq \max_{i \in \{1,...,K\}} \KL(\tau, \nu_i)
\end{align*}

\end{document}