### 2017-09-10 Thesis/Draft.Rmd
Add References

---
title: "On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 11pt
urlcolor: blue
---

\newpage

\vspace*{10\baselineskip}

"In general, we look for a new law by the following process: First we guess it. Then we – now don't laugh, that's really true. Then we compute the consequences of the guess to see what, if this is right, if this law that we guessed is right, to see what it would imply. And then we compare the computation results to nature, or we say compare to experiment or experience, compare it directly with observations to see if it works. If it disagrees with experiment, it's wrong. In that simple statement is the key to science. It doesn't make any difference how beautiful your guess is, it doesn't make any difference how smart you are, who made the guess, or what his name is. If it disagrees with experiment, it's wrong. That's all there is to it."

- Richard Feynman in *Richard Feynman Messenger Lectures at Cornell University: The Character of Physical Law. Chapter 7, Seeking New Laws*.

\newpage

# Introduction \label{chap:Introduction}

The thesis at hand joins a large literature concerned with the problem of sequential sampling. One is familiar with the idea of "batch sampling", where the entire sample is gathered first and analyzed only after the last sample has become available. However, as noted in Chernoff (1958), many situations arise in which one would like to gather feedback from samples as they are drawn, so as to influence the choice of the next sample. The two fundamental characteristics of the latter sampling process are that it is sequential and adaptive. Since observations are drawn in sequence, one has a chance to analyze past observations in order to adapt the decision of what sample one would like to observe next.

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples, or where only a limited amount of samples can be collected. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that as soon as possible more patients receive the better drug and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions (which we call arms in the multi-armed bandit problem in analogy with the one-armed bandit slot machine). These distributions have different means, and one can draw samples sequentially from the distributions in order to estimate their means. To minimize the cumulative regret measure, one would like to play the distribution with the highest mean as much as possible (*Exploitation*). At the same time, however, one has to sample all arms at least a little, so as to initially find the best arm (*Exploration*). Any algorithm thus has to handle a trade-off between exploration and exploitation to minimize the regret. How can we organize the sampling in order to achieve this task?

While Gittins (1979) presents an Bayesian-optimal solution to the cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002) have lead to a blooming literature on multi-armed bandit problems that has been picked up to solve modern problems in website and advertising optimization, where advertisers try to allocate their advertising optimally so as to reduce regret from showing poorly converting advertisements. Several algorithms like the KL-UCB algorithm based on Kullback-Leibler divergence upper confidence intervals (Cappé et al., 2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012)  have followed and been shown to be optimal as they reach the lower bound introduced in Lai and Robbins (1985) and Burnetas and Katehakis (1996).

These algorithms have in common that they compute after every new sample an index for each of the unknown distributions. This index is based solely on the past (observed) data, and then used to pick the distribution from which the next sample should be drawn. Using these indexes, algorithms can sample sequentially and adapt their strategies to observations as they are collected.

Similar algorithms based on indexes can be used on related problems. Instead of minimizing cumulative regret, it can be of interest to find the best arm, that is, the distribution with the largest mean -- with as few samples as possible and with confidence as large as possible. This is particularly useful when there is not a cost from testing arms with lower mean, but when drawing any sample is costly. Thus one needs to explore the distributions as efficiently as possible. Exploitation is not necessary, and the setting is thus referred to as *Pure Exploration*. The need for strategies other than uniform sampling arises: One does not want to continue to sample distributions that one is confident in. This setting is related to the problem faced in Bayesian Optimization: Training large neural networks is costly in terms of time and computation, and so the parameter space needs to be explored with as few samples as possible. (Zhang et al., 2017; Brochu et al., 2010). The problem of best arm identification is also very related to statistical hypothesis testing as noted by Chernoff (1958) since one has to make a decision about how the means of distributions relate to each other. The main difference being that the bandit setting includes the testing procedure in the sampling process.

Given the focus on exploration, the pure exploration problem is usually described with a loss function different from the one in the cumulative regret settings. Audibert et al. (2010) introduce a soft measure called *simple regret* based on which they describe UCB policies. Other authors investigate binary measures that are equal to 1 when the best arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016). These binary measures correspond to the probability of the identified arm being the best, and thus inherently related to hypothesis testing. The problem of best arm identification can readily be extended to the so called Top-$m$ problem of identifying the best $m$ of $K$ arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the name is reminiscient of the cumulative regret multi-armed bandit, the problem is more similar to the top-$m$ problem. But instead of identifying the best $m$ out of $K$ arms, one aims to find all distributions out of $K$ that have a mean at least as large as a fixed threshold. Given that an implicit threshold arises in the top-$m$ problem between the $m$-th best and the $(m+1)$-th best arm (Kaufmann et al., 2016), the two problems share many characteristics. We discuss this in more detail in \autoref{chap:KullbackLeiblerPureExploration}. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above or below the threshold, and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

In the following, we first present the thresholding bandit problem in detail in \autoref{chap:ThresholdingBanditProblem}. To then extend upon the algorithm proposed by Locatelli et al. (2016) and discussed in \autoref{sec:TheAPTAlgorithm}, we first explore in \autoref{chap:KullbackLeiblerPureExploration} how the Kullback-Leibler divergence has been employed in the best arm identification literature by Garivier and Kaufmann (2016) and Kaufmann et al. (2016) as this will lead us to our own extension. We shortly describe in \autoref{sec:VarianceBasedAlgorithms} two variance-based algorithms for the thresholding bandit problem introduced in Mukherjee et al. (2017) and Zhang et al. (2017), before introducing the new Simple Likelihood Ratio algorithm in \autoref{sec:LRforUnivariateExponentialFamily}. We end in \autoref{chap:Experiments} with experiments to examine the performance of different algorithms.

\newpage

# The Thresholding Bandit Problem \label{chap:ThresholdingBanditProblem}

## An Example

To introduce the thresholding bandit problem, consider a common situation among internet-based companies. These companies often collect user feedback to offer reviews and ratings on websites or in apps, or collect it to improve their services. Netflix lets users rate movies with "Thumbs up" and "Thumbs down", apps like Clue ask users a question per day to learn to predict the user's health and mood. An online shop might give customers a one-question survey after they make a purchase. Finally, Foursquare and Yelp ask users to rate coffee shops, bars, and restaurants.

Consider the problem the latter companies face. Whenever a user reports that she visited a restaurant, the app would like to ask questions about her experience. To give a simple example, assume users are not willing to answer more than one question per restaurant visit, and the app would like to find out two things: Does the restaurant have good service? Is the restaurant trendy?

If the feedback that users can give is restricted to binary "Yes" or "No" answers, then we can model the feedback as Bernoulli random variables

$$
X_{\text{service}} \sim \text{Ber}(\mu_{\text{service}}) \quad \text{and} \quad X_{\text{trendy}} \sim \text{Ber}(\mu_{\text{trendy}}).
$$

The app will then display a "Good service" or "This place is trendy" description for the restaurant if $\mu_{\text{service}}$ and $\mu_{\text{trendy}}$ are estimated to be above a certain threshold, for example $\tau = 0.5$. That is, more than half of the people would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information should be shown as quickly as possible. But even though the app has thousands of users, not that many people visit a specific restaurant, and even less people report this back to the app. Thus the company only has a limited budget of $T$ potential users from which to collect samples $X_{\text{service}}$ and $X_{\text{trendy}}$ from. How can the company ask these $T$ questions to be as confident as possible in its classification, that is, in the information it displays?

A straightforward way of sampling users would be to ask the two questions in a round-robin fashion: Ask for "Service", then for "Trendy", then for "Service", ... Given 100 users, 50 users give feedback on the service level, 50 users report on the trendiness level. This however is unlikely to be optimal with respect to the confidence after 100 samples. Given that the company asks the users sequentially, there is the option to adapt the sampling strategy over time: If the first ten users report $\sum_{s=1}^{10}X_{\text{service},s}=10$, while $\sum_{s=1}^{10}X_{\text{trendy},s}=5$, then the company might already be quite confident that $\mu_{\text{service}} \geq \tau = 0.5$, while this is less clear for $\mu_{\text{trendy}}$. Instead of continuously sampling both questions, it might be more useful to collect more feedback on the trendiness level so as to figure out its position relative to the threshold. After all, the company would like to use the 100 samples to maximize its confidence in both displayed tips.

Finding the optimal sequential sampling strategy to classify distributions against a threshold with as much confidence as possible after a limited budget of samples is the problem described above, and the problem described by the thresholding bandit problem which we discuss for the remainder of this thesis. In the following section, we define the general setting of the thresholding bandit problem.

## General Setting \label{sec:Setup}

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, she would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms, where $X_{i} \sim \nu_i$. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which she would like to observe the reward $X_{I_t,t}$. Often, adaptive sampling strategies makes the decision of which arm to pull by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then is $B_i(t)$ a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

Which index proves to be useful is determined by the loss function with which the decisions of the algorithm are evaluated. Many problems are concerned with finding the arm with the largest mean $\mu_i^*$ while collecting loss either cumulatively during (Auer and Cesa-Bianchi, 2012) or once at the end of the sampling (Kaufmann et al., 2016). The thresholding bandit problem considers a slightly different loss function. Consider a fixed threshold value $\tau \in \mathbb{R}$ and an epsilon interval $\tau \pm \epsilon$ around the threshold for $\epsilon \geq 0$. As mentioned in the introduction, the learner is tasked with classifying arms as having a mean $\mu_i$ larger or smaller than $\tau$. As in Locatelli et al. (2016), whenever we say that "an arm is over the threshold", we say that the mean of the arm is over the threshold. To describe which arms are over the threshold, we also introduce the following set notation: Let $\mathcal{S}_{u}$ be the set of distributions in $\mathbb{A}$ with mean larger than some $u \in \mathbb{R}$: $\mathcal{S}_{u} := \{i \in \mathbb{A}, \mu_i \geq u\}$. Consequently, the complement is given by the arms with mean below $u$, $\mathcal{S}^C_{u} := \{i \in \mathbb{A}, \mu_i < u\}$.

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner returns a set of arms that she believes to be over the threshold. When evaluating the classification of the learner, arms with mean inside of the epsilon interval $\tau \pm \epsilon$ will not be considered. The learner only has to correctly distinguish the means from the threshold up to the precision given by $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$.

The learner's classification is returned as the set $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subset \mathbb{A}$ after $T$ rounds; that is, the set of arms estimated to be over the threshold. The classification is evaluated against the loss given as (Locatelli et al., 2016):

$$
\mathcal{L}(T) = \mathbbm{1}{\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big)}
$$

In order to perform well, a learner has to identify all arms that are over the threshold as such, and to not claim arms as being over the threshold if they actually are below the threshold. Due to the precision $\epsilon$, the loss is not influenced by arms which are very close to the threshold $\tau$ as measured by $\epsilon$, that is, arms inside of the $\tau \pm \epsilon$ interval. If either $\hat{\mathcal{S}}_{\tau}$ contains an arm below $(\tau - \epsilon)$, or $\hat{\mathcal{S}}^C_{\tau}$ contains an arm above $(\tau + \epsilon)$, a loss of $1$ is accrued. Else, all arms outside the epsilon interval are classified correctly and the loss is equal to $0$. In the case of $\epsilon = 0$, all arms need to be classified correctly to not collect any loss.

Algorithms for this thresholding problem can be compared based on their expected loss. Let $\mathbb{E}$ be the expectation with regard to the samples $X_{i,1}, ..., X_{i,T}$ collected by the algorithm. The expected loss is then given as

$$
\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor(\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big),
$$

which corresponds to the probability of making at least one wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not, are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting can be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting we have just introduced, as well as an upper bound on the performance of their APT algorithm which we will introduce in the following chapter. As in their paper, we define $R$-sub-Gaussian and assume it in the description of the their algorithm and bounds:

\begin{definition}[$R$-sub-Gaussian distribution] 
\label{definition:rsubgaussian}
Let $R > 0$. A distribution $\nu$ is $R$-sub-Gaussian if for all $t \in \mathbb{R}$ we have
\begin{equation*}
\mathbb{E}_{X \sim \nu}[\exp(tX - t\mathbb{E}[X])] \leq \exp(R^2t^2/2).
\end{equation*}
\end{definition}

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions offer concentration inequalities for the concentration of the empirical mean around the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. The family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the assumption of $R$-sub-Gaussianity, we can now proceed to describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016) for the thresholding bandit problem.

## The Anytime Parameter-free Thresholding Algorithm \label{sec:TheAPTAlgorithm}

In the following we describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016), which is a simple yet effective algorithm and displays some very favorable characteristics when compared to alternative algorithms.

We fix the threshold $\tau$ and epsilon interval around the threshold given by $\epsilon$ upfront, and supply them as arguments to the algorithm: $(\tau, \epsilon)$. Based on the general thresholding problem, we demand that the algorithm draws samples $X_{i,t}$ for $T$ rounds from the distributions $\nu_i, i\in \mathbb{A}$. For a given arm $i$ and a given time $t$, we let $T_i(t)$ represent the number of samples drawn from arm $i$ until time $t$ (including). Consequently, the empirical mean of arm $i$ at time $t$ is given by $\hat{\mu}_i(t) = \frac{1}{T_i(t)} \sum_{s=1}^{T_i(t)} X_{i,s}$ where $X_{i,s}$ is the feedback sampled from distribution $\nu_i$ when the arm is pulled for the $s$-th time. In the case of the APT algorithm, the decision $\hat{\mathcal{S}}_{\tau}$ is based on the empirical means $\hat{\mu}_i(T)$ estimated from the samples drawn up to time $T$: $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) = \{i \in \mathbb{A}, \hat{\mu}_i(T) \geq \tau\}$.

As we see in algorithm \ref{alg:APT}, the APT algorithm initiates by pulling each arm once. Then, for each $t > K$, it calculates for every arm the values $T_i(t)$ and $\hat{\mu}_i(t)$. Based on the empirical mean, the algorithm computes the estimated gap between the mean and the threshold as $\hat{\Delta}_i(t) = \hat{\Delta}_i^{\tau, \epsilon}(t) = |\mu_i(t) - \tau| + \epsilon$. The decision of which arm is sampled next is then based on the value $B_i(t+1) = \sqrt{T_i(t)} \hat{\Delta}_i(t)$: the algorithm pulls arm $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$ minimizing this decision quantity. After overall $T$ pulls, the algorithm returns the sets described in the general setup, $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}^C_{\tau}$.

\IncMargin{1em}
\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t)} (|\hat{\mu}_i(t) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}
\DecMargin{1em}

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms that are estimated to be further away from the threshold $\tau$ as measured by the absolute distance to their empirical mean $\hat{\mu}_i$, $\hat{\Delta}_i(t)$, are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to less often pull arms which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easier to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we would know whether to expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic as it is not known upfront), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

$$
\Delta_i := \Delta_i^{\tau,\epsilon} = |\mu_i - \tau| + \epsilon
$$

Given this gap, Locatelli et al. (2016) define the complexity $H_{\tau, \epsilon}$ for the thresholding bandit problem naturally as

$$
H := H_{\tau, \epsilon} = \sum_{i=1}^K \frac{1}{(\Delta_i^{\tau,\epsilon})^2}.
$$

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require less precision, that is, we increase $\epsilon$, the hardness decreases. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we do not know neither $\mu_i$ nor $H$ upfront, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

At this point, one can question whether the measure of complexity $H$ proposed in Locatelli et al. (2016) is a good measure of the "true" problem complexity. In the following \autoref{sec:UpperBoundLocatelli}, we will show the upper bound on the probability of error for the APT algorithm and see how the complexity $H$ arises naturally for the APT algorithm. In \autoref{sec:LowerBoundLocatelli} we then show the lower bound presented in Locatelli et al. (2016) and that the APT algorithm is optimal in that it matches the lower bound. We discuss the results and the APT algorithm further in \autoref{sec:DiscussionLocatelli}.

## Upper Bound for the APT Algorithm \label{sec:UpperBoundLocatelli}

In this chapter we present the upper bound on the probability of error that the APT algorithm incurs, as proven in Locatelli et al. (2016). The upper bound is on the probability of error that the algorithm has after $T$ rounds of sampling for a problem with arms that are assumed to be $R$-sub-Gaussian. The bound depends also on the complexity of the problem, and we show in the proof that given the index of the APT algorithm as defined in \autoref{sec:TheAPTAlgorithm}, the complexity $H$ is an appropriate measure.

We now state the theorem as in Locatelli et al. (2016) and present a proof.

\begin{theorem}[Theorem 2, Locatelli et al., 2016] \label{theorem:LocatelliTheorem4}
Let $K \geq 0$, $T \geq 2K$, and consider a problem $\mathcal{B}$. Assume that all arms $\nu_i$ of the problem are R-sub-Gaussian with means $\mu_i$. Let $\tau \in \mathbb{R}, \epsilon \geq 0$. Algorithm APT's expected loss is upper bounded on this problem as 
\begin{equation*} \mathbb{E}(\mathcal{L}(T)) \leq \exp
\Big(-\frac{1}{64R^2}\frac{T}{H} + 2 \log((\log(T) + 1)K)\Big) \end{equation*}

where we remind that $H = \sum_i (|\mu_i - \tau | + \epsilon)^{-2}$ and where
$\mathbb{E}$ is the expectation according to the samples of the problem.
\end{theorem}

### Proof of Theorem \ref{theorem:LocatelliTheorem4} \label{sec:UpperBoundProof}

In order to upper bound the expected loss of the APT algorithm, we want to define an event on which the algorithm correctly seperates arms from the threshold, that is, a separation condition holds on this event. We then want to show that this *favorable* event does indeed hold with large probability. Consequently, APT's probability of error will correspond to the probability that the favorable event does not hold, as this leads to a failure of the seperation condition, and consequently to an incorrect classification of at least one arm by the algorithm. We start by defining the separation condition.

### Separation Condition

As noted in Zhong et al. (2017), a simple requirement to correctly classify arms at the final round is the following. For every arm $i \in \{1, ..., K\}$ after the final round $T$, require
$$
\hat{\mu}_{i}(T) \geq \mu_i - \frac{\mu_i - \tau + \epsilon}{2} \quad \text{if} \quad \mu_i \geq \tau
$$
or
$$
\hat{\mu}_{i}(T) \leq \mu_i + \frac{\tau - \mu_i + \epsilon}{2} \quad \text{if} \quad \mu_i < \tau
$$
Or simply in terms of the gap between the arm and the threshold as used in the index by the APT algorithm:
$$
| \hat{\mu}_{i}(T) - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
$$

Intuitively, this condition makes sense as it ensures that the current mean estimate is closer to the true mean than to the threshold. Consequently, there is no arrangement possible in which the arm is classified wrong while the separation condition holds. Assuming $\mu_i < \tau$ and $\epsilon = 0$, an incorrect classification would require $\mu_i < \tau \leq \hat{\mu_i}$. But then  $| \hat{\mu}_{i}(T) - \mu_i | \geq |\mu_i - \tau|$, which violates the separation condition.

Thus, we need to show that this condition holds for every arm $i \in \{1,...,K\}$ with large probability after the final round $T$. In different words, we need to derive an upper bound on the probability with which this condition does not hold.

### The Favorable Event

The separation condition describes what event needs to hold in order for the APT algorithm to make a correct classification: The smaller the deviation of the empirical mean from the population mean, the more difficult it is for the classification to be wrong. Moreover, if the mean and empirical mean are identical, then the classification is correct independent of the threshold. Also, the closer the mean to the threshold, the better the estimate has to be for a correct classification to be likely. To bound the absolute deviation of the empirical mean from the true mean, we define the favorable event $\xi$ as

$$
\xi = \Big\{\forall i \in \mathbb{A}, \forall s \in \{1,...,T\} : |\frac{1}{s} \sum_{t=1}^{s}X_{i,t} - \mu_i| \leq \sqrt{\frac{T \delta^2}{H s}} \Big\}.
$$

The probability that this favorable event occurs should be large, or equivalently, the probability that the deviation is small should be large. Then it follows that the separation condition holds with large probability.

We know that $X_i$ are $R$-sub-Gaussian by assumption. That is, we have

$$
\mathbb{E}[\exp (tX_{i,s}-t\mathbb{E}[X_{i,s}])] = \mathbb{E}[\exp (t(X_{i,s}-\mathbb{E}[X_{i,s}]))] \leq \exp (R^2t^2/2)
$$

where the RHS is the moment-generating function of the random variable $X_{i,s}-\mathbb{E}[X_{i,s}]$. While Höffding's Lemma and Höffding's Inequality are commonly applied to bounded random variables $Z$ with $\mathbb{P}(a\leq Z \leq b) = 1$ and $\mathbb{E}[Z] = 0$, we see that $Z$ are $R$-sub-Gaussian with $R^2 = \frac{(a-b)^2}{4}$. Using this last equality, we can easily apply Höffding's Inequality to our case.

\begin{lemma}[Höffding's Inequality (Kloft Lecture Notes)] \label{lemma:HoeffdingsInequality}
Let $X_1, ..., X_s$ be independent, identically distributed random variables with mean $\mathbb{E}[X_t] = \mu$ and $\mathbb{P}(a_t \leq X_t - \mathbb{E}[X] \leq b_t) = 1$. Denote their empirical mean by $\bar{X}_s=\frac{1}{s}\sum_{t=1}^{s}X_t$. Then, for any $\gamma > 0$, we have the following concentration inequality:

\begin{equation*}
\mathbb{P}(|\bar{X}_s - \mu| \geq \gamma) \leq 2\exp (-2s^2\gamma^2/\sum_{t=1}^s(b_t - a_t)^2).
\end{equation*}
\end{lemma}

If we replace $\sum_{t=1}^s(b_t - a_t)^2$ by $4sR^2$, we have the concentration inequality for our $R$-sub-Gaussian random variables:

$$
\mathbb{P}(|\bar{X}_s - \mu| \geq \gamma) \leq 2\exp (-\frac{s\gamma^2}{2R^2}).
$$

Writing $\gamma = \sqrt{\frac{T\delta^2}{Hs}}$, we get for the favorable event $\xi_i = \{|\bar{X}_s - \mu| \leq \sqrt{\frac{T\delta^2}{Hs}}\}$ of an individual arm $i$, for each $i \in \mathbb{A}$ and each $s \in \{1, ..., T\}$:

$$
\mathbb{P}(\xi_i) = \mathbb{P}\Big(|\bar{X}_{i,s} - \mu_i| \leq \sqrt{\frac{T\delta^2}{Hs}}\Big) \geq 1 - 2\exp (-\frac{T\delta^2}{2R^2H}),
$$
where $T$ is again the total budget of samples, and $H$ the complexity. Now let

$$
\tilde{\xi} = \Big\{\forall i \in \mathbb{A}: |\frac{1}{s} \sum_{t=1}^{s}X_{i,t} - \mu_i| \leq \sqrt{\frac{T \delta^2}{H s}} \Big\}.
$$

To derive the probability of event $\tilde{\xi}$, we apply a union bound on $\mathbb{P}(\cup_{i=1}^K \xi_i^C)$:

\begin{align*}
\mathbb{P}(\tilde{\xi}) & = 1-\mathbb{P}(\xi^C) \\
& = 1-\mathbb{P}(\cup_{i=1}^K \xi_i^C) \\
& \geq 1 - \sum_{i=1}^K \mathbb{P}(\xi_i^C) \\
& \geq 1 - 2K\exp (-\frac{T\delta^2}{2R^2H})
\end{align*}

We could use the same union bound a second time to bound over all $s \in \{1, ..., T\}$ and derive for the favorable event $\xi$:

$$
\mathbb{P}(\xi) \geq 1 - 2TK \exp(-\frac{T\delta^2}{2R^2H})
$$

Locatelli et al. (2016) use instead of Höffding's inequality a Sub-Gaussian martingale inequality including a peeling argument in order to derive a better bound of

$$
\mathbb{P}(\xi) \geq 1 - 2(\log(T) +1)K \exp(-\frac{T\delta^2}{2R^2H})
$$

This bound is non-trivial for much smaller values of $T$, and we will use it in the following.

### Need to Bound $T_i$

We now have that with large probability the favorable event holds. On the favorable event, we have for every arm $i$:

$$
| \hat{\mu}_i(t) - \mu_i| \leq \sqrt{\frac{T\delta^2}{HT_i(t)}}
$$

The deviation of arm $i$ at time $t$ depends on how large $H$, $T$, and $T_i(t)$ are. While $t$, $H$, $T$ are deterministic, the number of times arm $i$ has been pulled at round $t$, $T_i(t)$, is a random variable. In order to know how large the probability of the favorable event is (and thus the error probability), we need to know the value of $T_i(t)$, or at least bound a range of potential values. Put differently: To ensure a deviation smaller than $\sqrt{\frac{T\delta^2}{HT_i(t)}}$ for arm $i$, we need to ensure that arm $i$ has been pulled at least $T_i(t)$ times. As usual in multi-armed bandits, and different from uniform sampling, the number of pulls is a random variable that also depends on how often other arms are pulled. In general, we do not want each arm to be pulled an equal amount, but each arm should be pulled some times, and not too many times.

Thus, in order to check that the deviation of arm $i$ really is small enough for the separation condition to hold on the favorable event, we need to bound the number of pulls of every arm from below.

### Characteristics of a Frequently Pulled Arm (Lower Bound on $T_k(T)$)

Given that we do not follow a uniform sampling strategy, we know that at time $T$, some arms will have been pulled more frequently than others. In particular, some arms will have been pulled more than $T/K$ times. How often an arm is pulled can be related to the individual arm's contribution to the overall complexity measure. Let $h_i$ be a measure of arm $i$'s individual problem complexity and write for the overall problem complexity $H = \sum_{i = 1}^{K} h_i$. Then arm $i$ contributes a share of $h_i/H$ to the overall problem complexity. We can now show that there exists an arm that has been pulled at least as often as the share of its complexity contribution.

Given the initialization of the APT algorithm, every arm is pulled once initially. Consequently, there is a budget of $T-K$ rounds left after the initialization. As shown in Locatelli et al. (2016), there exists an arm $k$ that has been pulled after the initialization ($\geq 2$ pulls) and for which at the final round $T$ it holds: $T_k(T) - 1 \geq \frac{(T-K)h_k}{H}$. This can be shown by a contradiction. Assume that the opposite holds, that is

$$
\forall i \in \{1,...,K\}: T_i(T) - 1 < \frac{(T-K)h_i}{H}
$$

Consequently, at time $T$, we would have pulled across all arms $\sum_{i=1}^KT_i(T)-1$ times. If the previous assumption holds, we would not have used the entire budget:

$$
\sum_{i=1}^K T_i(T)-1 < \sum_{i=1}^K \frac{(T-K)h_i}{H} = (T-K)\sum_{i=1}^K \frac{h_i}{H} = T-K
$$

This is a contradiction, since as an requirement to the algorithm we have that $\sum_{i=1}^K T_i(T)-1 = (T-K)$, that is, we use the entire budget of samples.

Furthermore, we have assumed $T>2K$ in the statement of Theorem \ref{theorem:LocatelliTheorem4}. Plugging this in gives 
$$
T_k(T) - 1 \geq \frac{(T-K)h_k}{H} \geq \frac{(T-1/2T)h_k}{H} = \frac{Th_k}{2H}
$$

Now we know how often the arm has been pulled at the very end. That does not mean that $k$ has been pulled at time $T$. Thus we consider the time $t \leq T$ at which arm $k$ is pulled for the last time. We get

$$
T_k(t) \geq T_k(T) - 1 \geq \frac{Th_k}{2H}.
$$

This gives us a lower bound on how often arms are pulled that are pulled frequently relative to their contribution to the problem complexity. Now we need a lower bound on $T_i(T)$ for every other arm $i$.

### Lower Bound on $T_i(T)$

Consider again as before the round $t$ at which arm $k$ is pulled for the very last time. Since the APT algorithm minimizes at every round $s \geq K+1$ the index $B_j(s)$, and pulls the arm with the smallest index, we know that arm $k$ must have had the smallest index at round $t$. We have:

$$
B_k(t) \leq B_i(t) \Leftrightarrow \sqrt{T_k(t)}\hat{\Delta}_k(t) \leq \sqrt{T_i(t)} \hat{\Delta}_i(t)
$$

One can already expect that this can be used to derive a lower bound on $T_i(t)$ on the RHS, and this is exactly what we do. On the LHS, we plug in the lower bound for $T_k(t)$ derived before. Additionally bounding $\Delta_k(t)$ from below, and $\hat{\Delta}_i(t)$ from above using the favorable event will give the desired lower bound.

Thus assume we are on the favorable event $\xi$ with large probability. we consequently have

$$
| \hat{\mu}_i(t) - \mu_i| \leq \sqrt{\frac{T\delta^2}{HT_i(t)}}.
$$

As shown in Locatelli et al. (2016), one can use the reverse triangle inequality to get:

\begin{align*}
|\hat{\mu}_i(t) - \mu_i| & = |(\hat{\mu}_i(t) - \tau) - (\mu_i - \tau)| \\
& \geq || \hat{\mu}_i(t) - \tau | - |\mu_i - \tau|| \\
& \geq |(|\hat{\mu}_i(t) - \tau| + \epsilon) - (|\mu_i - \tau| + \epsilon)| \\
& \geq |\hat{\Delta}_i(t) - \Delta_i|
\end{align*}

If we combine this with the favorable event, we get

$$
\Delta_k - \sqrt{\frac{T\delta^2}{HT_k(t)}} \leq \hat{\Delta}_k(t) \leq \Delta_k + \sqrt{\frac{T\delta^2}{HT_k(t)}}
$$

and similar for $\hat{\Delta}_i$. We can now lower bound the LHS of the indexes as

\begin{align*}
B_k(t) & \geq \Big(\Delta_k - \sqrt{\frac{T\delta^2}{HT_k(t)}}\Big) \sqrt{T_k(t)} \\
& \geq \Big(\Delta_k - \sqrt{\frac{2}{h_k}}\delta \Big)
\end{align*}

At this point, it is useful to have a well chosen complexity measure. If we set $h_i = \frac{1}{\Delta_i^2}$ as in Locatelli et al. (2016), we get

$$
\Big(\Delta_k - \sqrt{2}\delta \Delta \Big) \sqrt{\frac{T}{2H\Delta_k^2}} \leq B_k(t)
$$

which can be now be reduced to a form that is not specific to $k$:

$$
\Big( \frac{1}{\sqrt{2}} - \delta \Big) \sqrt{\frac{T}{H}} \leq B_k(t).
$$

Now we need to remove $\hat{\Delta}_i$ in order to bound $T_i(t)$. Thus, we bound the RHS again using the favorable event much as we did for the LHS:

$$
B_i(t) = \hat{\Delta}_i \sqrt{T_i(t)} \leq \Big(\Delta_i + \sqrt{\frac{T\delta^2}{HT_i(t)}}\Big) \sqrt{T_i(t)} \leq \Delta_i \sqrt{T_i(t)} + \delta \sqrt{\frac{T}{H}}
$$

As in Locatelli et al. (2016) we use the fact that by definition $\Delta$ and $\hat{\Delta}$ are positive in order to square both sides and write

$$
(1 - 2\sqrt{2}\delta)^2 \frac{T}{2H\Delta^2_i} \leq T_i(t) \leq T_i(T)
$$

We thus have derived a lower bound on the amount of pulls for any arm. That means that we can show that no arm is pulled an insufficient amount of times with large probability, and that consequently the favorable event hold and the estimated mean is sufficienctly separated from the threshold.

### Check Separation Condition \label{sec:CheckSeparationCondition}

As the last step, we need to check whether the lower bound on the pulls of any arm suffices to ensure that the separation condition holds on the favorable event. That means, given $T_i(T) \geq (1-2\sqrt{2}\delta)^2 \frac{T}{2H\Delta_i^2}$, do we have $|\hat{\mu}_i(T) - \mu | \leq \frac{\Delta}{2}$ on the favorable event $\xi = \Big\{|\hat{\mu}_i - \mu | \leq \sqrt{\frac{T\delta^2}{H T_i(T)}}\Big\}$? To check, we plug the lower bound into the favorable event and get

$$
|\hat{\mu}_i - \mu | \leq \sqrt{\frac{T\delta^2}{H T_i(T)}} \leq \Delta_i \frac{\sqrt{2}\delta}{1-2\sqrt{2}\delta}
$$
Using $\delta = (4\sqrt{2})^-1$ as in Locatelli et al. (2016), we have that the separation condition holds exactly:

$$
|\hat{\mu}_i - \mu | \leq \Delta_i \frac{\sqrt{2}\delta}{1-2\sqrt{2}\delta} \leq \frac{1}{2}\Delta_i.
$$

### Upper Bound on Error Probability

We have seen that on the favorable event $\xi$, which holds with large probability $\mathbb{P}(\xi)$, the separation condition holds as all arms have been pulled a sufficient number of times, such that their estimated mean is close to the true mean. All arms are classified correctly, and the algorithm does not suffer a loss. On the other hand, we make an error whenever the favorable event does not hold after all; that is, the deviation bound for the mean does not hold. At least one arm is misclassified and the arm suffers a loss. This occurs with probability $1-\mathbb{P}(\xi)$. Consequently, the expected loss is upper bounded by

$$
\mathbb{E}[\mathcal{L}(T)] = 1 - \mathbb{P}(\xi) \leq 2(\log(T)+1)K \exp(-\frac{1}{64R^2}\frac{T}{H}).
$$

We now present the lower bound given in Locatelli et al. (2016) for the thresholding bandit problem. In order to derive the bound, a problem is specified to show that the error of the lower bound is attained even in this specific problem. In particular, for a set of problems with Gaussian distributions of variance $1$, the largest probability of error on these problems that the best algorithm exhibits is still at least as large as the lower bound.

## Lower Bound for the Thresholding Bandit Problem \label{sec:LowerBoundLocatelli}

Locatelli et al. (2016) also provide a general lower bound on the probability of error in the thresholding bandit problem. The bound is derived for a specified set of problems through a change in distribution as common in the multi-armed bandit literature (compare for example Audibert et al., 2010, or Kaufmann et al., 2016). For any of these problems, the largest probability of error that the best algorithm incurs is still lower bounded by the bound given below.

\begin{theorem}[Theorem 1, Locatelli et al., 2016] 
\label{theorem:Locatelli2016Theorem1}
For any sequence of gaps $(d_k)_k$, define a finite set of problems with Gaussian distributions with means corresponding to the gaps, and with variance 1. Lower bound the largest probability of error among these problems for the best possible algorithm as follows. Let $K,T \geq 0$. Let for any $i \leq K$, $d_i \geq 0$. Let $\tau \in \mathbb{R}, \epsilon > 0$.

For $0 \leq i \leq K$, we write $\mathcal{B}^i$ for the problem where the distribution of arm $j \in \{1, \dots, K\}$ is $\mathcal{N}(\tau+d_i+\epsilon, 1)$ if $i \neq j$ and $\mathcal{N}(\tau-d_i-\epsilon, 1)$ otherwise. For all these problems, $H := H_{\tau, \epsilon} = \sum_i (d_i+2\epsilon)^{-2}$ is the same by definition.

It holds that for any bandit algorithm,

\begin{equation*}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) \geq \exp(-3T/H-4 \log(12(\log(T)+1)K)),
\end{equation*}

where $\mathbb{E}_{\mathcal{B}^i}$ is the expectation according to the samples of problem $\mathcal{B}^i$.
\end{theorem}

A proof of the lower bound is given in the Appendix in \autoref{sec:AppendixAPTLB}.

If we write the lower bound as $LB = 12(\log(T)+1)K)\exp(-3T/H)\exp(-4) = C_{LB}(\log(T)+1)K\exp(-T/H)$ and the upper bound as
$UB = (\log(T)+1)K \exp(2) \exp(-\frac{1}{64R^2}\frac{T}{H}) = C_{UB}(\log(T)+1)K\exp(-T/H)$, we see that $LB$ and $UB$ match up to multiplicative constants and the $(\log(T)+1)K$ term which is negligiable in comparison with the exponent for large $T$ as shown in Locatelli et al. (2016). This makes the APT algorithm order optimal.

## Discussion \label{sec:DiscussionLocatelli}

One of the APT algorithm's main features is its simplicity. The sampling rule is straightforward, and, as highlighted in Locatelli et al. (2016), it does not require any knowledge about the type of sub-Gaussianity $R$, or the complexity $H$ upfront. This makes the algorithm immediately applicable in practice while it still enjoys its upper bound on the expected regret. This is in particular worth highlighting when comparing APT to algorithms used in the best arm identification or top-$m$ setting. As alluded to in \autoref{sec:TheAPTAlgorithm}, many algorithms that fix the sampling budget upfront (i.e., fixed budget algorithms as UCB-E or KL-LUCB-E) require knowledge about the complexity upfront (Audibert et al., 2010; Kaufmann and Kalyanakrishnan, 2013). The problem however being that to know the complexity, the experimenter has to know all means up to a permutation of the arms. This, of course, is not given in practice. Thus alternative algorithms try to estimate the complexity online which can be difficult as especially in the beginning plug-in estimates based on the estimated means may lead to wrong estimates of the complexity, which in turn leads to poor budget allocation based on mistaken beliefs (compare Garivier and Kaufmann, 2016). On the other hand, the sequential sampling strategy has to rely on some estimate. While the APT algorithm does not explicitly estimate the complexity, the gaps that serve as index are essentially the same. Empirically, we see in \autoref{chap:Experiments} that the problem described by Garivier and Kaufmann (2016) does also occur for the APT strategy. Faced not with sub-Gaussian but with more heavy-tailed distributions, this sensitivity is likely to increase.

Furthermore, as mentioned in Locatelli et al. (2016), the assumptions for the bounds no longer hold as arms become less sub-Gaussian; the complexity $H$ and the term $\hat{\Delta}_i(t)$ in the algorithm's index are both motivated by sub-Gaussian distributions. The more a problem deviates from those, the less appropriate the quantities become. In particular, the empirical mean will be less (or not at all) concentrated around the true mean and more impacted by outliers which are more likely under non sub-Gaussian distributions. Locatelli et al. (2016) thus refers to Bubeck et al. (2013a) which consider solutions for multi-armed bandits that are based on heavy-tail distributions. By using more robust estimators of the mean, one could equip the APT algorithm to deal with such distributions.

Both Mukherjee et al. (2017) as well as Zhong et al. (2017) use estimates of the variance of the arms for their algorithms for the thresholding bandit problem and try to thereby improve upon the performance of the APT algorithm. While their simulations show promising results, their bounds rely on the assumption of bounded reward distributions. Thus their algorithms are not necessarily able to improve upon the assumption of sub-Gaussian distributions made by Locatelli et al. (2016). We will continue to discuss their results in \autoref{sec:VarianceBasedAlgorithms}.

Yet another idea, again mentioned in Locatelli et al. (2016) is to show bounds based on the Kullback-Leibler divergence of the specific distribution directly. While the approximation through the Gaussian Kullback-Leibler divergence in the lower bound proof is convenient and leads to the simple APT algorithm, it might not stay appropriate for distributions whose Kullback-Leibler divergence tends to differ from the Gaussian. This, for example, is the case for Bernoulli distributions at very small parameters $\mu$. We have for the Kullback-Leibler divergence of two Gaussian distributions $\nu_i$ and $\nu_j$ of mean $\mu_i$ and $\mu_j$ with equal variance:

$$
\KL_{\mathcal{N}}(\nu_i || \nu_j) = 2(\mu_i - \mu_j)^2.
$$

Writing $\mu_j = \tau$ and $\epsilon = 0$, we see that the APT algorithm essentially measures the KL-divergence between two Normal distributions. For two Bernoulli distributions $\nu_i$ and $\nu_j$, parameterized by $\mu_i$ and $\mu_j=\tau$, we write

$$
\KL_{\text{Ber}}(\nu_i||\nu_j) = \mu_i \log(\frac{\mu_i}{\tau}) + (1-\mu_i) \log(\frac{1-\mu_i}{1-\tau})
$$

In the case of $\tau = 0.5$, the latter KL-divergence is symmetric in its first argument, and thus can be relatively well approximated by Pinsker's inequality:

$$
\KL_{\text{Ber}}(\nu_i||\nu_j) \geq 2(\mu_i - \tau)^2.
$$

This is equal to the Gaussian KL-divergence described before. The approximation, however, suffers heavily for very small values of $\mu_i$ and $\tau$. In particular, the Bernoulli KL-divergence becomes asymmetric in its first argument when $\tau \neq 0.5$. Thus, in contrast to the complexity measure $H$ based on the Gaussian KL-divergence, a complexity measure based on the Bernoulli KL-divergence would evaluate two arms as differently difficult if their absolute distance to the threshold is the same, but they lie on different sides of the threshold.

Thus similarly to how UCB algorithms in the cumulative regret setting have been extended to KL-UCB algorithms (Cappé et al., 2013), the APT algorithm offers room for improvement, too. As a more explicit example of such adaptation, consider the case of top-$m$ identification described in Kaufmann et al. (2016). The authors develop a lower bound based on KL-divergence, and note that "This information-theoretic lower bound permits to interpret the quantity $H(\nu)$ [...] as a subgaussian approximation". Given that Bernoulli distributions with very small or very large parameter values arise naturally in many applications, an improvement in this regard would be valuable.

In what follows, we first examine in \autoref{chap:KullbackLeiblerPureExploration} how information theoretic quantities are employed in the best arm identification and top-$m$ problem as detailed in Kaufmann and Garivier (2016) and Kaufmann et al. (2016). Before introducing a new algorithm for the thresholding bandit problem based on Kullback-Leibler divergences in \autoref{sec:LikelihoodRatioForExponentialFamily}, we first describe the variance-based approaches in \autoref{sec:VarianceBasedAlgorithms}.

```{r KLApproximationComparison, echo = FALSE, message = FALSE, out.width = '80%', fig.align = "center", fig.cap= "The approximation of the Bernoulli KL-divergence by the Gaussian KL-divergence works well if the second argument is equal to 0.5. Else we observe poor approximation in both value and symmetry. This becomes more obvious as the arguments move away from 0.5."}
library(dplyr)
library(tidyr)
library(ggplot2)
KL_ber <- function(p,q) p*log(p/q) + (1-p)*log((1-p)/(1-q))
KL_gauss <- function(p,q) 2*(p-q)^2

kldf <- data.frame(mu = rep(seq(0.00001, 0.99999, 
                                length.out = 10000), 2),
                   tau = c(rep(0.5, 10000), rep(0.05, 10000)))

kldf_long <- kldf %>% mutate(Gaussian = KL_gauss(mu, tau),
                             Bernoulli = KL_ber(mu, tau)) %>%
  gather(key = KL_Type, value = KL, -mu, -tau) %>%
  mutate(tau = as.character(tau))

kldf2 <- data.frame(mu = rep(seq(0.00001, 0.09999, 
                                length.out = 10000), 1),
                   tau = c(rep(0.05, 10000)))

kldf2_long <- kldf2 %>% mutate(Gaussian = KL_gauss(mu, tau),
                              Bernoulli = KL_ber(mu, tau)) %>%
  gather(key = KL_Type, value = KL, -mu, -tau) %>%
  mutate(tau = as.character(tau))

kldf_total <- cbind.data.frame(level = c(rep("Global", 40000),
                                         rep("Local", 20000)),
                               rbind.data.frame(kldf_long, kldf2_long))

ggplot(kldf_total, aes(x = mu, y = KL, 
                      group = interaction(KL_Type, tau))) +
  facet_grid(level~., scales = "free") +
  geom_line(aes(color = tau, linetype = KL_Type)) +
  labs(x = "First argument of KL-divergence",
       y = "KL-divergence"#,
       #title = "Comparison of Bernoulli KL-divergence and Gaussian Approximation"
       )  +
  theme_bw() +
  #theme(legend.position = "bottom") +
  scale_color_discrete(name = "Second argument \nof KL-divergence") +
  scale_linetype_discrete(name = "Type of \nKL-divergence") +
  guides(color = guide_legend(title.position = "top"),
         linetype = guide_legend(title.position = "top"))
```

\newpage

# Information Theoretic Approaches to Pure Exploration Bandits \label{chap:KullbackLeiblerPureExploration}

The previous chapter introduces the thresholding bandit problem as described in Locatelli et al. (2016), and the APT strategy as a first algorithm that is simple, yet order-optimal for large budgets. However, we end with the conclusion that a KL-divergence based measure of the distance between the distributions and the threshold might be more useful than the simple gap $\Delta^2_i = (\mu_i - \tau_i)^2$ employed so far. Indeed, citing Kaufmann et al. (2016), we note that the complexity $H$ -- and thus the gaps -- used so far can be considered a sub-Gaussian approximation of a better complexity measure based on information theoretic measures like the Kullback-Leibler divergence.

In order to give more intuition to why that is the case, and in which ways information theoretic measures can be introduced to measure the complexity of bandit problems, we now turn to Kaufmann et al. (2016) and Garivier and Kaufmann (2016). The latter describes a strategy based on Kullback-Leibler divergences for the best-arm identification problem, while the first derives complexity measures based on the divergence. Their work offers methods that can inform our goal of extending the thresholding bandit problem. In what follows, we adopt the notation of Kaufmann et al. (2016) to ease a comparison with their arguments.

## Information Theoretic Complexity Measures in Top-$m$ Problems \label{sec:KaufmannEtAl2016}

Instead of the thresholding bandit problem, Kaufmann et al. (2016) consider the problem of identifying the best $m$ out of $K$ arms, that is, the top-$m$ problem. They describe their bandit model $\nu = (\nu_1,...,\nu_K)$ by $K$ arms characterized by probability distributions $\nu_i$, $i \in \mathbb{A} = \{1,...,K\}$, with mean $\mu_i$. A learner picks at any time step $t=1,2,...,$ an arm $I_t \in \{1,...,K\}$ for which she observes the independent reward $X_{t} \sim \nu_{I_t}$. The learner aims to find the $m$ out of $K$ arms with the largest expected reward. This set of arm indexes is denoted by $\mathcal{S}_m^*$. Kaufmann et al. write the means $(\mu_{1}, ..., \mu_{K})$ sorted in decreasing order as the K-tuple $(\mu_{[1]}, ..., \mu_{[K]})$. Based on this, they require that $\mu_{[m]} > \mu_{[m+1]}$ such that the solution $\mathcal{S}_m^*$ is unique. The class of bandit problems $\nu$ for which the latter condition holds is called $\mathcal{M}_m$ by Kaufmann et al. At the end of the sampling procedure, after $T$ samples, the learner is required to return a classification $\hat{\mathcal{S}}_m$. This classification is then evaluated by the loss function $\mathcal{L}(T) = \mathbb{I}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m)$. This binary loss is similar to the loss in the thresholding bandit. We can now define a strategy of the learner as $\mathcal{S} = ((I_t), T, \hat{\mathcal{S}}_m)$: The strategy defines how an arm is chosen, how many samples are observed, and how the arms are eventually classified.

When introducing the thresholding bandit setting, we did not mention the fact that there exists an alternative setting in which the budget $T$ is not fixed upfront. This setting is called *fixed confidence*, while the case we considered so far is called *fixed budget*. In the fixed budget setting, we run algorithms for $T$ samples and then compare them based on their expected loss, $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m) =: p_t(\nu)$ as we have seen in $\autoref{sec:Setup}$. Alternatively, consider a $\delta$-PAC strategy $\mathcal{S}(\delta)$for which it holds for every problem $\nu \in \mathcal{M}_m$ that the classification holds at least with confidence level $1-\delta$, that is, $\mathbb{P}(\mathcal{S}^*_m = \hat{\mathcal{S}}_m) \geq 1-\delta$. The strategy has to achieve this level of confidence in its prediction before it is allowed to stop sampling. This is the fixed confidence case, in which the budget $T=T_{\delta}$ is no longer fixed upfront, but becomes a random variable. Algorithms are then compared by their expected number of draws $\mathbb{E}_{\nu}[T_{\delta}]$ and we prefer those which achieve the confidence level $1-\delta$ with a smaller expected budget.

In the fixed confidence setting, a family of strategies $S = (\mathcal{S}(\delta))_{\delta \in (0,1)}$ is called PAC if for every $\delta$, $\mathcal{S}(\delta)$ is $\delta$-PAC. On the other hand, in the fixed budget setting, a family of strategies $S = (\mathcal{S}(t))_{t \in \mathbb{N}}$ is called consistent if $p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$ for every problem $\nu \in \mathcal{M}_m$.

Given this setup, we can now follow the arguments of Kaufmann et al. (2016) to derive complexity measures for the top-$m$ problem. As alluded to earlier, these complexity measures will rely on the KL-divergence in contrast to the complexity $H$ used by Locatelli et al. (2016). When discussing $H$, its motivation is that it represents the difficulty of the problem by measuring the distance between arms and the threshold. Instead, Kaufmann et al. approach complexity by how many samples (what error probability) one will need in the fixed confidence (fixed budget) setting. In order to arrive at their new measure of complexity, it makes sense to first follow their arguments to a general lower bound on the sample complexity $\mathbb{E}_{\nu}[T]$ which is given in their Theorem 4. Here, $\mathcal{P}$ refers to a set of probability measures for which Kaufmann et al. (2016) assume certain continuity conditions in their Assumption 3, and $\nu_i \in \mathcal{P}$. In the following, $d(x,1-x) = x\log(\frac{x}{1-x}) + (1-x) \log(\frac{1-x}{x})$ is the binary KL-divergence.

\begin{theorem}[Theorem 4, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem4}
Let $\nu \in \mathcal{M}_m$, where $\mathcal{M}_m$ is the set of uniquely identifiable bandit models described above. Assume that $\mathcal{P}$ satisfies the continuity conditions in Assumption 3. Any $\delta$-PAC algorithm on $\mathcal{M}_m$ satisfies, for $\delta \leq 0.15$:

\begin{equation*}
\mathbb{E}_{\nu}[\tau] \geq \Big[ \sum_{a \in \mathcal{S}^*_m} \frac{1}{\KL(\nu_a || \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}^*_m} \frac{1}{\KL(\nu_a || \nu_{[m]})} \Big] \der(1-\delta, \delta),
\end{equation*}
\end{theorem}

To prove this statement for a given $\delta$-PAC strategy $\mathcal{S}$, we assume certain continuity conditions (compare Kaufmann et al. (2016)) under which there exists for every arm $i \in \{1, \dots, K\}$ an alternative model $\nu' = (\nu_1, \dots, \nu_{i-1}, \nu_i', \nu_{i+1}, \dots, \nu_K)$. In this alternative model, in accordance with the continuity conditions, only arm $i$ is modified. After the modification, the arm $\nu_{m+1}$ is now closer to $\nu_i$ as measured by the Kullback-Leibler divergence than $\nu_i'$ is to $\nu_i$, if $\mu_i > \mu_{m+1}$ (consequently, $\mu_i' < \mu_{m+1}$). On the other hand, if $\mu_a < \mu_{m}$, then $\nu_a'$ will be moved away from $\nu_i$ just enough so that $\nu_i$ is closer to $\nu_m$ than to $\nu_i'$ (consequently, $\mu_i' > \mu_m$). These changes in distribution make the original set of optimal arms, $\{1,\dots,m\}$, no longer optimal under the alternative model $\nu'$. To compare this to the thresholding bandit problem, assume that arms $\nu_{m}$ and $\nu_{m+1}$ indirectly define a threshold $\tau \in (\mu_{m+1},\mu_m)$. Then, by creating an alternative problem as described here, we flip arm $i$ around the implied threshold $\tau$ (compare the lower bound proof in \autoref{sec:AppendixAPTLB}).

Given the strategy returns $\hat{S}_m$, we know that for the event $\mathcal{E} = (\hat{S}_m = \{1, \dots, m\})$ (which is element of the $\sigma$-Algebra $\mathcal{F}_{T}$ induced by the stopping time $T$), that $\mathbb{P}_{\nu}(\mathcal{E}) \geq 1-\delta$ and $\mathbb{P}_{\nu'}(\mathcal{E}) \leq \delta$ (because fixed confidence $\delta$!). To give the lower bound on the expected number of samples, $\mathbb{E}_{\nu}[\tau]$, we employ the general Lemma given by the authors for such changes in distribution in bandit models.

\begin{lemma}[Lemma 1, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlLemma1}
Let $\nu$ and $\nu'$ be two bandit models with $K$ arms such that for all $i$, the distributions $\nu_i$ and $\nu_i'$ are mutually absolutely continuous. For any almost-surely finite stopping time $\sigma$ with respect to $(\mathcal{F}_t)$,

\begin{equation*}
\sum_{i=1}^{K} \mathbb{E}_{\nu} [N_i(\sigma)] \KL(\nu_i|| \nu_i') \geq \sup_{\mathcal{E} \in \mathcal{F}_{\sigma}} d(\mathbb{P}_{\nu}(\mathcal{E}), \mathbb{P}_{\nu'}(\mathcal{E})).
\end{equation*}
\end{lemma}

We apply this Lemma directly on the expected number of draws for each arm, $\mathbb{E}_{\nu}[N_i]$. For each arm $a$, we thus have $\KL(\nu_i, \nu_i') \mathbb{E}_{\nu}[N_i] \geq d(1-\delta, \delta)$. Using the definition of the alternative model above, we get for every small but fixed $\alpha > 0$, every arm $i \in \{1, \dots, K\}$ and $j \in \{m+1, \dots, K\}$:

\begin{equation*}
\mathbb{E}_{\nu}[T_i] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_i || \nu_{m+1}) + \alpha} \quad \text{and} \quad \mathbb{E}_{\nu}[T_j] \geq \frac{\der(1-\delta, \delta)}{\KL (\nu_j || \nu_{m}) + \alpha}.
\end{equation*}

Let $\alpha \to 0$, sum the individual arms' bounds to get the bound on the overall number of draws:

\begin{equation*}
\mathbb{E}_{\nu}[T] = \sum_{i=1}^K \mathbb{E}_{\nu}[T_i] \geq \Big( \sum_{i \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_i|| \nu_{[m+1]})} + \sum_{i \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_i|| \nu_{[m]})} \Big) d(1-\delta, \delta)
\end{equation*}

Given that we will be naturally interested in either very high or very low confidence of our algorithm, that is, $P_\nu(\mathcal{E})$ is either close to 0 or 1, it is possible to use the following approximation to the binary relative entropy given in Kaufmann et al. (2016):

\begin{equation}\label{bre_approx}
\forall x \in [0,1], \quad d(x,1-x) = d(1-x,x) \geq \log \frac{1}{2.4x}.
\end{equation}

This allows us to write the lower bound on the expected stopping time as:

\begin{equation*}
\mathbb{E}_{\nu}[T] \geq \Big( \sum_{i \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_i|| \nu_{[m+1]})} + \sum_{i \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_i|| \nu_{[m]})} \Big) \log \big(\frac{1}{2.4\delta} \big).
\end{equation*}

Given this lower bound, it is clear that the complexity of the top-$m$ problem can be lower bounded by the KL-divergence between arms and the implied threshold between the $m$-th best arm and the $(m+1)$-th best arm. The last equation gives a lower bound on the effort (in terms of expected number of samples) required to achieve a certain confidence level, where a problem that requires a larger effort is described as more complex. Indeed, in the case of fixed confidence problems, Kaufmann et al. (2016) describe problems as having a larger *sampling complexity*. Following Kaufmann et al. (2016), we define the complexity $\kappa_C(\nu)$ in the fixed confidence setting as

\begin{equation}
\kappa_C(\nu) = \inf_{S \, \text{PAC}} \limsup_{\delta \to 0} \frac{\mathbb{E}_{\nu}[T_{\delta}]}{\log \frac{1}{\delta}}. \label{eq:KappaC}
\end{equation}

This definition makes sense if we consider the fact that the confidence level $\delta$ is experimenter chosen and thus not relevant in describing the complexity. Thus, by including $\log(1/\delta)$ explicitly in \eqref{eq:KappaC}, we get rid of the confidence term in the lower bound on $\mathbb{E}_{\nu}[T]$ after plugging it into \eqref{eq:KappaC}. We arrive at the lower bound on the complexity measure given in Kaufmann et al. (2016):

\begin{equation*}
\kappa_C(\nu) \geq \sum_{a \in \mathcal{S}_m^*} \frac{1}{\KL(\nu_a || \nu_{[m+1]})} + \sum_{a \notin \mathcal{S}_m^*} \frac{1}{\KL(\nu_a || \nu_{[m]})}
\end{equation*}

Given this description, we would expect that a complexity measure in a fixed confidence thresholding bandit probem might involve replacing $\mu_{[m]}$ and $\mu_{[m+1]}$ by the actual threshold $\tau$ that divides the group above the threshold from the group below the threshold. But given the setting in Locatelli et al. (2016), we are mainly interested in complexity measures and algorithms for the fixed budget problem. The results of Kaufmann et al. (2016) are more limited for this case. While the authors are able to provide a bound on their complexity measure in the fixed budget case, it only applies to the case of *two-armed* bandit models.

Consider a consistent fixed budget strategy $\mathcal{S}$ ($p_t(\nu)$ tends to zero as $t \rightarrow \infty$), and define the complexity in the fixed budget setting as follows:

\begin{equation}
\kappa_B(\nu) = \inf_{A \, \text{consistent}} \Big(\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu)\Big)^{-1}.
\end{equation}

While we included the experimenter-fixed $\delta$ in the definition of $\kappa_C$, we now include the fixed $T$ in the definition of $\kappa_B$ to disregard it in the complexity measure after plugging in the lower bound on $p_t(\nu)$. The resulting lower bound given in Theorem 12 of Kaufmann et al. (2016). We see that the best exponent with which the error $p_t(\nu)$ decreases as $t \rightarrow \infty$ is based on Kullback-Leibler divergences:

\begin{theorem}[Theorem 12, Kaufmann et al., 2016] \label{theorem:KaufmannEtAlTheorem12}
Let $\nu = (\nu_1, \nu_2)$ be a two-armed bandit model such that $\mu_1 > \mu_2$. In the fixed budget setting, any consistent algorithm satisfies:

\begin{equation*}
\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu) \leq \inf_{(\nu_1', \nu_2') \in \mathcal{M}: \mu_1' < \mu_2'} \max \big\{\KL(\nu_1' || \nu_1), \KL(\nu_2' || \nu_2)\big\}
\end{equation*}
\end{theorem}

For a proof, consider first the two different bandit models, $\nu$ and $\nu'$, where we fix w.l.o.g. $i^* = 1$ as the best arm in $\nu$, while $i^* = 2$ in $\nu'$. We require the algorithm $\mathcal{S}$ to be consistent for both problems. Since we are in the fixed budget setting, the stopping time is fixed at $T = t$. The correct decision differs between the two problems, and we focus on the event $A = (\hat{S}_1 = 1)$. Now, we can apply Lemma 1 from Kaufmann et al. (2016) as before, as it gives a measure of how different two bandit models are (it bounds the change in distribution). Here, we want to know how different $\nu$ is from $\nu'$, and, given we have only two arms,

\begin{equation*}
\mathbb{E}_{\nu'}[T_1(t)]\KL(\nu_1'|| \nu_1) + \mathbb{E}_{\nu'}[T_2(t)]\KL(\nu_2'|| \nu_2) \geq \der(\mathbb{P}_{\nu'}(A),\mathbb{P}_{\nu}(A)))
\end{equation*}

The expectations were chosen with respect to $\nu'$ in order to allow for the following: We have $p_t(\nu) = 1 - \mathbb{P}_{\nu}(A)$ and $p_t(\nu') = \mathbb{P}_{\nu'}(A)$ (since $A$ is wrong for $\nu'$). But since algorithm $\mathcal{A}$ is consistent for both $\nu$ and $\nu'$, we know that there should exist some number of samples such that the probability of event $A$ is bigger in setting $\nu$ (where it is the correct decision) than in setting $\nu'$ (where this would be a wrong decision and failure). Thus, for every $\epsilon > 0$, there exists $t_0(\epsilon)$ such that for all $t \geq t_0(\epsilon)$, $\mathbb{P}_{\nu'}(A) \leq \epsilon \leq \mathbb{P}_\nu(A)$. Consequently, if we plug in the bound of $\epsilon$ for $\mathbb{P}_{\nu'}(A)$, then

\begin{align*}
\mathbb{E}_{\nu'}[T_1(t)]\KL(\nu_1' || \nu_1) + \mathbb{E}_{\nu'}[T_2(t)]\KL(\nu_2' || \nu_2) & \geq \der(\epsilon,1-p_t(\nu))) \\
& = \epsilon \log \frac{\epsilon}{1-p_t(\nu)} + (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} \\
& \geq (1-\epsilon) \log \frac{1-\epsilon}{p_t(\nu)} + \epsilon \log \epsilon.
\end{align*}

Let $\epsilon$ go to zero so that

\begin{align*}
\mathbb{E}_{\nu'}[T_1(t)]\KL(\nu_1' || \nu_1) + \mathbb{E}_{\nu'}[T_2(t)]\KL(\nu_2' || \nu_2)
& \geq \log \frac{1}{p_t(\nu)} = -\log p_t(\nu).
\end{align*}

Dividing both sides by $t$ and taking the limsup, we arrive at

\begin{align*}
\limsup_{t \to \infty} - \frac{1}{t} \log p_t(\nu)
& \leq \limsup_{t \to 0} \frac{1}{t} \sum_{i=1}^2 \mathbb{E}_{\nu'}[T_i(t)] \KL(\nu_i' || \nu_i) \\
& \leq \max_{i=1,2} \KL(\nu_i' || \nu_i),
\end{align*}

where the last line follows from the fact that the $\frac{1}{t} \sum_{i=1}^2 \mathbb{E}_{\nu'}[T_i(t)]$ part creates a weighted average of the two divergences $\KL(\nu_i', \nu_i)$. As noted in Kaufmann et al. (2016), one arrives at the final result by choosing $\mu_1'$, $\mu_2'$ so that $| \KL(\nu_1', \nu_1) - \KL(\nu_2', \nu_2) |$ is as small as possible.

As noted by Kaufmann et al. (2016), the advantage of this lower bound is that it makes the complexity in the fixed confidence and the fixed budget setting directly comparable. The disadvantage being that the result only holds for two-armed models. While the authors are not able to derive directly comparable lower bounds for the top-$m$ case, they derive a first lower bound in the Gaussian case with equal variances.

Before we discuss algorithms given in Kaufmann et al. (2016) that match the lower bounds, we show how based on the proof of Theorem 12 in Kaufmann et al. (2016) very similar lower bounds can be derived for the fixed budget thresholding problem.

### Lower Bound for Fixed Budget Thresholding Problem \label{sec:LBTBP}

In order to derive a bound for the fixed budget problem, Kaufmann et al. (2016) focus on the setting with two arms which is the most simple version of the top-$m$ problem. Less than two arms does not make sense, since the single arm is by definition the best arm. In the case of the thresholding bandit, however, a single-armed problem makes still at least limited sense in that we still have to classify this arm correctly with respect to the threshold. Thus, we can show the following corollary which gives a bound on the complexity of the single-armed thresholding bandit.

\begin{corollary}
Consider the setting in \autoref{sec:Setup}. Let $\nu = \nu_1$ be a one-armed thresholding bandit model where $\tau$ is the threshold fixed upfront. Let $\nu$ be from the one-parameter exponential family of distributions such that $\nu_1$ can be parameterized by its mean, $\mu_1$, and the Kullback-Leibler divergence between distributions $\nu_1$ and $\nu_1'$ can be written as $\KL(\nu_1', \nu_1) = \KL(\mu_1', \mu_1)$. In the fixed budget setting, any consistent algorithm ($p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$) satisfies for $\mu_1' \rightarrow \tau$:

$$
- \frac{1}{t} \log p_t(\nu) \leq \KL(\mu_1' || \mu_1)
$$
\end{corollary}\label{corollary:LB1TBP}

This is basically the Chernoff-Stein Lemma for the hypothesis test of two alternatives (compare chapter 11 in Cover and Thomas, 2006, or \autoref{theorem:ChernoffStein}), in which the optimal error exponent is based on the Kullback-Leibler divergence. We arrive at this result as the one-armed bandit does not require sequential sampling: The single arm receives all samples. $\mathbb{E}_{\nu'}[N_1(t)]$ is not a random variable but equals $T$. The single-armed bandit is a simple hypothesis test.

But as we see in the following corollary, we can extend the lower bound to the $K$-armed fixed budget thresholding bandit problem, which stands in contrast to the fact that for the top-$m$ problem, it is only possible to show the lower bound for two arms. The result is based on an alternative setting $\nu'$ in which all arms have means $\mu' = \tau$ such that their expected number of pulls corresponds to uniform sampling, $\mathbb{E}_{\nu'}[N_i(t)] = T/K$.

\begin{corollary}
Consider the setting in \autoref{sec:Setup}. Let $\nu = (\nu_1, ..., \nu_K)$ be a $K$-armed thresholding bandit problem in which $\tau$ is the threshold fixed upfront. For $i \in \{1,...,K\}$, let $\nu_i$ be from the one-parameter exponential family of distributions such that $\nu_i$ can be parameterized by its mean, $\mu_i$, and the Kullback-Leibler divergence between distributions $\nu_i$ and $\nu_i'$ can be written as $\KL(\nu_i, \nu_i') = \KL(\mu_i, \mu_i')$. In the fixed budget setting, any consistent algorithm satisfies for $\mu_i' \rightarrow \tau$:

$$
\limsup_{t \rightarrow \infty} -\frac{1}{t} \log p_t(\nu) \leq \max_{i \in \{1,...,K\}} \KL(\mu_i' || \mu_i)
$$

or

$$
- \frac{1}{t} \log p_t(\nu) \leq \frac{1}{K}\sum_{i=1}^{K} \KL(\tau || \mu_i),
$$
\end{corollary}\label{corollary:LBKTBP}

We present the proofs of \autoref{sec:LBTBP} in the appendix in \autoref{sec:corollaryLBTBPproof}. Corresponding to the complexity measures defined in equation $(1)$ of Kaufmann et al. (2016), we can define the complexity $\kappa_{TBP}(\nu)$ in the thresholding bandit problem and lower bound it:

$$
\kappa_{TBP}(\nu) = \inf_{S \text{ consistent}} \Big(\limsup_{t \rightarrow \infty} - \frac{1}{t} \log p_t(\nu)\Big)^{-1} \geq K \Big(\sum_{i=1}^{K} \KL(\mu_i'||\mu_i)\Big)^{-1}
$$

### Strategies Matching the Top-$m$ Lower Bound

Kaufmann et al. (2016) do not present algorithms matching the lower bound of Theorem \ref{theorem:KaufmannEtAlTheorem17}. They do consider algorithms that connect to Theorem \ref{theorem:KaufmannEtAlTheorem4} and Theorem \ref{theorem:KaufmannEtAlTheorem12}. Given that we aim to extend the fixed budget thresholding bandit problem, we are most interested in the matching strategies for Theorem \ref{theorem:KaufmannEtAlTheorem12}. However, given that this lower bound applies only to two-armed bandits, the algorithms proposed by Kaufmann et al. (2016) are of limited practical interest. Furthermore, the proposed strategies require knowledge that is in general not available upfront: For Gaussian distributions, the knowledge of the (potentially different) variance is assumed which enables a strategy that pulls each arm proportional to its variance.

Even more difficult are the assumptions in the case of Bernoulli two-armed bandit models: The proposed algorithm requires knowledge about the true parameters of the distributions. That this limits the practical interest is also noted by the authors of Kaufmann et al. (2016). Thus, they suggest to use uniform sampling strategies in the case of two-armed Bernoulli bandits, and show that there is not much to be lost by doing so.

Thus, Kaufmann et al. (2016) joins other papers like Audibert et al. (2010) and Kaufmann and Kalyanakrishnan (2013) which struggle to find matching strategies in the fixed budget setting that do not rely on knowledge of parameters or the complexity of the problem.

In contrast, Kaufmann et al. (2016) are able to show that their complexity measure derived for the fixed confidence setting, $\kappa_C(\nu)$, can be upper-bounded using the bound given for the KL-LUCB algorithm in Kaufmann and Kalyanakrishnan (2013). In the case of exponential family bandits, the complexity $\kappa_C(\nu)$ is bounded as given in equation (6) in Kaufmann et al. (2016):

\begin{equation*}
\sum_{i \in S^*_m} \frac{1}{\KL(\mu_i || \mu_{[m+1]})} + \sum_{i \notin S^*_m} \frac{1}{\KL(\mu_i || \mu_{[m]})} \leq \kappa_C(\nu) \leq 24 \min_{\mu \in [\mu_{[m+1]},\mu_{[m]}]} \sum_{i=1}^{K} \frac{1}{\KL^*(\mu_i || \mu)},
\end{equation*}

where again we write with slight abuse of notation $\KL(\nu_{\mu}|| \nu_{\mu'}) = \KL(\mu || \mu')$ as the shorthand for the KL-divergence of two distributions from the exponential family parameterized by $\mu$ and $\mu'$ respectively. $\KL^*(\mu || \mu') = \KL(\mu^* || \mu)$ is the Chernoff information between the two distributions, where $\mu^*$ is chosen so that $\KL(\mu^* || \mu) = \KL(\mu^* || \mu')$.

## Information Theoretic Approach to Best Arm Identification \label{sec:OptimalBestArmIdentification}

Instead of discussing the KL-LUCB algorithm for the fixed confidence top-$m$ problem, we shift our focus onto the fixed confidence best arm identification problem, as discussed in Garivier and Kaufmann (2016). The authors also derive a lower bound on the sample complexity using Lemma \autoref{theorem:KaufmannEtAlLemma1}. Furthermore, they propose a matching strategy for $K \geq 2$ arms.

The setting differs slightly from the setting considered in Kaufmann et al. (2016). In particular, Garivier and Kaufmann (2016) focus on distributions $\nu_{\theta}$ from the one-parameter canonical exponential family of distributions, which can be parameterized by their mean $\mu$, and thus can be written as $\nu_{\mu}$; as before, we simplify the notation of the KL-divergence for these distributions to $\KL(\nu || \nu') = \KL(\mu || \mu')$.

Furthermore, if we let $\mathcal{S}$ to be a set of bandit models with $K$ distributions each, then each model is fully defined by a vector $\bm{\mu} = (\mu_1, \dots, \mu_K)$. Assume that every $\bm{\mu} \in \mathcal{S}$ has a unique maximum given by $i^*(\bm{\mu})$ such that $\mu_{i^*} > \max_{i \neq i^*(\bm{\mu})} \mu_i$. In order to fix the confidence $1-\delta$, we specify the risk $\delta$. Similar to the top-$m$ case considered before, strategies corresponding to this risk are called $\delta$-PAC if for every $\bm{\mu} \in \mathcal{S}$, the strategy ends surely after a finite number of samples, $\mathbb{P}_{\bm{\mu}}(T_{\delta} < \infty) = 1$, and the error probability is at least as small as the fixed risk level, $\mathbb{P}_{\bm{\mu}}(\hat{i}_{T_{\delta}} \neq i^*) \leq \delta$. In order to derive the lower bound on the expected number of samples $\mathbb{E}[T]$ through a change in distributions, we introduce the set of alternative problems for which the optimal arm is different from $\bm{\mu}$:

\begin{equation*}
\text{Alt}(\bm{\mu}) := \{\bm{\lambda} \in \mathcal{S}: i^*(\bm{\lambda}) \neq i^*(\bm{\mu})\},
\end{equation*}

where $\bm{\lambda}$ is the vector of means defining the distributions of the alternative problem. Lastly, we have the set of probability distributions on the arms $\{1,...,K\}$: $\Sigma_K = \{w \in \mathbb{R}_{+}^K: w_1 + \dots + w_K = 1\}$. Each $w_i$ corresponds to the fraction of overall samples that arm $i$ receives (which are random variables similar to $T_i$ in \autoref{sec:UpperBoundProof}). We now state the lower bound:

\begin{theorem}[Theorem 1, Garivier and Kaufmann, 2016] \label{theorem:GarivierKaufmannTheorem1}
Let $\delta \in (0,1)$. For any $\delta$-PAC strategy and any bandit model $\bm{\mu} \in \mathcal{S}$,

\begin{equation*}
\mathbb{E}_{\bm{\mu}} [T_{\delta}] \geq T^*(\bm{\mu}) d(\delta,1-\delta),
\end{equation*}

where

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{i=1}^K w_i \text{\em{KL}}(\mu_i|| \lambda_i)\right).
\end{equation}
\end{theorem}

We see that the two bounds given in Theorem \ref{theorem:KaufmannEtAlTheorem4} and Theorem \ref{theorem:GarivierKaufmannTheorem1} on the sample complexity differ only in the way in which they sum the Kullback-Leibler divergences over the $K$ arms. In contrast to Theorem \ref{theorem:KaufmannEtAlTheorem4}, here the contribution of each arm through its KL-divergence is weighted. The difference is that while we were able to clearly describe the change of distribution in the top-$m$ case for each arm -- we specified an alternative model for each arm that yields a new set of $m$ optimal arms -- we now consider all alternative models that yield a different best arm than the model $\bm{\mu}$. In the previous theorem, we sum over the complexity of each individual arm, since the change in distribution only affects one arm at a time, thus leading to a KL-divergence of 0 for the other arms while still changing the optimal set. Now, we consider all instances of alternative models $\bm{\lambda}$, and thus also those in which several arms are modified (and thus have a KL-divergence unequal 0 in the result of Lemma 1).

To proof Theorem \ref{theorem:GarivierKaufmannTheorem1}, we let $\delta \in (0,1)$, $\bm{\mu} \in \mathcal{S}$, and only consider $\delta$-PAC strategies. As before, $T_i(t)$ describes the random number of pulls of arm $i$ until time $t$, for all $t\geq 1$. Again, Lemma \ref{theorem:KaufmannEtAlLemma1} states how the risk $\delta$ relates to the expected number of draws of each arm given two models which differ in their optimal arm. How much the two models differ is evaluated based on the KL-divergence for each pair of arms:

\begin{equation*}
\forall \lambda \in \mathcal{S}: i^*(\lambda) \neq i^*(\mu), \sum_{i=1}^K \KL(\mu_i,\lambda_i)\mathbb{E}_{\bm{\mu}}[T_i(T_{\delta})] \geq d(\delta, 1- \delta).
\end{equation*}

As mentioned, we now do not consider a specific alternative for each arm, but instead look at the infimum over all the possible alternatives, and use the fact that $\mathbb{E}_{\bm{\mu}}[T_{\delta}] = \sum_{i=1}^K \mathbb{E}_{\bm{\mu}}[T_i(T_{\delta})]$:

\begin{align*}
\der(\delta, 1-\delta) & \leq \inf_{\lambda \in \text{Alt}(\bm{\mu})} \mathbb{E}_{\bm{\mu}}[T_{\delta}] \Big(\sum_{i=1}^K \frac{\mathbb{E}_{\bm{\mu}}[T_i]}{\mathbb{E}_{\mu}[T_{\delta}]} \KL(\mu_i, \lambda_i)\Big) \\
& = \mathbb{E}_{\bm{\mu}}[T_{\delta}] \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{i=1}^K \frac{\mathbb{E}_{\bm{\mu}}[T_i]}{\mathbb{E}_{\mu}[T_{\delta}]} \KL(\mu_i, \lambda_i)\Big) \\
& \leq \mathbb{E}_{\bm{\mu}}[T_{\delta}] \sup_{w \in \Sigma_K} \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{i=1}^K w_i \KL(\mu_i, \lambda_i)\Big).
\end{align*}

As stated in Garivier and Kaufmann (2016), the term $\frac{\mathbb{E}_{\bm{\mu}}[T_i]}{\mathbb{E}_{\mu}[T_{\delta}]}$ gives the fraction of draws each arm receives. This value, however, depends on the used algorithm. To make it hold for any $\delta$-PAC algorithm, it is replaced by the supremum over $\Sigma_K$, which is the distribution over these fractions for any $\delta$-PAC strategy.

This lower bound can be interpreted as a minimax game^[See the presentation by Aurélien Garivier at COLT 2016: <https://youtu.be/eLyrrYqJlkE>], where the experimenter has to choose the share of samples she attributes to each arm in a way that is optimal given that nature will respond and choose the most difficult alternative model given her weights. Thus, to make the lower bound practical, it is a matter of finding the optimal weights $w^*(\bm{\mu})$. This step did not occur in the lower bound for the top-$m$ problem since the alternative models were specified by changing the distribution of one arm with respect to the "border" between the two groups of arms.

The existence of the optimal weights suggests an algorithm that keeps track of the share of samples that is attributed to each arm and makes sure that it stays close to the optimum until some decision rule assures that the $\delta$-PAC condition is fulfilled.

Indeed, the optimal weights will lead to a sampling structure in which arms are drawn such that their individual probabilities of error stay at the same level. An arm more distant from the optimal arm in terms of Kullback-Leibler divergence should receive a smaller share of the samples to achieve the same error probability as an arm closer to the optimal arm.

This is what happens when we look for proportions $w \in \Sigma_K$ that minimize the cost from a change in distributions, that is, the term $F(\bm{\mu}, \bm{w}) = \inf_{\bm{\lambda} \in \text{Alt}(\bm{\mu})} \Big(\sum_{i= 1}^K w_i \KL(\mu_i, \lambda_i)\Big)$. In the proof of Lemma 3 of Garivier and Kaufmann, it is shown that nature, selecting the alternative model $\bm{\lambda}$, will always pick an alternative model in which only a single arm $i$ is changed such that $\lambda_i > \lambda_1$ since

\begin{align*}
T^*(\bm{\mu})^{-1} & = \sup_{w \in \Sigma_K} \min_{i \neq 1} \inf_{\bm{\lambda} \in \mathcal{S}: \lambda_i > \lambda_1} \sum_{i=1}^K w_i \KL(\mu_i, \lambda_i)  \\
& = \sup_{w \in \Sigma_K} \min_{i \neq 1} \inf_{\bm{\lambda}: \lambda_i \geq \lambda_1} [w_1 \KL(\mu_1, \lambda_1)+w_i \KL(\mu_i, \lambda_i)]. \Alignyesnumber\label{eqn:ComplexityTrackStopStrategy}
\end{align*}

Consequently, $w$ will be chosen so as to minimize $T^*(\bm{\mu})$ no matter which arm $i$ is picked. It is a minimax game, so the ideal solution is to keep for each arm $i \neq 1$ the term $w_1 \KL(\mu_1, \lambda_1) + w_i \KL(\mu_i, \lambda_i)$ at the same level. Garivier and Kaufmann show that this can be written in terms of the Jenson-Shannon divergence, defined for every $\alpha \in [0,1]$ as

\begin{equation*}
I_{\alpha}(\mu_1, \mu_2) := \alpha \KL(\mu_1 || \alpha \mu_1 + (1-\alpha) \mu_2) + (1-\alpha) \KL(\mu_2 || \alpha \mu_1 + (1-\alpha) \mu_2).
\end{equation*}

Using this notation, the optimal proportions can be written as 
\begin{equation*}
w^*(\bm{\mu}) = \arg \max_{w \in \Sigma_K} \min_{i \neq 1} (w_1 + w_i) I_{\frac{w_1}{w_1+w_i}}(\mu_1, \mu_i)
\end{equation*}

which let's us furthermore conclude that these quantities are equal for all $i,j \in \{2,\dots, K\}$ with
\begin{equation*}
(w_1^* + w_i^*) I_{\frac{w_1^*}{w_1^*+w_i^*}}(\mu_1, \mu_i) = (w_1^* + w_j^*) I_{\frac{w_1^*}{w_1^*+w_j^*}}(\mu_1, \mu_j),
\end{equation*}

which is simply the statement in Lemma 4 of Garivier and Kaufmann (2016).

And so by assuring the reader that there are ways to compute $w_i^*$ and $T^*$ given $\bm{\mu}$ (see chapter 2.2 in Garivier and Kaufmann (2016)), we can now introduce the *Track-and-Stop* strategy given in Garivier and Kaufmann (2016) that is able to match the lower bound. The main idea of the *Track-and-Stop* algorithm is to keep track of the optimal shares for each arm by equalizing the error probabilies for each arm.

When going from theory to algorithm, the obvious problem is that we no longer know the true parameters $\bm{\mu}$ and cannot calculate the optimal weights. Instead, we have to estimate the maximum likelihood estimates $\hat{\mu}(t) = (\hat{\mu}_1(t), \dots, \hat{\mu}_K(t))$ of the true mean at the current round $t$ and need to base our sample weights on them. It is however problematic to rely on the sample means as we are inherently unsure about them. And so wrong estimates, which naturally occur most easily in the beginning when not many samples have been pulled, are prone to lead to wrong weights. Thus, if a wrong estimate $\hat{\mu}_i$ for arm $i$ leads to a small weight $w_i$, then after a few more rounds neither the estimate nor the weight will be correct. Thus, Kaufmann and Garivier conclude that "a naive plug-in sampling rule sometimes fails". They propose to enforce a certain minimum level of exploration to ensure that arms that erroneously received a very small weight will be further explored regardless.

A very intuitive solution to this problem is proposed in Lemma 8 of Garivier and Kaufmann (2016) with the $D-Tracking$ rule. Whenever some arm has received a very small number of pulls $T_i(t)$ given the number of iterations $t$ so far, and the number of arms $K$ that have to share the samples, pull the arm that has received the smallest amount of samples so far. Else, pull the arm whose number of pulls so far $T_i(t)$ is the farthest away from what its optimal share $t\cdot w^*_i(\hat{\bm{\mu}}(t))$ would be. Given this sampling rule, the authors can show in their Proposition 9 that the empirical proportion of draws converges in probability to the optimal weights given the true means.

When combining this sampling rule with an appropriate stopping rule, the authors are able to prove in Theorem 14 an upper bound on the expected sampling complexity $\mathbb{E}_{\mu}[T_{\delta}]$. The stopping rule ensures that the algorithm explores sufficiently to promise a risk smaller than $\delta$, while stopping as soon as possible. Given that the sampling rule is independent of this stopping rule, the rule boils down to a statistical test at every iteration and thus is a sequential test based on the current estimate of $\hat{\bm{\mu}}$. Chernoff's Stopping Rule proposed in Garivier and Kaufmann (2016) calculates for each arm $i\in \mathbb{A} = \{1,...K\}$ the Generalized Likelihood Ratio statistic with every arm $j \in \mathbb{A}^{\setminus i}$, which for $\hat{\mu}_i(t) \geq \hat{\mu}_j(t)$ can be written as

\begin{equation*}
Z_{i,j}(t) = - Z_{j,i} = T_i(t) \KL(\hat{\mu}_i(t) || \hat{\mu}_{i,j}(t)) + T_j(t) \KL(\hat{\mu}_{j}(t) || \hat{\mu}_{i,j}(t)),
\end{equation*}

where

\begin{equation*}
\hat{\mu}_{i,j}(t) := \frac{T_i(t)}{T_i(t) + T_j(t)}\hat{\mu}_i(t) + \frac{T_j(t)}{T_i(t) + T_j(t)}\hat{\mu}_j(t).
\end{equation*}

Consequently, a sensible idea for a stopping rule would be that there has to be an arm $i \in \{1,...,K\}$ for which $Z_{i,j} \geq 0 \text{ for all }  j \in \{1,...,K\}^{\setminus i}$ (which is always fulfilled if there is a uniquely best arm), and we demand that for this arm the smallest of the $(K-1)$ test statistics has to be larger than an exploration threshold $\beta(t, \delta)$. The final sampling round $T_{\delta}$ of the algorithm is given by:

\begin{equation*}
T_{\delta} = \inf \{t \in \mathbb{T}: \exists i \in \mathbb{A}, \forall j\in \mathbb{A}^{\setminus i}, Z_{i,j}(t) > \beta(t, \delta)\}
\end{equation*}

Demanding that the optimal exploration threshold increases in rounds $t$ (if we have more samples, require more certainty; punish for failing previous tests) and decreases in risk $\delta$ at a diminishing rate, the optimal exploration rate for Bernoulli arms derived in Theorem 10 of Garivier and Kaufmann (2016) is intuitive:

\begin{equation*}
\beta(t, \delta) = \log \Big(\frac{2t(K-1)}{\delta}\Big).
\end{equation*}

With this, we can state the final theorem of the chapter, which gives the upper bound matching the lower bound in Theorem \ref{theorem:GarivierKaufmannTheorem1} for small risk values $\delta$ using the D-Tracking rule and Chernoff's stopping rule with the exploration term given in the previous equation:

\begin{theorem}[Theorem 14, Garivier and Kaufmann, 2016] \label{theorem:GarivierAndKaufmannTheorem14}
Let $\bm{\mu}$ be an exponential family bandit model. Let $\alpha \in [1, e/2]$ and $r(t) = O(t^{\alpha})$. Using Chernoff's stopping rule with $\beta(t,\delta) = \log(r(t)/\delta)$ and the sampling rule D-Tracking,

\begin{equation*}
\limsup_{\delta \to 0} \frac{\mathbb{E}_{\bm{\mu}}[T_{\delta}]}{\log(1/\delta)} \leq \alpha T^*(\mu), \quad \text{with } T^*(\mu) \text{ as defined in equation \ref{eqn:ComplexityTrackStopStrategy}}.
\end{equation*}
\end{theorem}

## Conclusion

Overall, the two papers presented in this chapter make significant contributions to identifying the top-$1$ or the top-$m$ arms in a multi-armed bandit setting. Their contribution is primarily based on including the Kullback-Leibler divergence and Chernoff information as information theoretic measures in the derivation of bounds and the matching algorithms. Thereby, they go beyond a sum of squared gaps which is most relevant only in the case of Gaussian distributions. While the contribution in the fixed budget setting is limited to the two-armed bandit case, the authors derive general lower bounds in the fixed confidence setting. All bounds presented are primarily based on Lemma \ref{theorem:KaufmannEtAlLemma1}. Lemma \ref{theorem:KaufmannEtAlLemma1} is able to measure the "transportation cost" (see Garivier and Kaufmann, 2016) of moving from one distribution to another one in terms of Kullback-Leibler divergence. It is the bound on the expected likelihood ratio term that appears when changes in distribution are applied.

In the case of the fixed budget top-$m$ arm identification problem, we see that the derivation of a lower bound on the complexity term is mostly limited by the difficulty of defining an alternative problem $\nu'$ to $\nu$ that fits the Lemma \ref{theorem:KaufmannEtAlLemma1}. But given the intuition that in top-$m$ problems the $m$-th best arm and the $(m+1)$-th best arm define an implied threshold between the two groups of arms, we were able to apply the ideas of Kaufmann et al. (2016) to our fixed budget thresholding bandit problem. In contrast to their problem, arms are classified against the known threshold $\tau$, which makes it possible to define an alternative problem $\nu'$ in which we do not have to change arms with respect to each other, but only with respect to $\tau$. This has lead us to the simple lower bound on the problem complexity for the thresholding bandit problem with exponential family distributions given in Corollary \ref{corollary:LBKTBP}.

In the best arm identification problem discussed by Garivier and Kaufmann (2016), we were introduced to the idea of comparing each arm to the best arm, and evaluating the difference between the best arm and the other arms based on the Generalized Likelihood Ratio test. This idea seems transferrable to the thresholding bandit problem. Indeed, it might even be easier as the best arm does not have to be estimated; instead, we would compare each arm to the threshold $\tau$. Intuitively, one can frame this problem for each arm individually as an hypothesis test between the maximum likelihood estimate of the mean and the threshold $\tau$. We will come back to this idea in \autoref{sec:LikelihoodRatioForExponentialFamily}.

Lastly, we observe that Garivier and Kaufmann (2016) support their algorithm's main strategy by an additional D-Tracking rule which protects them against poor estimates in the beginning of the sampling. Such rule should be mostly unnecessary, for example, the cumulative regret setting, as algorithms there rely on upper confidence bound estimates. This is not the case for the Track-and-Stop strategy introduced in Garivier and Kaufmann, and neither for our Simple Likelihood Ratio strategy introduced in \autoref{sec:1DSLR}. In \autoref{sec:Simulation2}, we show empirically that the performance of the Simple Likelihood Ratio strategy can be further improved by adding the D-Tracking rule.

\newpage

# Extensions to the Thresholding Bandit Problem \label{chap:Extensions}

## Introduction \label{sec:ExtensionsIntroduction}

In \autoref{chap:ThresholdingBanditProblem} we introduced the thresholding bandit problem as proposed in Locatelli et al. (2016) alongside the APT algorithm. We saw that the upper bound for the APT algorithm matches the lower bound for the thresholding bandit problem given in Locatelli et al. (2016). The algorithm also exhibits favorable characteristics such as its upper bound holding at any stopping time, and that it does not require any parameter to be tuned upfront. However, we have seen that the index $B^{APT}$ and the complexity term $H$ employed by the APT algorithm are the most effective for $R$-sub-Gaussian distributions. Both are related to the KL-divergence between two Gaussian distributions of equal variance. Given that this sub-Gaussian approximation can be poor in certain cases, we motivated information theoretic measures in an algorithm's index (compare \autoref{sec:DiscussionLocatelli}).

Information theoretic measures are one way of taking not only the sample mean into account to estimate differences between distributions -- as done by the APT algorithm. Another logical extension is to consider higher moments of a distribution, like its variance. After all, arms do not necessarily have the same variance, even if they are sub-Gaussian (two arms might be $R_1$-sub-Gaussian and $R_2$-sub-Gaussian). And given two arms for which we estimate the same mean, we would prefer to draw more samples from the arm with the larger variance. The simplest example of this are two Normal distributions with different variances.

For these reasons, in the following, we first present in \autoref{sec:VarianceBasedAlgorithms} two algorithms which incorporate empirical variances as described in Mukherjee et al. (2017) and Zhong et al. (2017). We then go back to the information theoretic ideas and introduce a new algorithm based on the concept of likelihood ratios in \autoref{sec:LikelihoodRatioForExponentialFamily}. Finally, we compare the performance of different algorithms based on simulations and real world data sets in \autoref{chap:Experiments}.

## Variance Based Algorithms for the Thresholding Bandit \label{sec:VarianceBasedAlgorithms}

### Augmented-UCB Algorithm \label{sec:AugUCBAlgorithm}

Combining learnings from the APT and other algorithms, Mukherjee et al. (2017) propose the Augmented-UCB (AugUCB) algorithm based on variance estimates for the thresholding bandit problem. The setting they consider is very similar to the setting in Locatelli et al. (2016). The main difference being that, while Locatelli et al. (2016) assume $R$-sub-Gaussian reward distributions, Mukherjee et al. (2017) assume rewards to be bounded in $[0,1]$, which makes the AugUCB applicable to fewer distributions. The AugUCB algorithm itself differs significantly from the APT algorithm. The two main ideas it employs are, first, upper confidence bounds based on the Bernstein inequality and empirical variance estimates to compute the index based on which an arm is chosen at every round. Second, if an arm estimate is "far enough" away from the threshold as measured by the confidence bounds, it will be eliminated, that is, no longer be pulled until the end of the budget. Both ideas are based not only on the empirical mean estimate $\hat{\mu}_i(t)$, but also on a term $s_i$ that is increasing linearly in the estimated standard deviation. While the APT algorithm plays the arm minimizing the quantity $\sqrt{T_i(t)} \hat{\Delta}_i(t)$, AugUCB plays the arm minimizing the quantity $\hat{\Delta}_i(t) - 2s_i$, where $s_i$ is based on the variance estimate $\hat{v}_i$, and an exploration term $\psi_m$ linear in $T$, as well as an elimination parameter $\rho$ (compare Algorithm 1 in Mukherjee et al., 2017):

\begin{equation*}
s_i = \sqrt{\frac{\rho \psi_m (\hat{v}_i + 1) \log(T \epsilon_m)}{4n_i}}.
\end{equation*}

Consequently, the authors also introduce complexity measures which take the variance into account. From Gabillon et al. (2011) they use $H_{\sigma, 1}= \sum_{i=1}^K \frac{\sigma_i + \sqrt{\sigma_i^2 + (16/3)\Delta_i}}{\Delta_i^2}$. Furthermore, with $\stackrel{\sim}{\Delta}_i^2 = \frac{\Delta_i^2}{\sigma_i + \sqrt{\sigma_i^2 + (16/3)\Delta_i}}$, where $\stackrel{\sim}{\Delta}_{(i)}$ is an increasing ordering of $\stackrel{\sim}{\Delta}_i$, the authors define a complexity corresponding to $H_{CSAR,2}$ in Chen et al. (2014):

\begin{equation*}
H_{\sigma,2} = \max_{i \in \mathcal{A}} \frac{i}{\stackrel{\sim}{\Delta}_{(i)}^2}.
\end{equation*}

While the latter complexity appears in the provided upper bound on the AugUCB algorithm's expected loss, it could be replaced by $H_{\sigma, 1}$ since $H_{\sigma, 2} \leq H_{\sigma,1}$. $H_{\sigma,1}$ is more easily compared to the complexity $H$ defined in Locatelli et al. (2016). Essentially, every arm's complexity contribution is additionally scaled by a term increasing in the arm's variance (see also Figure 3 below).

In the case of the APT algorithm the number of times a specific arm has been drawn $T_i$ measures how concentrated the estimated mean is around the true mean (and thus for large $T_i$ an arm is no longer drawn even if $\hat{\mu}$ is close to $\tau$). This exploration is regulated by the combination of the variance estimate $\hat{v}_i$ and the exploration factor $\psi_m$ in the case of the AugUCB algorithm. The terms ensure that certain arms are no longer pulled when the algorithm is sufficiently confident in the location of the true mean. Indeed, the AugUCB algorithm goes a little further than the APT algorithm in that it stops to sample arms when a separation condition (mean separated from threshold) is estimated to hold. The same separation condition is used in the authors' upper bound proof and is reminiscient of the separation condition used in the proof of the upper bound in Locatelli et al.'s (2016) (compare \autoref{sec:CheckSeparationCondition}).

Therefore, when using a very similar favorable event in the proof of their upper bound, Mukherjee et al. (2017) have to bound the probability that an arm is not removed from the active set after a certain number of rounds -- even though the arm is above (respectively below) the threshold. If this probability is small, then the algorithm is able to be confident in an arm's true mean by employing variance estimates. In order to derive the bound, Mukherjee et al. (2017) use the Bernstein inequality.

It has to be noted that Mukherjee et al. (2017) have to assume that the distributions have rewards bounded in $[0,1]$ in order to derive the bound on the probability that an arm has not been eliminated correctly in round $T_i$. Thus, the distributions for which their upper bound holds are a subset of the $R$-sub-Gaussian distributions considered in Locatelli et al. (2016). It would otherwise not be possible to upper bound the $s_i$ term by $\Delta_i/4$ as Mukherjee et al. do in their proof: With bounded rewards, the arms' variances are also bounded in $[0,1]$ such that $\hat{v}_i \in [0,1]$. Yet, the simulations by Mukherjee et al. are performed on Gaussian arms with variance smaller than one, which at least makes empirical variances smaller 1 likely.

Based on this assumption, the authors show that the probability that an arm has not been eliminated after $m_i$ rounds is given by (compare equation (2) in Mukherjee et al., 2017):

\begin{equation} \label{Mukherjee2017Equation2}
\mathbb{P}(\hat{\mu}_i > \mu_i + 2s_i) \leq \mathbb{P} (\hat{\mu}_i > \mu_i + 2\bar{s}_i) + \mathbb{P}(\hat{v}_i \geq \sigma^2_i + \sqrt{\rho \epsilon_{T_i}})
\end{equation}

Compare Mukherjee et al. (2017) for the notation. Then it becomes clear that the left part of the RHS gives the probability that the mean estimate is poor, while the right term gives the probability that the variance is overestimated. The first term is bounded by the Bernstein inequality. In the cumulative regret setting, there exists a similar decomposition for algorithms based on upper confidence bounds. There, the expected number of pulls of a suboptimal arm is bounded. Compare for example chapter 3.1 in Kaufmann et al. (2012b), in which the expected number of draws of an suboptimal arm is decomposed into the probability that the optimal arm's mean is underestimated, and the probability that the upper confidence bound of the suboptimal arm is still very large. To give bounds for UCB algorithms, certain convergence rates are shown for these two probabilities. Mukherjee et al. do similarly for their upper confidence algorithm in the thresholding setting.

### Anytime Parameter-free Empirical Variance Guided Algorithm \label{sec:EVTAlgorithm}

A second algorithm for the thresholding bandit problem employing empirical variance estimates is introduced in Zhong et al. (2017). While they propose different algorithms for different setting of the thresholding bandit problem, we focus here on their strategy that can be compared directly with the problem described in Locatelli et al. (2016).

This strategy is the parameter-free version of their Empirical Variance Guided Algorithm (EVT$_{pf}$). The EVT$_{pf}$ algorithm is an extension of the APT algorithm that takes the empirical variance of the arms into account and deduces an upper bound through empirical Bernstein bounds in a proof following the model of the one described in Locatelli et al. (2016) (compare \autoref{sec:UpperBoundProof}).

In order to describe $EVT_{pf}$, we adapt a formulation used in Zhong et al. (2017) to describe the index used by the algorithms in a more generally. Write the index $B_i(t+1)$ of arm $i$ at round $t$ as

$$
B_i(t+1) = S_i(t) D_i(t),
$$

where $D_i(t)$ is some measure of the distance between the mean $\mu_i$ or the distribution $\nu_i$ of arm $i$ from the threshold $\tau$, while $S_i(t)$ is some confidence value measuring how confident we can be in $D_i(t)$. Thus, for the APT algorithm, if we let $\epsilon = 0$ as in Zhong et al. (2016), we have

$$
D_i^{APT}(t) = \hat{\Delta}_i(t) = |\hat{\mu}_i(t) - \tau| \quad \text{and} \quad S_i^{APT}(t) = \sqrt{T_i(t)}.
$$

We see that $D_i(t)$ is the probably most straightforward distance between the sample mean and the threshold, while $S_i(t)$ increases over time, as we become more confident in our sample mean, and thus ensures that we do not pull the same arm forever. Based on this, we can now easily compare APT to EVT$_{pf}$ for which the index can be written as:

$$
D_i^{EVT_{pf}}(t) = \sqrt{\hat{\sigma}^2_{i}(t) + \hat{\Delta}_i(t)} - \hat{\sigma}_i(t) \quad \text{and} \quad S_i^{EVT_{pf}} = \sqrt{T_i(t)}
$$

We see that both algorithms employ the same confidence term. The empirical variance describes the shape of the distribution and thus is a measure of how far the distribution is away from the threshold. It consequently makes sense to have it appear in $D_i^{EVT}$. And in contrast to the APT algorithm, the EVT$_{pf}$ algorithm takes more information about the arms into account than just their mean. As mentioned previously, this should especially be valuable whenever arms have different variances. Roughly similar to how the Kullback-Leibler divergence measures distances between distributions, $D_i^{EVT}$ including the empirical variance serves as a crude measure of distance between threshold and distribution. We note that for a Bernoulli distributed random variable, $D_i^{EVT}$ is always decreasing as the mean approaches the threshold (compare figure 2). Similar should hold in general for bounded variables as the variance and the mean are linked by the distance of the mean to the bound of the support (in contrast to, for example, a normal distributed random variable where a change in the mean does not correspond to a change in the variance). This is important as the bounds of the EVT$_{pf}$ algorithm only hold for random variables bounded in $[0,1]$. Thus, it is limited to the same set of distributions as the AugUCB algorithm.

```{r EVT_Index_Bernoulli, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", out.width='80%', fig.cap = "This graph displays the distance part of the index of the APT, SLR, and EVT algorithms. While EVT's distance is based on both the empirical mean and empirical variance, the empirical mean determines the variance in the case of Bernoullis. We see that the distance of both EVT and SLR has an asymmetric behavior."}
bfun <- function(mu, tau) sqrt((mu*(1-mu))^2 + abs(tau-mu)) - mu*(1-mu)
KL_ber <- function(p,q) p*log(p/q) + (1-p)*log((1-p)/(1-q))
KL_gauss <- function(p,q) 2*(p-q)^2

evtdf <- data.frame(mu = seq(0.0000001,0.5,length.out = 10000),
                    tau = rep(0.1, 10000)) %>%
  mutate(APT = KL_gauss(mu, tau),
         SLR = KL_ber(mu, tau),
         EVT = bfun(mu, tau)) %>%
  gather(key = Algorithm, value = Distance, -mu, -tau) %>%
  mutate(tau = as.character(tau))

ggplot(evtdf, aes(x = mu, y = Distance, 
                  group = Algorithm, color = Algorithm,
                  linetype = Algorithm)) +
  geom_line() +
  labs(#title = "Comparison of Distance Measures by Algorithm",
       subtitle = "For Bernoulli arms and a threshold at 0.1.",
       x = "Empirical Mean Estimate") +
  theme_bw() +
  theme(legend.position = "bottom")
```

By defining a new complexity measure $H^{\text{EVT}_{pf}}$ for the EVT$_{pf}$ algorithm, Zhong et al. (2017) are able to derive an upper bound on the expected loss of the algorithm. Let the complexity increase in the variance and decrease in the distance of the mean from the threshold such that

$$
H^{\text{EVT}_{pf}} = \sum_{i=1}^K h_i^{\text{EVT}_{pf}} = \sum_{i=1}^K (\sigma_i^2\Delta_i^{-2} + \Delta_i^{-1}).
$$

Then one can state the following theorem with notation as in \autoref{sec:Setup}.

\begin{theorem}[Theorem 5, Zhong et al., 2017] 
\label{theorem:ZhongEtAl2017Theorem5}
Let $K > 0$, $T > 2K$, and consider a thresholding bandit problem. Assume that all arms of the problem are bounded in $[0,1]$. Then the probability of an error is bounded as

\begin{equation*}
\mathbb{E}[\mathcal{L}[T]] \leq TK \exp\Big(-C\frac{T}{H^{\text{EVT}_{pf}}}\Big),
\end{equation*}

where $C$ is some positive constant.
\end{theorem}

The authors approach the proof very similar as Locatelli et al. (2016) in the proof of their lower bound. They define favorable events that hold with large probability, and ensure that the same separation condition as in Locatelli et al. holds given that each arm is pulled sufficiently often. Additionally to a favorable event that bounds the deviation between sample mean and population mean, they define a second event on which the sample standard deviation is close to the population standard deviation. This is bounded in large probability by an empirical Bernstein inequality (Maurer and Pontil, 2009). The union of the two events then serves as the favorable event in the remainder of their proof.

Given a lower bound adapted to the new complexity $H^{\text{EVT}_{pf}}$, Zhong et al. (2017) show that the EVT$_{pf}$ algorithm is optimal in that its upper bound matches the lower bound. Furthermore, they compare their algorithm in experiments against the APT algorithm. The experiments use data generated from uniform distributions. Each experiment compares 100 arms with variances picked at random from a range of possible variances. Over three experiments, the range of variances is varied from small to large. They report the largest advantage of their variance based algorithms in comparison to APT when the variances are very small; given they only consider problems bounded on a support of $[0,1]$, $\sigma_i \ll 1$. In this case the problem complexity $H^{\text{EVT}_{pf}}$ can be much smaller than $H^{APT}$ -- the EVT$_{pf}$ algorithm profits more from distributions that are simple to estimate in terms of their variance than the APT algorithm does.

When comparing the two complexities for Bernoulli random variables against a threshold, we see that the smaller complexity of the EVT$_{pf}$ algorithm implies in particular that EVT$_{pf}$ could be able to cope with distributions of parameter $\mu$ close to $\tau$ much better than APT (compare Figure 3).

```{r Hevt_Hapt_Comparison, echo = FALSE, message = FALSE, fig.align="center", out.width = '70%', fig.cap = "Comparing the complexity terms used in the APT, AugUCB, and EVT algorithms, we observe that the problem complexity of the EVT algorithm increases much less rapidly than for the other algorithms. Thus, the EVT should be able to cope with arms close to the threshold better than the APT algorithm. The complexity displayed for the AugUCB is an upper bound on the complexity used in the upper bound of Mukherjee et al. (2017)."}
hevt <- function(mu, tau) {
  (mu*(1-mu))^2 * (abs(mu-tau))^{-2} + abs(mu-tau)^{-1}
}

hapt <- function(mu, tau) {
  abs(mu-tau)^{-2}
}

haugucb <- function(mu, tau) {
  vars <- mu*(1-mu)
  (sqrt(vars) + sqrt(vars + (16/3)*(abs(mu-tau))))/(abs(mu-tau)^2)
}

hcompdf <- data.frame(mu = seq(0.1,0.9,length.out = 10000),
                      tau = rep(0.5, 10000)) %>%
  mutate(APT = hapt(mu, tau),
         EVT = hevt(mu, tau),
         AugUCB = haugucb(mu, tau)) %>%
  gather(key = Algorithm, value = Complexity, -mu, -tau)

ggplot(hcompdf, aes(x = mu, y = Complexity, group = Algorithm,
                    color = Algorithm, linetype = Algorithm)) +
  geom_line() +
  geom_vline(aes(xintercept = 0.5), linetype = 2) +
  ylim(c(0,400)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(x = "Mean", title = "Comparison of Complexity Terms",
       subtitle = "Complexity of single Bernoulli arm for APT and EVT and a threshold at 0.5.")
```

## Likelihood Ratio for Sequential Sampling \label{sec:LikelihoodRatioForExponentialFamily}

We now present a new sequential sampling algorithm for the fixed budget thresholding bandit problem that builds on the idea of the likelihood ratio statistic for hypothesis tests. We first describe the general idea of likelihood ratio based tests shortly, to then introduce the new algorithm for the thresholding bandit problem, which we call Simple Likelihood Ratio (SLR) algorithm.

### Likelihood Ratio-based Hypothesis Tests \label{sec:LRTests}

In a standard parametric testing problem, we have an i.i.d. sample $\bm{x} = (x_1, ..., x_T)$, where each $x_s$ is a realization from the random variable $X$. We assume that $X$ follows some probability distribution $Q(\theta)$ paramaterized by a vector $\theta$. Let $q(x;\theta)$ be the density function corresponding to the distribution. We would like to test whether an alternative parameterization is more likely than the parameterization given by the null hypothesis. That is, we test

$$
H_0: \quad \theta = \theta_{H_0} \qquad \text{vs.} \qquad H_1: \quad \theta = \theta_{H_1}
$$

In order to assess whether the $H_0$ hypothesis can be rejected, we compute the likelihood $L(x|\theta)$ of the observed sample under both parameterizations: $L_{H_0} = L(x|\theta = \theta_{H_0}) = \prod_{s=1}^T q(x_s|\theta_{H_0})$ and $L_{H_1} = L(x|\theta = \theta_{H_1}) = \prod_{s=1}^T q(x_s|\theta_{H_1})$. The likelihood ratio test statistic is then given as

$$
LR_T(\theta_{H_0}|| \theta_{H_1}) = \frac{L_{H_0}}{L_{H_1}} = \prod_{s=1}^T\frac{q(x_s|\theta_{H_0})}{ q(x_s|\theta_{H_1})}.
$$

and is used in the optimal test of two hypotheses as shown by the Neyman-Pearson lemma (see chapter 11, Theorem 11.7.1 in Cover and Thomas, 2006).

The intuition behind the test is that, if the true parameterization is $H_1$ and not $H_0$, then the likelihood $L_{T,H_1}$ should be larger than $L_{T,H_0}$. A small enough likelihood ratio would thus lead us to a rejection of the null hypothesis, while a large likelihood ratio is evidence in favor of it.

Furthermore, the likelihood ratio is inherently linked to information theoretic quantities and divergence measures. Assuming that $Q$ is a discrete distribution with support $\chi$, one can write the negative log-likelihood ratio as (compare Cover and Thomas, 2006)

\begin{align*}
-\log(LR_T(\theta_{H_0}|| \theta_{H_1}) & = \sum_{i=1}^T \log(\frac{q(x_s|\theta_{H_0})}{ q(x_s|\theta_{H_1})}) \\
& = \sum_{a \in \chi} T q(a|\hat{\theta}) \log(\frac{q(a|\theta_{H_0})}{ q(a|\theta_{H_1})}) \\
& = \sum_{a \in \chi} T q(a|\hat{\theta}) \log(\frac{q(a|\theta_{H_0})}{ q(a|\theta_{H_1})}\frac{q(a|\hat{\theta})}{ q(a|\hat{\theta})}) \\
& = T \KL(q(\hat{\theta})||q(\theta_{H_0})) - T \KL(q(\hat{\theta})||q(\theta_{H_1})),
\end{align*}

where $\hat{\theta}$ is the maximum likelihood estimate of the parameter $\theta$ given the sample $\bm{x}$. If we hold the $H_0$ hypothesis fixed and set $\theta_{H_1} = \hat{\theta}$, then with $\KL(Q(\hat{\theta})||Q(\hat{\theta})) = 0$:

$$
-\log(LR_T(\theta_{H_0}|| \theta_{H_1})) = -\log(LR_T(\theta_{H_0}|| \hat{\theta})) = T\KL(Q(\hat{\theta}||Q(\theta_{H_0}))
$$

Furthermore, we have for any distributions $Q_{H_0}$ and $Q_{H_1}$ for a sample $\bm{x}$ sampled from the $H_0$ distribution, that

$$
\frac{1}{T} \log(LR_T(\theta_{H_0}||\theta_{H_1})) \rightarrow \KL(Q(\theta_{H_0})||Q(\theta_{H_1})) \quad \text{in probability},
$$

by the law of large numbers (or asymptotic equipartition property) for the Kullback-Leibler divergence as in Theorem 11.8.1 in Cover and Thomas (2006). Lastly, based on the characteristics shown before, for frequentist tests the Kullback-Leibler divergence between $Q(\theta_{H_0})$ and $Q(\theta_{H_1})$ gives the best possible exponent for the error of the second type given a fixed error of the first type. This is the result of the Chernoff-Stein Lemma (Theorem 11.8.3, Cover and Thomas, 2006), which is reminiscient of the complexity $\kappa_B$ derived by Kaufmann et al. (2016) and as described in \autoref{chap:KullbackLeiblerPureExploration}:

\begin{theorem}[Chernoff-Stein Lemma] \label{theorem:ChernoffStein}
Let $X_1, X_2, ..., X_n$ be i.i.d. distributed according to $Q$. Consider the hypothesis test between two alternatives, $Q=P_1$ and $Q=P_2$, where $\KL(P_1||P_2) < \infty$. Let $A_n \subseteq \chi^n$ be an acceptance region for hypothesis $H_1$. Let the probabilities of error be $\alpha_n = P_1^n(A_n^C)$, and $\beta_n = P_2^n(A_n)$, and for $0< \epsilon <1/2$, define $\beta_n^{\epsilon} = \min_{A_n \subseteq \chi^n, \alpha_n < \epsilon} \beta_n$. Then

$$
\text{lim}_{n\rightarrow \infty} \frac{1}{n} \log \beta_n^{\epsilon} = - \KL(P_1||P_2)
$$
\end{theorem}

Indeed, our Corollary \autoref{corollary:LB1TBP} for the one-armed thresholding bandit problem corresponds directly to this result, and shows that the thresholding bandit can be viewed as a sequential testing of separate hypotheses.

### Likelihood Ratio as Index in Sequential Sampling

We now employ the concept of the likelihood-ratio test to a new sequential sampling strategy for the thresholding bandit problem. In contrast to the cumulative regret problem or the best arm identification problem, the thresholding bandit problem offers a natural $H_0$ hypothesis for the arms' distributions: the threshold $\tau$. Given that we want to classify the arms against the threshold, and given that $\tau$ is known and fixed, we can set a null hypothesis that is a function of $\tau$. In the simplest case, the distributions are parameterized by their mean $\theta_i = \mu_i$, and we set $\theta_{i,H_0} = \tau$ for all arms. The likelihood ratio then quantifies how "significant" of a difference there is between the threshold $\tau$ and the current sample mean $\hat{\mu}_i(t)$ of an arm, assuming distributions for which the maximum likelihood estimate of the mean is given by the empirical mean. The likelihood ratio will be smaller than 1 and decrease in sample size even if the threshold and sample mean are constant.

These characteristics make the likelihood ratio a natural index for adaptive sampling strategies (compare also Chernoff, 1958; Garivier and Kaufmann, 2016). Consider the setting described in \autoref{sec:Setup} and \autoref{sec:TheAPTAlgorithm}, and write as index:

$$
B_i^{LR}(t+1) = LR(\tau|| \hat{\mu}_i(t)) = \prod_{i=1}^T\frac{q(x_i|\tau)}{ q(x_i|\hat{\mu}_i(t))}.
$$

We have $B_i^{LR}(t) \leq 1$, $t \uparrow \rightarrow B_i^{LR}(t) \downarrow$, and $|\hat{\mu}_i(t) - \tau| \uparrow \rightarrow B_i^{LR}(t) \downarrow$. Consequently, by choosing the arm that maximizes $B_i^{LR}(t)$, we choose an arm for which we either have not collected many samples, or for which the sample mean is close to the threshold $\tau$. In this sense, a strategy based on the likelihood ratio is reminiscient of the index employed by the APT algorithm. Indeed, assuming the samples are normally distributed with equal variance for all arms under both $H_0$ and $H_1$, $-\log(B_i^{LR}(t))$ equals the index in APT. Why then would we prefer the likelihood based strategy?

Given that the APT index and the likelihood-ratio based index are only equivalent for normally distributed arms with equal variances, the new strategy promises efficiency improvements whenever the true distribution is known and different from the normal distribution. The most important example are Bernoulli arms with small means as motivated above. While the likelihood ratio can in theory be used for any distributions, and arms could even have different classes of distributions, the class of distribution has to be known upfront for every arm. Given the role of the threshold $\tau$ and the goal of classification of the mean against $\tau$, the likelihood ratio strategy is most naturally employed on distributions parameterized by their mean. In the following descriptions, we thus focus on one-parameter distributions of the exponential family. Additionally, we also propose a version for the Normal distribution with unknown mean. Before we describe the algorithm, we show that the likelihood ratio has a convenient form for this distribution family.

### Likelihood Ratio for One-Parameter Exponential Family of Distributions \label{sec:LRforUnivariateExponentialFamily}

We now focus on a thresholding bandit problem where the arms are distributed according to $\nu_{\theta_1}, ..., \nu_{\theta_K}$, which are all from the same canonical exponential family of distributions with scalar parameter $\theta_i \in \Theta$. The probability density function of the absolutely continuous distribution $\nu_{\theta_i}$ with regard to a dominating measure $\rho$ on $\mathbb{R}$ is given by $q(x|\theta) = h(x) \exp(T(x)\theta - b(\theta))$. As described in Cappé et al. (2013), citing Lehmann and Casella (1998), the derivative $b'(\theta)$ of $b(\theta)$ is an increasing continuous function such that the expected value $\mathbb{E}(\nu_{\theta}) = b'(\theta)$ for all $\theta \in \Theta$. Given that $b'(x)$ is strictly convex, we have a one-to-one relation between $b'(\theta)$ and $\mathbb{E}(\nu_{\theta})$ through the continuous inverse $b'^{-1}(\theta)$. Consequently, one can just as well parameterize the distribution by its mean $\mu = \mathbb{\nu_{\theta}}$ and write $\nu_{b'^{-1}(\mu)}$ with the probability density function (compare Cappé et al., 2013)

$$
q(x|b'^{-1}(\mu)) = h(x) \exp(S(x) b'^{-1}(\mu) - b(b'^{-1}(\mu))).
$$

This lets us write the likelihood ratio for the exponential family distributions as

$$
LR_T(b'^{-1}(\mu_0)||b'^{-1}(\mu_1)) = \prod_{s=1}^T \frac{\exp(S(x_s) b'^{-1}(\mu_0) - b(b'^{-1}(\mu_0)))}{\exp(S(x_s) b'^{-1}(\mu_1) - b(b'^{-1}(\mu_1)))}
$$

or if we write it as negative log-likelihood ratio:

$$
-\log(LR_T) = T\Big(b\big(b'^{-1}(\mu_0)) - b(b'^{-1}(\mu_1)\big)\Big) - \Big(b'^{-1}(\mu_0)-b'^{-1}(\mu_1)\Big)\sum_{i=1}^nS(x_i)
$$

As we will see now, the last equation takes a convenient form for many common distributions.

**Example 1: Bernoulli Distribution**

We have $S(x_s) = x_s$, $\theta = \log(\mu/(1-\mu))$, $\Theta = (0,1)$, $b(\theta) = \log(1+\exp(\theta))$, and $\mu \in (0,1)$. Then $b'(\theta) = \mu$, and we can write

\begin{align*}
-\log(LR_T(\mu_0||\mu_1)) & = T(\log\Big(\frac{1-\mu_0}{1-\mu_1}\Big)) - \Big(\log(\frac{\mu_0}{1-\mu_0}) - \log(\frac{\mu_1}{1-\mu_1})\Big)\sum_{s=1}^{T}x_i \\
& = n \Big( (1-\hat{\mu})\log\Big(\frac{1-\mu_1}{1-\mu_0}\Big) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}) \Big)
\end{align*}

We observe that, if we let $\mu_0 = \tau$ and $\mu_1 = \hat{\mu}$ (compare \autoref{sec:LRTests}), then

$$
-\log(LR_T(\tau||\hat{\mu})) = T\KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) =  = T \KL(\hat{\mu}||\tau) = n \Big(\hat{\mu}\log\big(\frac{\hat{\mu}}{\tau}\big) + (1-\hat{\mu})\log\big(\frac{1-\hat{\mu}}{1-\tau}\big)\Big)
$$

**Example 2: Poisson Distribution**

We have $S(x_s) = x_s$, $\theta = \log(\mu)$, $\Theta = \mathbb{R}$, $b(\theta) = \exp(\theta)$, and $\mu \in (0, +\infty)$. Then $b'(\theta) = \mu$, and we can write:

$$
-\log(LR_T(\mu_0||\mu_1)) = T(\mu_0 - \mu_1) - (\log(\mu_0) - \log(\mu_1))\sum_{s=1}^T x_s = T\Big((\mu_0 - \mu_1) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}\Big)\Big)
$$

We observe that, if we let $\mu_0 = \tau$ and $\mu_1 = \hat{\mu}$ (compare \autoref{sec:LRTests}), then

$$
-\log(LR_T(\tau||\hat{\mu})) = T \KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = T \KL(\hat{\mu}||\tau) = T\Big((\tau - \hat{\mu}) + \hat{\mu}\log\Big(\frac{\hat{\mu}}{\tau}\Big)\Big)
$$

**Example 3: Normal distributions with known variance $\sigma^2$**

We have $S(x_s) = x_s$, $\theta = \mu/\sigma^2$, $\Theta = \mathbb{R}$, $b(\theta) = \sigma^2\theta^2/2$, $\mu \in \mathbb{R}$. Then $b'(\theta) = \mu$, and we can write

$$
-\log(LR_T(\mu_0||\mu_1)) = T (\frac{\mu_0^2}{2\sigma^2} - \frac{\mu_1^2}{2\sigma^2}) - (\frac{\mu_0}{\sigma^2} - \frac{\mu_1}{\sigma^2})\sum_{s=1}^Tx_s = T\frac{1}{2\sigma^2}\Big(\mu_0^2-\mu_1^2 - 2\mu_0\hat{\mu} + 2\mu_1\hat{\mu}\Big)
$$

And again, even though the distribution is continuous (compare \autoref{sec:LRTests}), if we let $\mu_1 = \hat{\mu}$ and $\mu_0 = \tau$, then

$$
-\log(LR_T(\tau||\hat{\mu})) = T\KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = T \KL(\hat{\mu}||\tau) = T\frac{(\hat{\mu} - \tau)^2}{2\sigma^2}.
$$

**Example 4: Gamma distributions with known shape parameter $\alpha$**

We have $S(x_s) = x_s$, $\theta = -\alpha/\mu$, $\Theta = (-\infty, 0)$, $b(\theta) = -\alpha \log(-\theta)$, $\mu \in (0,+\infty)$. Then $b'(\theta) = -\mu$, and we can write

$$
-\log(LR_T(\mu_0||\mu_1)) = T\alpha(-\log(\frac{\alpha}{\mu_0}) - \log(\frac{\mu_1}{\alpha})) - (-\frac{\alpha}{\mu_0} + \frac{\alpha}{\mu_1})\sum_{s=1}^{T}x_s = T\alpha \Big(\log\big(\frac{\mu_0}{\mu_1}\big) - \frac{\hat{\mu}}{\mu_1} + \frac{\hat{\mu}}{\mu_0}\Big).
$$

And again, even though the distribution is continuous (compare \autoref{sec:LRTests}), if we let $\mu_1 = \hat{\mu}$ and $\mu_0 = \tau$, then

$$
-\log(LR_T(\tau||\hat{\mu})) = T \KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = T \KL(\hat{\mu}||\tau) = T\alpha \Big(\frac{\hat{\mu}}{\tau} - 1 - \log\big(\frac{\hat{\mu}}{\tau}\big)\Big).
$$

For $\alpha = 1$, the Gamma distribution is equal to the Exponential distribution.

Given these results for the exponential family of distributions, we can now define the Simple Likelihood Ratio algorithm.

## 1D-Simple Likelihood Ratio Algorithm \label{sec:1DSLR}

We now define the Simple Likelihood Ratio Algorithm for the one-parameter exponential family of distributions (1D-SLR). Consider again the setting in \autoref{sec:Setup} though without the assumption of $R$-sub-Gaussian distributions. Instead, assume that the distributions $\nu_i$ of arms $i \in \mathbb{A}$ are all the same distribution from the one-parameter exponential family of distributions. We write this class of distributions as $\nu(\mu, \theta)$. While $\mu$ is the unknown mean parameter, $\theta$ describes potentially *known* parameters, as for example $\sigma_1, ..., \sigma_K$ for Normal distributions, or $\alpha_1, ..., \alpha_K$ for Gamma distributions. Additionally to $\nu(\mu, \theta)$, the 1D-SLR algorithm receives as input the definition of the problem $(\tau, \epsilon)$. As initialization, the algorithm pulls each arm once. At time $t>K$, the algorithm updates the number of pulls for arm $i$ at time $t$, $T_i(t)$, for each arm, as well as the empirical mean $\hat{\mu}_i(t)$ of arm $i$ after $T_i(t)$ pulls. The empirical mean is calculated for each arm $k\in \{1,...,K\}$ as $\hat{\mu}_i(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)} x_{i,s}$, with $T_i(t) = \sum_{s=1}^{t}\mathbb{I}(I_s = i)$. It serves as maximum likelihood estimate of the arm's mean. As in Locatelli et al. (2016), $X_{i,s}$ is the sample received when pulling arm $i$ for the $s$-th time. Depending on the distribution, the algorithm then computes the index

\begin{align*}
B_i^{SLR}(t+1) & = \mathbb{I}(\hat{\mu}_i(t)\geq \tau) \Big(-\log \big(LR_{T_i(t)}(\tau-\epsilon||\hat{\mu}_i(t))\big)\Big) \\
& \quad +\: \mathbb{I}(\hat{\mu}_i(t) <\tau) \Big(-\log\big(LR_{T_i(t)}(\tau+\epsilon||\hat{\mu}_i(t))\big)\Big) \\
& = T_i(t) \Big( \mathbb{I}(\hat{\mu}_i(t) \geq \tau)\cdot \KL(\hat{\mu}_i(t)||\tau+\epsilon) \\
& \quad \quad \quad +\: \mathbb{I}(\hat{\mu}_i(t) < \tau)\cdot \KL(\hat{\mu}_i(t)||\tau+\epsilon) \Big) \Alignyesnumber\label{eqn:SLRIndex}.
\end{align*}

The next arm to be pulled is the minimizer of the index $B_i^{SLR}(t+1)$, $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$. At the end of the horizon $T$, the algorithm outputs the set of arms $\hat{S}_\tau = \{i \in \mathbb{A},  \hat{\mu}_i(T) \geq \tau\}$.

\IncMargin{1em}
\begin{algorithm}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$, class of distributions $\nu(\mu, \theta)$, $\theta_i \: \forall i \in \mathbb{A}$}
\Output{$\hat{S}_\tau = \{i \in \mathbb{A},  \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i\leq K} B_i^{SLR}(t)$ as in \autoref{eqn:SLRIndex}}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{SLR algorithm for one-parameter exponential family of distributions.}\label{algo_slr_1d}
\end{algorithm}\DecMargin{1em}

## 2D-SLR Algorithm for Normal Distributions

As soon as we move away from one-parameter distributions parameterized by their mean, it becomes less clear how the likelihood ratio for $H_0: \mu = \tau$ and $H_1: \mu = \hat{\mu}$ should be calculated. In the case of Normal distributions with unknown mean and unknown variance, given that the variance and mean estimate are linked, we can use 
$$
H_0: \mu_0 = \tau \land \sigma_0^2 = \sigma^2_{\tau} = \frac{1}{n}\sum_{i=1}^{n}(x_i-\tau)^2 \quad \text{vs.} \quad H_1: \mu_1 = \hat{\mu} \land \sigma_1^2 = \hat{\sigma}^2 = \frac{1}{n}\sum_{i=1}^{n}(x_i-\hat{\mu})^2
$$

In the following, we assume $\epsilon = 0$ for simplicity. At each round $t$, the algorithm computes $T_i(t)$ and $\hat{\mu}_i(t)$ as before, and additionally the empirical variances $\sigma_{\tau,i}^2(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)}(x_{i,s}-\tau)^2$ and $\hat{\sigma}_{i}^2(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)}(x_{i,s}-\hat{\mu})^2$, where $\hat{\sigma}_{i}^2(t) < \sigma_{\tau,i}^2(t)$ if $\hat{\mu} \neq \tau$ since $\hat{\mu}$ is the maximum likelihood estimate. The index is then

\begin{align*}
B_i^{2DSLR}(t+1) & = -\log\big(LR_T(\tau, \sigma_{\tau,i}^2(t)||\hat{\mu}_i(t),\hat{\sigma}^2_{i}(t))\big) \\
& = T_i(t) \log\Big(\frac{\sigma_{\tau,i}^2(t)}{\hat{\sigma}_i^2(t)}\Big) \\
& = T_i(t) \KL\big(N(\hat{\mu}_i(t),\hat{\sigma}^2_{i}(t))||N(\tau, \sigma_{\tau,i}^2(t))\big) \Alignyesnumber\label{eqn:SLR2DIndex}
\end{align*}

The next arm to be pulled is the minimizer of the index $B_i^{2DSLR}(t+1)$, $I_{t+1} = \arg \min_{i \leq K} B_i(t+1)$. At the end of the horizon $T$, the algorithm outputs the set of arms $\hat{S}_\tau = \{i \in \mathbb{A}, \hat{\mu}_i(T) \geq \tau\}$.

\IncMargin{1em}
\begin{algorithm}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{\emph{$\tau$, $\epsilon$, $\nu(\mu, \theta) = \mathcal{N}(\mu,\sigma^2) \, \forall i \in \mathbb{A}$, where $\mu$ and $\sigma^2$ are unknown}}
\Output{$\hat{S}_\tau = \{i \in \mathbb{A},  \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm twice}\;
\For{$t = 2K+1$ \KwTo $T$}{
\For{$i = 1$ \KwTo $K$}{
  \emph{Compute $\hat{\sigma}^2_{i}(t)$}\;
  \If{$\hat{\sigma}^2_{i}(t) = 0$}{\emph{Set $\hat{\sigma}^2_{i}(t) = 1/T$}}
}
  \emph{Pull arm $I_t = \arg \min_{i\leq K} B_i^{SLR_{2D}}(t)$ as in \autoref{eqn:SLR2DIndex}}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{2D-SLR algorithm for Normal distributions with unknown mean and variance.}\label{algo_slr_2d}
\end{algorithm}\DecMargin{1em}

## Discussion

At the beginning of this thesis, we set out to find an algorithm that improves upon the Anytime Parameter Free algorithm introduced in Locatelli et al. (2016) for the thresholding bandit problem. We aimed to make the algorithm perform better in problems with arms that are not well approximated as sub-Gaussian; in particular a problem with Bernoulli arms of small means, and with a threshold close to 0. We thus tried to find a strategy that takes more information about the arms into account than only the sample mean. With the Simple Likelihood Ratio strategy, we propose an algorithm that fulfills this goal at the cost of being distribution specific.

The SLR strategy requires the knowledge of the class of distribution upfront, just as other algorithms require parameters. While this is likely to be a negligiable problem in the case of $\{0,1\}$ observations whose mean can be modeled reasonably by Bernoulli distributions, it can be more difficult to know the true class of distributions upfront when $\nu_i$ are continuous distributions. We have shown however, that the SLR strategy can be adapted to all one-parameter distributions in the exponential family of distributions. The algorithm even offers a natural adaptation to the problem of normal distributions with unknown mean and unknown variance.

Furthermore, the SLR strategy reduces to the APT strategy in the case where arms are normally distributed with known, equal variances. In this case we have as index (compare example 3 above): 
$$
B_i^{SLR}(t) = T_i(t)\frac{(\hat{\mu}_i(t) - \tau)^2}{2\sigma^2}.
$$
In order to decide at time $t$ which arm to draw next, we necessarily compare the indexes of arm $i$ and arm $j$. Assume w.l.o.g. that $B_i^{SLR}(t) < B_j^{SLR}(t)$. Then:

\begin{align*}
B_i^{SLR}(t) < B_j^{SLR}(t) & \Leftrightarrow 
T_i(t)\frac{(\hat{\mu}_i(t) - \tau)^2}{2\sigma^2} < T_j(t)\frac{(\hat{\mu}_j(t) - \tau)^2}{2\sigma^2} \Alignyesnumber\label{eqn:SLRWithKnownVar}\\
& \Leftrightarrow 
\sqrt{T_i(t)}\sqrt{(\hat{\mu}_i(t) - \tau)^2} < \sqrt{T_j(t)}\sqrt{(\hat{\mu}_j(t) - \tau)^2} \\
& \Leftrightarrow 
B_i^{APT}(t) < B_j^{APT}(t)
\end{align*}

Thus, SLR achieves the same upper bound as APT on this problem. Given that in all other cases the SLR is able to take more information about the distributions into account through the Kullback-Leibler divergence (in the case of the exponential family of distributions), the algorithm is expected to improve upon the APT strategy.

As mentioned before, this improvement comes only at the cost of knowing the correct distribution upfront. In particular, as for the APT algorithm, there are no parameters which need to be tuned upfront as is often the case for upper-confidence bound based algorithms (see for example UCB-E in Audibert et al., 2010).

Additionally, the index used in the SLR algorithm offers the same interpretation considered in \autoref{sec:EVTAlgorithm}. Write the index again as
$$
B_i^{SLR}(t) = S_i^{SLR}(t) D_i^{SLR}(t),
$$
where $S_i^{SLR}(t)$ is the confidence term and $D_i^{SLR}(t)$ is the measure of distance between the distribution of arm $i$ and the threshold $\tau$. The index of the Simple Likelihood Ratio strategy for one-parameter exponential family distributions can then be written with
$$
S_i^{SLR}(t) = T_i(t) \quad \text{and} \quad
D_i^{SLR}(t) = \KL(\hat{\mu}_i(t)||\tau),
$$
where we employ the number of pulls so far for a given arm as the confidence measure, while the Kullback-Leibler divergence fulfills its role as distance measure between distributions (see also Figure 2). The larger the Kullback-Leibler divergence, the less inclined we are to draw the arm; the same holds, if we have pulled an arm many times. Since the KL-divergence reduces to the squared distance between means for Normal distributions of known variance, the SLR and APT algorithms are identical in this case.

As we see in the experiments that follow, the SLR strategy is able to improve significantly upon the APT strategy in many cases.

\newpage

# Experiments \label{chap:Experiments}

We now compare the algorithms introduced in the previous chapters in simulations and experiments against each other to evaluate their performance. We first perform experiments on synthetic data in \autoref{sec:SyntheticData}. We then use Bernoulli time series data collected by the online shop Amorelie to evaluate the algorithms on real world data in \autoref{sec:RealData}.

We compare the empirical performance of the SLR algorithm not only against the APT algorithm, but also against the variance-based algorithms. Thus we also compare the variance-based algorithms for the first time directly against each other. As benchmark, the uniform sampling strategy is included. Additionally, we include the Bayes-UCB algorithm introduced in Kaufmann et al. (2012) for the cumulative regret setting. See \autoref{sec:AppendixBUCB} in the appendix for how we adapt the Bayes-UCB algorithm to the thresholding bandit problem.

## Synthetic Data \label{sec:SyntheticData}

Each of the following five simulations is repeated for 5000 iterations. Depending on the simulation, the observations are generated randomly by a given exponential family of distributions. As we are primarily concerned with improving upon APT with respect to non-sub-Gaussian distributions, the first three simulations focus on Bernoulli distributions, especially those with small means. We then go on to show with experiments on data from the Exponential distribution and Poisson distribution that the SLR strategy can be adapted to other distributions as well.

### Simulation 1: Bernoulli Arms and $\tau = 0.5$

The first experiment is equal to experiment 3 in Locatelli et al. (2016) to achieve a baseline comparison. We compare ten Bernoulli arms that move away from the threshold at $\tau = 0.5$ geometrically with $\mu_{1:4} = 0.4-0.2^{1:4}$, $\mu_5 = 0.45$, $\mu_6 = 0.55$, and $\mu_{7:10} = 0.6 + 0.2^{5-1:4}$. We run the standard APT algorithm as well as the SLR algorithm with $\epsilon = 0.1$. We compare them to the AugUCB algorithm with the parameter choice of $\rho = 1/3$ as used by its authors in Mukherjee et al. (2017). We run the parameter-free EVT$_{pf}$ algorithm without feedback delay, thus making it directly comparable to the other algorithms. The Bayes-UCB algorithm is run with $c=1$ and takes as prior a Beta distribution with parameters $\alpha_0 = \tau$ and $\beta_0 = 1 - \tau$, thus leading to a mean of the prior distribution at $\frac{\alpha_0}{\alpha_0+\beta_0}=\tau$ while keeping the prior as weak as possible. Lastly, we run the uniform sampling strategy as a benchmark.

```{r loc2_data_summary, echo = FALSE, message = FALSE}
get_apt_complexity <- function(x, tau, epsilon) {
  gaps <- abs(x-tau)+epsilon
  sum(1/gaps^2)
}

mean_loc2 <- c(0.4-0.2^(1:4),
               0.45, 0.55,
               0.6+0.2^(5-(1:4)))

H_loc2 <- get_apt_complexity(mean_loc2, 0.5, 0.1)

knitr::kable(t(mean_loc2), digits = 3, col.names = paste0("V", 1:10), caption = "Mean of arms in simulation 1.")
```

This first experiment is a problem in which all algorithms should be able to perform well. The problem is not very difficult with $H_1^{APT} \approx$ `r round(H_loc2)`, and due to $\tau = 0.5$, the Bernoulli distribution can be well approximated by a Gaussian. This is the perfect environment for the APT algorithm.

As the next figure shows, essentially all methods converge rapidly to an error rate of very close to zero, where the error rate is the share of the 5000 simulations on which the classification of the algorithm is wrong. Only the uniform strategy that clearly lacks behind. We also observe that the variance based algorithms (AugUCB and EVT) lag behind at first. AugUCB is able to adjust its sampling scheme only after 1000 iterations but then rapdily improves to the level of APT. EVT takes about 250 samples more to start a quick descend to the performance of the other algorithms. All of this, however, happens after the algorithms are below an error rate of 0.01. While the SLR algorithm improves upon APT, the improvement is a slight one, and not very large in terms of budget difference. This underlines that the scenario at hand is ideal for the APT algorithm.

```{r simulation1_results, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulation 1."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(paste0(current_path, "loc2_comp_BUCB_horizon_4000.Rda"))
load(paste0(current_path, "loc2_comp_APT_2000.Rda"))
load(paste0(current_path, "loc2_comp_LR_2000.Rda"))
load(paste0(current_path, "loc2_comp_AugUCB_2000.Rda"))
load(paste0(current_path, "loc2_comp_UNIFORM_2000.Rda"))
load(paste0(current_path, "loc2_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB","SLR"), 
                   each = 1991), rep("B-UCB", 3991), rep("EVT", 1981))
index <- c(rep(10:2000, times = 4), 10:4000, 20:2000)
alg_res <- log(c(loc2_comp_UNIFORM_2000, loc2_comp_APT_2000, 
                 loc2_comp_AugUCB_2000, loc2_comp_LR_2000,
                 loc2_comp_BUCB_horizon_4000, loc2_comp_EVT))
loc2_comp_res <- data.frame(index = index, names = alg_names, 
                            results = alg_res)
ggplot(loc2_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = names, color = names)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(subtitle = "Dashed Line indicates Error = 0.1.",
       y = "log(Error)", x = "Round Index") + 
  coord_cartesian(xlim = c(0,2000)) +
  scale_color_discrete(name = "Algorithm") + 
  theme_bw() #+
  #theme(legend.position = "bottom")

```

\newpage

### Simulation 2: Bernoulli Arms and $\tau = 0.05$ \label{sec:Simulation2}

```{r loc6nt_data_summary, echo = FALSE, message = FALSE}
mean_loc6nt <- c(0.001, 0.005, 0.01, 0.015,
                 0.04, 0.06,
                 0.085, 0.09, 0.095, 0.099)

H_loc6 <- get_apt_complexity(mean_loc6nt, 0.05, 0)

knitr::kable(t(mean_loc6nt), digits = 3, col.names = paste0("V", 1:10), caption = "Mean of arms in simulation 2 classified against threshold of 0.05.")
```

In the second experiment, we compare the same algorithms as in the first experiment. Again, we classify 10 arms, but this time against a threshold of $\tau = 0.05$ and without an epsilon interval, $\epsilon = 0$ (which generally rewards algorithms for pulling arms close to the threshold). We choose the means of the arms to be much smaller too, and arrange them symmetrically around the threshold. While the distributions are bounded and thus still sub-Gaussian, the Normal approximation should be poor in comparison to the previous case. We expect an improvement for SLR compared to APT as SLR takes the Bernoulli characteristics into account. In particular, even though the arms are symmetrically distributed around the threshold in terms of absolute distance, they are not symmetrically around the threshold in terms of their Kullback-Leibler divergence (compare Figure 2). Thus we expect the SLR algorithm to not pull them symmetrically. Furthermore, we expect an improvement for the variance-based algorithms, as the variances are now smaller than in the first simulation, which should benefit the EVT algorithm given its theoretical analysis in Zhong et al. (2017).

The figure below shows the performance of all algorithms. The results underline that this problem is much harder in general ($H_2^{APT} \approx 24703$). The performance of the variance based algorithms is surprisingly poor; EVT performs even worse than uniform sampling until close to the horizon, while Augmented UCB takes a long time to adjust away from uniform sampling. Both algorithms perform worse than APT after 7000 samples.

The Simple Likelihood Ratio strategy improves considerably upon the alternative strategies. SLR reaches the average empirical error of 0.1 within the budget, and it improves at an exponential rate within the entire budget; in contrast to APT which appears to improve linearly in the observations. At the end of the budget, the difference in the empirical error is nearly 10% between the two strategies.

In order to assess how the SLR strategy improves upon the APT algorithm, it is useful to look at how the algorithms allocate their sampling budget across arms. Given that we run the simulation 5000 times, we can compare across simulations how often an algorithm pulls each arm at a given round. During the 5000 iterations, the algorithm should pull at any given iteration the arm most often that it thinks is the most difficult to classify. In Figure 6, we plot the resulting curves where the y-axis shows a 10-round moving average of the pulls to remove some variation.

The graph shows that both APT and SLR correctly single out the two arms closest to the threshold, and allocate most of the budget to them. However, while SLR keeps the two arms constant at two different levels (about 1900 for arm 5 and about 2500 for arm 6), APT at first pulls arm 5 more often than arm 6. Then, for *both* arms, the budget appears to converge to about 2000. Given that the two arms are symmetric around the threshold, it makes sense that APT would allocate the same exploration effort to them. And given that the variation of arm 5 away from the threshold is limited from below by zero, it is much easier for arm 6 to have an empirical mean larger than its true mean, and thus to have a larger index than arm 5 in the beginning. Consequently, APT would prefer to sample arm 5. This, however, is wrong in terms of Kullback-Leibler divergence. Measured by the divergence, arm 6 is closer to the threshold than arm 5, thus harder to classify, and consequently preferred by the SLR algorithm. Furthermore, the SLR algorithm appears to quickly find the correct share that it has to allocate to the two arms. This is in contrast to APT which continues to adjust the shares across the entire budget.

Additionally, we can observe that the SLR strategy starts to drop arms much quicker than APT, and it drops arm 1 to 3 first, which again underlines the effect of the asymmetric Kullback-Leibler divergence. If we consider arm 3 and arm 8, we have $d_3(0.01||0.05) \approx 0.025$ while $d_8(0.09||0.05) = 0.014$. The divergence is much smaller for arm 8. The asymmetry is even more dramatic for arm 1 and arm 10 with $d_1(0.001||0.05) \approx 0.046$ and $d_{10}(0.099||0.05) \approx 0.020$: In terms of the divergence, the distance between arm 1 and the threshold is twice as far as between arm 10 and the threshold. APT considers them to be equally distant from the threshold.

One interesting observation, that we will observe in all simulations with Bernoulli arms, is the superior performance of the Bayes-UCB algorithm adapted from Kaufmann et al. (2012) which was originally designed for cumulative regret problems. Nonetheless, the algorithm proves again and again to be the best performing. Compare with \autoref{sec:AppendixBUCB} for more analysis on the behavior of the Bayes-UCB algorithm in simulation 2. In particular, consider Figures 17 and 18 which show for individual arms the distribution of pulls for that arm across simulations. The distributions highlight that both the APT and SLR fail in many iterations of the simulations to pull arms frequently enough to estimate their correct mean. The APT algorithm pulls the most difficult arm (Arm 6) in more than 5% of the simulations only once (during initialization). Consider arm 6 returns $X_{6,1} = 1$ as its first observation. Then $B_6(t+1) = 1 \cdot |1-0.05| = 0.95$. It will then not be pulled for a long time because its index is very large. Indeed, consider some arm that has been pulled 4000 times with $\mu_i(4000) = 0.03$. Then $B_i(t+1) = \sqrt{2000} \cdot |0.05-0.03| \approx 0.89$. Even though we have basically no evidence for the performance of arm $6$, we would continue to pull arm $i$. Similar holds for the SLR algorithm. For both, the distribution of pulls for arm 6 is very flat across the interval of $[1,6000]$ pulls. In contrast, the distribution of pulls appears normally distributed for the Bayes-UCB algorithm with the mode between 2100 and 2300 pulls. The numerical problem that arises for the APT and SLR algorithms might be weakened for the Bayes-UCB algorithm by the combination of a (weak) prior that shrinks the mean estimates, and the way in which quantiles are based on the posterior. Also, the fact alone that it is based on confidence bounds of the empirical mean, instead of the empirical mean itself. We show in the appendix that the problem does not only occur for arm 6.

Empirically, part of this problem can be alleviated by including the D-Tracking rule proposed by Garivier and Kaufmann (2016) and discussed in \autoref{sec:OptimalBestArmIdentification} into the SLR algorithm, as it ensures that no arm is ``left behind'' in terms of samples. Figure 5 shows a significant improvement for the SLR-D algorithm in comparison to the SLR algorithm without D-Tracking. However, the algorithm still does not reach the performance of Bayes-UCB.

Lastly, we remark on the shape of the error curve of the uniform strategy. Even though the uniform strategy samples arms at random from the set of arms with the least observations at any round, we observe a sawtooth pattern for simulation 2 and 3. This is an artifact of the comparison against the threshold of $0.05$ and $0.01$ respectively, combined with the discrete $0$ and $1$ observations. Take for example an arm with a true mean of $\mu = 0.06$ compared against $\tau = 0.05$. Let's say the arm observes the first 1 after 15 samples, $X_{15} = 1$, such that $\hat{\mu}(15) = 1/15 \approx 0.067$. At this point, it is correctly classified as above the arm. Even if the next samples are zeroes, it will continue to be correctly classified until observation 20. At that point, $\hat{\mu}(20)$ = 1/20 = 0.05. By observing one more observation, the arm will be classified as below the threshold, since then $\hat{\mu}(21) = 1/21 < 0.05 = \tau$. The classification is suddenly wrong for no particular reason, and a sudden spike in the error curve occurs. In the graph of simulation 3 it is obvious that the spike occurs with a period length of about 100 samples, which corresponds to $\tau = 0.01=1/100$. This explains why we do not observe such a pattern in the first simulation with $\tau = 0.5$, or for the continuous Exponential distribution in simulation 4.

```{r simulation2_results, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulations 2 and 3. Note the different x-axes."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(paste0(current_path, "loc6nt_comp_BUCB_horizon_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_AugUCB_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_UNIFORM_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_APT_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_LR_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_LRD.Rda"))
load(paste0(current_path, "loc6nt_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB",
                     "SLR", "SLR-D", "B-UCB"), 
                   each = 6991), rep("EVT", 6981))
index <- c(rep(10:7000, times = 6), 20:7000)
alg_res <- log(c(loc6nt_comp_UNIFORM_7000, loc6nt_comp_APT_7000,
                 loc6nt_comp_AugUCB_7000, loc6nt_comp_LR_7000,
                 loc6nt_comp_LRD,
                 loc6nt_comp_BUCB_horizon_7000, loc6nt_comp_EVT))
loc6nt_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)

current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(file = paste0(current_path, "loc7nt_comp_UNIFORM_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_APT_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_LR_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_LRD_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_BUCB_horizon_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_AugUCB_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB", 
                     "SLR", "SLR-D", "B-UCB"), 
                   each = 9990), rep("EVT", times = 9979))
index <- c(rep(11:10000, times = 6), 22:10000)
alg_res <- log(c(loc7nt_comp_UNIFORM_10000, loc7nt_comp_APT_10000,
                 loc7nt_comp_AugUCB_10000, loc7nt_comp_LR_10000,
                 loc7nt_comp_LRD_10000,
                 loc7nt_comp_BUCB_horizon_10000, loc7nt_comp_EVT))
loc7nt_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)

sim23_res <- rbind.data.frame(loc6nt_comp_res, loc7nt_comp_res)
sim23_res$Simulation <- c(rep("Simulation 2", 48927), rep("Simulation 3", 69919))

ggplot(sim23_res, aes(x = index, y = results)) + 
  geom_line(aes(group = Algorithm, color = Algorithm)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  facet_grid(.~Simulation, scales = "free") +
  labs(subtitle = "Dashed Line indicates Error = 0.1.",
       y = "log(Error)", x = "Round Index") +
  scale_color_discrete(name = "Algorithm") +
  theme_bw() + 
  theme(legend.position = "bottom")
```

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '70%', fig.cap = "Budget allocation across arms for APT and SLR algorithms in simulation 2."}
load(file = "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/Thesis/experiment_postprocessing/loc6nt_ma.Rda")
library(dplyr)

rbind(loc6nt_APT_ma, loc6nt_LR_ma) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.3) +
  facet_grid(.~algorithm) +
  labs(x = "Index", y = "10 Rounds Moving Average"#,
       #title = "Average Number of Pulls Allocated to Each Arm per Round",
       #subtitle = "5000 iterations; a value of 1000 means that in 20% of 
#the iterations the arm was pulled in that round."
) +
  theme_bw() +
  theme(legend.position = "bottom")
```

\newpage

### Simulation 3: Bernoulli Arms and $\tau = 0.01$

In the third experiment, we lower the threshold further to $\tau = 0.01$ and arrange now $K=11$ arms on a $\log_{10}$ scale around it. Again, we set $\epsilon = 0$. We compare the same algorithms as previously with the settings adapted to $\tau$ and $\epsilon$ accordingly. Notice that the experiment horizon is now $T=10000$ rounds instead of $T=7000$ rounds in the previous experiment.

```{r loc7nt_data_summary, echo = FALSE, message = FALSE}
mean_loc7nt <- c(10^-4, 10^-3.5, 10^-3.25, 10^-3, 
                 10^-2.75, 10^-2.5, 10^-2.25, 10^-1.85, 10^-1.75,
                 10^-1.5, 10^-1)

mean_loc7nt_char <- c("10^-4", "10^-3.5", "10^-3.25", "10^-3", "10^-2.75", "10^-2.5", "10^-2.25", "10^-1.85", "10^-1.75", "10^-1.5", "10^-1")

#get_apt_complexity(mean_loc7nt, 0.01, 0)

knitr::kable(t(mean_loc7nt), digits = 4, col.names = paste0("V", 1:11), caption = "Mean of arms in simulation 3.")
```

Looking at the performance after 7000 rounds, we see that the problem appears to be slightly more difficult compared to the previous one ($H_3^{APT} \approx 210359$). We see that up until this point the variance-based strategies cannot improve upon APT. Also, SLR has not yet crossed the 10% empirical error. We observe that in the following rounds EVT decreases at an exponential rate and thus crosses the error of APT and continues to improve. At the end of the budget, the improvement of SLR compared to the other algorithms is again large, and likely to increase. However, Bayes-UCB is again the best performing algorithm and is improving at a faster rate than SLR. As in the previous experiment, Bayes-UCB and SLR show significant improvements on this Bernoulli problem with very small means and threshold compared to APT in particular. Furthermore, the observation that the SLR algorithm with D-Tracking rule performs better than without is replicated.

```{r simulation3_results, echo = FALSE, include = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulation 3."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(file = paste0(current_path, "loc7nt_comp_UNIFORM_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_APT_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_LR_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_LRD_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_BUCB_horizon_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_AugUCB_10000.Rda"))
load(file = paste0(current_path, "loc7nt_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB", "SLR", "SLR-D", "B-UCB"), 
                   each = 9990), rep("EVT", times = 9979))
index <- c(rep(11:10000, times = 6), 22:10000)
alg_res <- log(c(loc7nt_comp_UNIFORM_10000, loc7nt_comp_APT_10000,
                 loc7nt_comp_AugUCB_10000, loc7nt_comp_LR_10000,
                 loc7nt_comp_LRD_10000,
                 loc7nt_comp_BUCB_horizon_10000, loc7nt_comp_EVT))
loc7nt_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)
ggplot(loc7nt_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = Algorithm, color = Algorithm)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(subtitle = "Dashed Line indicates Error = 0.1.", 
       y = "log(Error)", x = "Round Index") +
  theme_bw() #+
  #theme(legend.position = "bottom")
```

### Simulations 4 and 5: Exponential and Poisson Distributions

In \autoref{sec:LRforUnivariateExponentialFamily}, we show that the likelihood ratio has a very useful form for the one-parameter exponential family of distributions, which allows us to use the SLR algorithm not only on Bernoulli arms, but also on Exponential or Poisson distributions, for example. In simulation 4, we consider Exponential distributions, in simulation 5, we test the algorithms on Poisson distributions.

We run the same algorithms as before except for Bayes-UCB (even though it should certainly be possible to adapt it to these case based on a Gamma prior). In the case of the Exponential arms, we classify against $\tau = 1$ and have sampled the arms' mean parameters from an Exponential distribution with rate $\theta = 1$. The Poisson arms are classified against a threshold of $\tau = 0.4$, with the mean parameters being sampled from an Exponential distribution with mean $\mu = 0.7$. In both cases, we classify 20 arms.

For the Exponential arms, APT crosses the 10% error mark first, followed by the AugUCB algorithm. All algorithms perform better than uniform sampling. After more than half of the budget, the SLR algorithm jumps in performance and performs better than APT and AugUCB. Indeed, the AugUCB basically stops improving after half of the budget. This is a consequence of it eliminating arms it believes to know very well. This happens at an error rate of less than $0.01$. On average, the AugUCB samples only for 2564 rounds of the budget of 4000 rounds across the 5000 simulations.

The early elimination (when compared to the Bernoulli experiments) might be caused by the fact that the AugUCB assumes the variance of the arms to be in $[0,1]$, which does not hold for half of the arms in the experiment. Similar goes for the EVT algorithm which is only able to catch up at the very end of the budget.

```{r simulation4_results, include = FALSE, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulation 4."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/exponential/"

load(paste0(current_path, "exp2_comp_LR.Rda"))
load(paste0(current_path, "exp2_comp_APT.Rda"))
load(paste0(current_path, "exp2_comp_EVT.Rda"))
load(paste0(current_path, "exp2_comp_UNIFORM.Rda"))
load(paste0(current_path, "exp2_comp_AugUCB.Rda"))

alg_names <- c(rep(c("Uniform", "APT","SLR", "AugUCB"), 
                   each = 3981), rep("EVT", 3961))
index <- c(rep(20:4000, times = 4), 40:4000)
alg_res <- log(c(exp2_comp_UNIFORM, exp2_comp_APT, exp2_comp_LR,
                 exp2_comp_AugUCB, exp2_comp_EVT))
exp2_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)
ggplot(exp2_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = Algorithm, color = Algorithm)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(subtitle = "Dashed Line indicates Error = 0.1.",
       y = "log(Error)", x = "Round Index") +
  theme_bw()# +
  #theme(legend.position = "bottom")
```

In the case of Poisson arms, the performance of the arms is more clear cut. As expected, SLR improves upon APT. APT outperforms the EVT algorithm, which occurs even though the variance of all arms considered in this experiment is in $[0,1]$. After about 1000 observations, APT and EVT appear to improve at about the same rate, but APT has a much better initial exploration phase. Given the previous simulations, it is surprising to see the Augmented-UCB algorithm to come ahead after $T=10000$ observations. After about half the budget, the same pattern as before occurs where there is a drastic change in sampling behavior. This suffices to outpace both APT and SLR.

```{r simulation5_results, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulations 4 and 5. Note the different x-axes."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/exponential/"

load(paste0(current_path, "exp2_comp_LR.Rda"))
load(paste0(current_path, "exp2_comp_APT.Rda"))
load(paste0(current_path, "exp2_comp_EVT.Rda"))
load(paste0(current_path, "exp2_comp_UNIFORM.Rda"))
load(paste0(current_path, "exp2_comp_AugUCB.Rda"))

alg_names <- c(rep(c("Uniform", "APT","SLR", "AugUCB"), 
                   each = 3981), rep("EVT", 3961))
index <- c(rep(20:4000, times = 4), 40:4000)
alg_res <- log(c(exp2_comp_UNIFORM, exp2_comp_APT, exp2_comp_LR,
                 exp2_comp_AugUCB, exp2_comp_EVT))
exp2_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)

current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/poisson/"

load(paste0(current_path, "pois2_comp_LR.Rda"))
load(paste0(current_path, "pois2_comp_APT.Rda"))
load(paste0(current_path, "pois2_comp_EVT.Rda"))
load(paste0(current_path, "pois2_comp_UNIFORM.Rda"))
load(paste0(current_path, "pois2_comp_AugUCB.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "SLR", "AugUCB"), each = 9981), 
               rep("EVT", 9961))
index <- c(rep(20:10000, times = 4), 40:10000)
alg_res <- log(c(pois2_comp_UNIFORM, pois2_comp_APT, 
                 pois2_comp_LR, pois2_comp_AugUCB, pois2_comp_EVT))
pois2_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)

exp45_res <- rbind.data.frame(exp2_comp_res, pois2_comp_res)
exp45_res$Distribution <- c(rep("Exponential", 19885), rep("Poisson", 49885))

ggplot(exp45_res, aes(x = index, y = results)) + 
  geom_line(aes(group = Algorithm, color = Algorithm)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  facet_grid(.~Distribution, scales = "free") +
  labs(#title = "Results of Experiment 5", 
       subtitle = "Dashed line at Error = 0.1.", 
       y = "log(Error)", x = "Round Index") +
  theme_bw() +
  theme(legend.position = "bottom")
```

## Experiments on Web Analytics Data \label{sec:RealData}

In the following, we present the results of two experiments performed on a real world data set. While most multi-armed bandit algorithms are assessed on synthetic data, as we have done in \autoref{sec:SyntheticData}, a real world data set is useful to observe how algorithms perform in presence of, for example, not perfectly i.i.d. samples. Every company of course tries to run their algorithms under settings as ideal as possible, but there will always be problems. Thus our experiments serve as a first robustness check on the algorithms introduced before.

### The Data Set

Data have been collected by the German online shop [Amorelie](https://amorelie.de) as part of standard website analytics. For 197 selected products, we observe the number of users visiting each product's detail page (PDP) during a given minute. The data set we use here has been collected during the entire month of November 2016. It can be considered time series data collected at minutely frequency. Thus the data contain 43200 observations per product, and overall 8510400 observations.

Since we want to run an offline comparison of bandit algorithms, it is important to have at each time point an observation for each arm. Given that, it is easy to evaluate the perfomance of algorithms that make different decisions at different point in times and thus do not observe the same samples. By having an observation for each arm at any given time point, each algorithm had at least the opportunity to observe a certain sample. Other papers (for example, Li et al., 2011) have solved this problem by using a sample drawn uniformly during, for example, an A/B test. Then, if the next observation in the sample is not from the same arm that the algorithm would like to draw from, observations are discarded until the next observation from the selected arm.

The choice of using what is basically multiple time series data as the sample offers a simple way of running the algorithms just as often on the data as for simulated data. The goal is to perform 5000 runs over the sample to confidently assess the performance. Since we would like to preserve potential time dependencies in the data, a simple resampling of the data is not an option to create the needed 5000 variations of the sample. Instead we perform time series cross validation. To preserve the autocorrelation, one can create different samples of $X$ observations by sliding a window of size $X$ over the time series. Moving the window one period (in our case one minute) at a time, one ensures that no sample shares start and end observation with another sample. In particular, this means for a given bandit strategy and two samples that the observations that the algorithm collects as feedback during the initialization phases differ from sample to sample and are never the same. This is because during initialization every strategy pulls arms $1, ..., K$ (or $1,...,2K$ for EVT) once in this order.

```{r, echo = FALSE}
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/arms_example_binary.Rda")

knitr::kable(arms_example_binary, caption = "Example of the binary time series data for five products.")
```

While the data set consists of count data (the number of visits to a PDP during a given minute), we binarize the data by applying a certain threshold. If the count is above this threshold, the new observation is $1$, else $0$. Thus we have a binary time series of length $60\cdot 24 \cdot 30 = 43200$ for each of the 197 products. Given that the number of visitors that reach a given PDP is likely to vary by time of day and by day, and in general over time, we expect the mean of the time series to not be stable over time. For example, less visitors visit during night than during the evening. This leads to a decrease in the average number of visits, and creates a daily seasonality. On the other hand, more users may visit the site during a weekend, thus leading to a spike on Saturdays and Sundays. Because of the latter time dependency, it is best practice to run online A/B tests for at least a week^[See for example the "A/B Testing in the Wild" presentation by Emily Robinson, <https://youtu.be/SF-ryGgLOgQ>], or an integer multiple of weeks, so as to take into account the weekly seasonality in the time series. For this reason, the experiments we present below are run on $60 \cdot 24 \cdot 7 = 10080$ observations for each arm, so as to have a sample for one full week.

We can look at summary statistics for the sample means of the 197 products over the month of November to get a first impression of the binary data we are dealing with. We see that most products have a small mean, which is very common also for actual conversion rate or click-through rate data that can be tested with A/B tests or multi-armed bandits. Yet the sample of products contains a few outliers with very large view rates, with a maximum rate of 78%. If we used a threshold $\gamma$ to binarize the count data, then this means that on average every fourth minute the count of visitors was below $\gamma$ for this product. On the other hand, we conclude that using thresholding bandit strategies to test the products against a threshold of for example $\tau = \frac{2}{60} = 0.0333$ or $\tau = \frac{5}{60} = 0.083$ is likely to create more complex problems than a threshold of, say, $\tau = \frac{15}{60} = 0.25$ would.

```{r amo_data_summary, echo = FALSE, message = FALSE}
load(file = "/Users/timradtke/Desktop/thesis_data/pv_summary.Rda")

knitr::kable(pv_summary, digits = 3, col.names = c("Mean", "Minimum", "25% Quantile", "Median", "75% Quantile", "Maximum"), caption = "Summary Statistics for View Rates of 197 Products")
```

### Experiment 1

In the first experiment, we pick ten arms randomly from the set of 197 arms. The experiment is run with a budget of $T=10080$ samples for each of the $5000$ iterations for which we repeat the experiment. Given the time series cross-validation idea described above, this means that in total, $15080$ observations of the $43200$ available observations per arm are used in training. In the thresholding bandit problem, we classify arms as below or above a threshold. Thus we can evaluate the strategies also based on whether their classifications are correct in a holdout sample (the time period after training). As this holdout sample, we can use the 10080 observations in our data set that follow the first 15080 observations used in training. In the table below, we present the count rates of the ten products for both training and test samples.

```{r experiment1_arm_means, echo = FALSE, message = FALSE}

load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo3_mean_firsthalf.Rda")
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo3_mean_secondhalf.Rda")

Training <- data_amo3_mean_firsthalf
Test <- data_amo3_mean_secondhalf
tau <- 2/60
TrainingMinusTau <- Training - tau

#knitr::kable(rbind(Training, Test, TrainingMinusTau)[,order(Training)], digits = 3, col.names = names(data_amo3_mean_firsthalf)[order(Training)],
#             caption = "Count rates for ten products for both training and test sample.")
knitr::kable(rbind(Training, Test)[,order(Training)], digits = 3, col.names = names(data_amo3_mean_firsthalf)[order(Training)],
             caption = "Training and test sample means for experiment 1.")

```

We see that while the rates do vary across training and test samples, they are relatively stable which is likely due to the fact that the samples come from an entire week's worth of observations. In particular, since we test against a threshold of $\tau = \frac{2}{60} = 0.0333$ in this experiment, the true classifications of arms do not differ across training and test data. A correct classification after training will be correct in the test set. For the rest of the analysis, we thus evaluate the performance using the training rates.

```{r, include = FALSE}
#When comparing the rates against the threshold, we see that three products are below the already small threshold. Product $V64$ seems to be especially difficult to classify correctly, even without time dependency. On the other hand, arms $V119$ and $V118$ should be much easier than the rest.
```

In order to visualize the time dependency that we expect for our samples, we plot a moving average of the mean of the time series for the first week of data. To remove daily seasonality, we select a sliding window of $60\cdot24 = 1440$ observations. To make the time dependency even more obvious, we contrast the actual data against synthetic data generated using the sample means of the actual data.

The graph again underlines that arm $V64$ is likely to be very difficult to classify given that its moving average crosses the threshold twice. The same is true for arm $V139$, which is not visible when looking at the overall mean alone. Furthermore, arms $V159$ and $V189$ come very close to the threshold at times.

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Moving average for the arms used in experiment 1."}
library(zoo)
library(tidyr)
library(dplyr)
library(ggplot2)
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo3_wide.Rda")

real_daily <- data.frame(index = 1:8641,
                                lapply(data_amo3_wide[1:10080,], FUN = rollapply, 
                                       width = 24*60, mean)) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 86410)) %>%
  rename(Arm = arm)

real_means <- colMeans(data_amo3_wide[1:10080,])
tau_amo3 <- 2/60

syn_data <- data.frame(rep(NA, times = 10080))
set.seed(124632728)
for(i in 1:length(real_means)) {
  syn_data[[i]] <- as.numeric(purrr::rbernoulli(10080, p  = real_means[i]))
}
names(syn_data) <- names(data_amo3_wide)

syn_daily <- data.frame(source = rep("Synthetic", 8641),
                        index = 1:8641,
                        lapply(syn_data, FUN = zoo::rollapply, width = 24*60, mean)) %>%
  gather(arm, mean, -index, -source) %>%
  rename(Arm = arm)

rbind(real_daily, syn_daily) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo3), linetype = 2) +
  facet_grid(.~source) + 
  labs(#title = "Moving average over 24*60 observations (1 day)", 
       subtitle = "Dashed line indicates the threshold.", 
       x = "Index", y = "One day moving average") +
  theme_bw() +
  theme(legend.position = "bottom")
```

These difficulties seem to have a large impact on the performance of all algorithms. Given that the data are binary observations, we can compare all algorithms in a setting in which one expects all assumptions to hold before having observed the data: The distributions are sub-Gaussian for the APT algorithm, bounded between $0$ and $1$ for the variance-based algorithms, and assumed to be Bernoulli for the SLR algorithm. Additionally, we use the Bayes-UCB algorithm as it performed best on the synthetic data. We set $\epsilon = 0$, and $\tau = 2/60$. For the AugUCB algorithm we use the parameters as in Mukherjee et al. (2017), that is, $\rho = 1/3$. For Bayes-UCB we set the prior to $\alpha_0 = \tau$ and $\beta_0 = 1-\tau$, and pick $c=1$.

```{r, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for experiment 1."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/"

load(paste0(current_path, "amo3_comp_BUCB.Rda"))
load(paste0(current_path, "amo3_comp_APT.Rda"))
load(paste0(current_path, "amo3_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo3_comp_AugUCB.Rda"))
load(paste0(current_path, "amo3_comp_LR.Rda"))
load(paste0(current_path, "amo3_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB","SLR","B-UCB"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 5), 20:10080)
alg_res <- log(c(amo3_comp_UNIFORM, amo3_comp_APT, amo3_comp_AugUCB, amo3_comp_LR, amo3_comp_BUCB, amo3_comp_EVT))
loc3_comp_res <- data.frame(index = index, Algorithm = alg_names, 
                            results = alg_res)
ggplot(loc3_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = Algorithm, color = Algorithm)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(#title = "Results of Real Data Experiment 1", 
       subtitle = paste0("Dashed Line indicates Error = 0.1."), 
       y = "log(Error)", x = "Round Index") +
  theme_bw()
```

Compared to the previous simulations with synthetic data, the results are dramatically poor. First of all, we see that all strategies perform poorly. All of them end above the 10% error rate mark after 1 week of data. Second, the uniform sampling strategy fares best when evaluated against the sample mean of all observations used during training. This of course is the opposite of what we would like to see, and the opposite of what we saw on the synthetic data, where we were able to improve upon the naive strategy. Third, again in contrast to the results on synthetic data, we observe non-monotonicity in the error curves. Given that the highs and lows for the different models run mostly parallel, it stands to reason that this effect is caused by the variation in means of the arms over time. Indeed, roughly comparing the error curves to the moving averages in the previous figure, the increase in errors after round $5000$ and the decreasing error after round $7500$ seems to line up with the fluctuations observed in the data. Given that we evaluate the models against a sample mean of all observations, it is not surprising that the uniform sampling strategy has a certain advantage when it comes to "factoring out" the seasonalities. In contrast, the adaptive sampling schemes try to adapt more or less quickly to the changes in the means--a feature that would be valuable in a cumulative regret multi-armed bandit setting. Here, however, it leads to wrong estimates of the overall sample mean.

If we focus on a comparison of adaptive sampling strategies and less on their globally bad performance, we see that the variance based strategies perform worse than APT and SLR. Very dramatic is the behavior of the Augmented-UCB algorithm, which first seems to perform well; this, however, is a result of it pulling arms mostly uniformly, until it starts to adapt its sampling and discarding arms after nearly 5000 rounds. The result is a dramatic increase in error which leads to the worst performance at the budget horizon. The differences between APT, SLR, and EVT are small, and their performance fluctuates in sync. None of them is really able to converge due to the changes in the underlying means. Advantageous only seems to be the fact that their behavior is less erratic than that of the AugUCB strategy given that they do not discard arms.

### Experiment 2

In order to underline the results of the previous experiment, we perform a second experiment using the data collected by Amorelie. We increase the threshold to $\tau = \frac{5}{60} = 0.833$ to make the problem in general slightly easier, and again pick 10 arms at random from the data set of 197 arms. We again use a sliding window of $10080$ observations to create $5000$ iterations and look at the training and test sample means as in the previous experiment.

Some arms, in particular $V110$ and $V150$, but also $V103$, $V149$, $V104$, and $V112$ are quite close to the threshold. Two arms, $V129$ and $V32$ appear to have basically no positive observation in the training set. On the other hand, arm $V154$ is larger than the threshold by a fair amount and should be easier to classify.

```{r experiment2_arm_means, echo = FALSE, message = FALSE}

load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_mean_firsthalf.Rda")
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_mean_secondhalf.Rda")

Training <- data_amo4_mean_firsthalf
Test <- data_amo4_mean_secondhalf
tau <- 5/60
TrainingMinusTau <- Training - tau

#knitr::kable(rbind(Training, Test, TrainingMinusTau)[,order(Training)], #digits = 3, col.names = names(data_amo4_mean_firsthalf)[order(Training)],
#             caption = "Count rates for ten products for both training #and test sample.")

knitr::kable(rbind(Training, Test)[,order(Training)], digits = 3, col.names = names(data_amo4_mean_firsthalf)[order(Training)],
             caption = "Training and test sample means for experiment 2.")

```

As in the previous experiment, we can also take a look at the one day moving averages of the arms to quickly assess the time dependency of the means. Given those and the previous experiment, the performance results of Experiment 2 shown in Figure 13 are not surprising. We recognize the same structure as previously. The uniform sampling strategy performs the best. The adaptive sampling strategies perform bad, do not converge, and end above the 10% error rate mark. This time, AugUCB is not (yet) the worst performing strategy, but equally erratic after it eventually adapts its sampling. This time around, it even has two spikes in its adaptive phase. Again, APT, SLR, and EVT perform in sync and more or less equally bad, with EVT slightly better than APT and SLR this time around. Interestingly, Bayes-UCB performs clearly worst of all strategies. Its performance changes at the same time as AugUCB starts to adapt its sampling strategy. This last fact is of particular interest, as Bayes-UCB is dramatically better than the other strategies in experiments on synthetic data.

On the right hand side of figure 13, the performance of the algorithms on synthetic data from Bernoulli distributions is depicted. These synthetic samples were sampled from distributions with mean equal to the sample means of the real data. We thereby create similar samples that are perfectly i.i.d. On this data, Bayes-UCB performs very well, ending at an average error rate of about 0.1%. This mirrors its performance observed in the previous simulations with synthetic data. Also all other adaptive strategies perform clearly better now, and their error rates are strictly decreasing. Interestingly, the uniform sampling scheme is now clearly worse than in the experiment on real data; it does no longer have the advantage of sampling the different mean phases at a representative rate. Excluding Bayes-UCB, we see that the SLR algorithm performs the best, followed by the variance-based algorithms. APT comes in at the same error rate as uniform sampling. While the difference between SLR and EVT is less than 2% (1.2% vs 3%), the difference to uniform sampling (10%) is relevant. The error rate of 10% is achieved by SLR about 1000 observations earlier than by EVT.

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Moving average of arms used in experiment 2. Dashed line indicates the threshold."}
library(zoo)
library(tidyr)
library(dplyr)
library(ggplot2)
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_wide.Rda")

real_daily <- data.frame(index = 1:8641,
                                lapply(data_amo4_wide[1:10080,], 
                                       FUN = rollapply, 
                                       width = 24*60, mean)) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 86410))

real_means <- colMeans(data_amo4_wide[1:10080,])
tau_amo4 <- 5/60

syn_data <- data.frame(rep(NA, times = 10080))
set.seed(124632728)
for(i in 1:length(real_means)) {
  syn_data[[i]] <- as.numeric(purrr::rbernoulli(10080, p  = real_means[i]))
}
names(syn_data) <- names(data_amo4_wide)

syn_daily <- data.frame(source = rep("Synthetic", 8641),
                        index = 1:8641,
                        lapply(syn_data, FUN = zoo::rollapply, width = 24*60, mean)) %>%
  gather(arm, mean, -index, -source)

rbind(real_daily, syn_daily) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo4), linetype = 2) +
  facet_grid(.~source) + 
  labs(#title = "Moving average over 24*60 observations (1 day)", 
       subtitle = "Horizontal bar indicates the threshold.", 
       x = "Index", y = "One day moving average") +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r, include = FALSE, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Rolling average of arms used in experiment 2."}
amo4_cummean_real <- data_amo4_wide[1:10080,] %>% tbl_df() %>%
  mutate_each(funs(cummean)) %>%
  mutate(index = 1:10080) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 100800))

amo4_cummean_syn <- syn_data %>% tbl_df() %>%
  mutate_each(funs(cummean)) %>%
  mutate(index = 1:10080) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Synthetic", 100800))

rbind(amo4_cummean_real, amo4_cummean_syn) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo4), linetype = 2) +
  facet_grid(.~source) + 
  labs(#title = "Running average of 10080 observations (1 week)", 
       subtitle = "Horizontal bar indicates the threshold.", 
       x = "Index", y = "Running average") +
  ylim(c(0,0.3)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r, include = FALSE, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Rolling average of arms used in experiment 2."}
amo4_cummean_real_12hrs <- data_amo4_wide[721:10800,] %>% 
  tbl_df() %>%
  mutate_each(funs(cummean)) %>%
  mutate(index = 1:10080) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 100800))

rbind(amo4_cummean_real_12hrs, amo4_cummean_syn) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo4), linetype = 2) +
  facet_grid(.~source) + 
  labs(title = "Running average of 10080 observations (1 week), 12hrs later", 
       subtitle = "Horizontal bar indicates the threshold.", 
       x = "Index", y = "Running average") +
  ylim(c(0,0.3)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r, include = FALSE, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Rolling average of arms used in experiment 2."}
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_shuffled_wide.Rda")

set.seed(214237)
shuffle_idx <- sample(1:(24*60), 24*60)
data_amo4_shuffled_wide[1:1440,] <- data_amo4_shuffled_wide[shuffle_idx,]

amo4_shuffled_cummean_real <- data_amo4_shuffled_wide[1:10080,] %>%
  tbl_df() %>%
  mutate_each(funs(cummean)) %>%
  mutate(index = 1:10080) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 100800))

rbind(amo4_shuffled_cummean_real, amo4_cummean_syn) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo4), linetype = 2) +
  facet_grid(.~source) + 
  labs(#title = "Running average of 10080 observations (1 week), 12hrs later", 
       subtitle = "Horizontal bar indicates the threshold.", 
       x = "Index", y = "Running average") +
  ylim(c(0,0.3)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance results of algorithms in experiment 2."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/"

load(paste0(current_path, "amo4_comp_BUCB.Rda"))
load(paste0(current_path, "amo4_comp_APT.Rda"))
load(paste0(current_path, "amo4_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo4_comp_AugUCB.Rda"))
load(paste0(current_path, "amo4_comp_LR.Rda"))
load(paste0(current_path, "amo4_comp_EVT.Rda"))

load(paste0(current_path, "amo4sim_comp_BUCB.Rda"))
load(paste0(current_path, "amo4sim_comp_APT.Rda"))
load(paste0(current_path, "amo4sim_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo4sim_comp_AugUCB.Rda"))
load(paste0(current_path, "amo4sim_comp_LR.Rda"))
load(paste0(current_path, "amo4sim_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB","SLR", "B-UCB"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 5), 20:10080)
alg_res <- log(c(amo4_comp_UNIFORM, amo4_comp_APT, amo4_comp_AugUCB, amo4_comp_LR, amo4_comp_BUCB, amo4_comp_EVT))
source_name <- rep("Real", 60416)
loc4_comp_res <- data.frame(source = source_name, 
                            index = index, names = alg_names, 
                            results = alg_res)

alg_names <- c(rep(c("Uniform", "APT", "AugUCB", "SLR", "B-UCB"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 5), 20:10080)
alg_res <- log(c(amo4sim_comp_UNIFORM, amo4sim_comp_APT, amo4sim_comp_AugUCB,
                 amo4sim_comp_LR, amo4sim_comp_BUCB, amo4sim_comp_EVT))
source_name <- rep("Synthetic", 60416)
loc4sim_comp_res <- data.frame(source = source_name, index = index, 
                               names = alg_names, results = alg_res)
loc4total <- rbind(loc4_comp_res, loc4sim_comp_res) %>%
  rename(Algorithm = names)

ggplot(loc4total, aes(x = index, y = results, 
                      group = Algorithm, color = Algorithm)) + 
  geom_line() +
  facet_grid(.~source) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(#title = "Results of Real Data Experiment 2", 
       subtitle = paste0("Dashed line indicates Error = 0.1."), 
       y = "log(Error)", x = "Round Index") +
  theme_bw() +
  theme(legend.position = "bottom")
```

Using a plot of the pulls at each round across the 5000 simulations in Figure 14, as we did for simulation 2, we can try to understand why some strategies perform well on synthetic data, but poorly on the time series data.

We observe that the patterns across algorithms are quite similar, while the patterns are very different across data sets. On the synthetic data, given that arm $V149$ is clearly the most difficult arm to classify due to its closeness to the threshold $\tau$, all algorithms focus a major share of their budget on this arm. The speed at which they start to do so, however, differs. Perhaps surprisingly, the best performing Bayes-UCB algorithm increases its focus on arm $V149$ the slowest, while in many iterations spending budget on exploring other arms. This is surprising as the Bayes-UCB algorithm's exploration factor was originally conceived for the cumulative regret problem. Thus one would expect it to explore *less* than other algorithms. But its performance does not seem to be easily explained by the lessened focus on the most difficult arm; else we would expect the APT algorithm to perform second best. However, both SLR and EVT perform better than APT while converging more quickly to a higher share of samples allocated to the most difficult arm.

Something that might be hidden by this aggregated view of all 5000 iterations are a share of iterations in which the EVT completely fails by exploring wrong arms for long periods. This effect might even be more drastic for the APT algorithm. Even though it seems as if its behavior is very similar to SLR and EVT, the overall performance is not better than the performance of uniform sampling. Taking this reasoning into account, one can hypothesize that the Bayes-UCB algorithm finds the most difficult arm uniformly across most iterations as the rounds increase, while other algorithms either find the most difficult arm right away or never and thus have many iterations that end with a misclassification.

For the real data, the difference between Bayes-UCB and the other algorithms is just as apparent. As is the effect of the time varying means of the arms on the rate at which they are sampled. APT, SLR, and EVT all focus quickly in many iterations on the still most difficult arm $V149$. However, in the case of the real data this behavior is not rewarded in the end, as this leads to biased samples and thus wrong estimation of the overall sample mean against which the algorithms are evaluated. We see in the curves of arms $V32$ and arm $V154$ that the EVT algorithm is less impacted by daily seasonalities, while APT clearly is.

```{r ShareOfPullsAtEachIteration, echo = FALSE, include = FALSE, fig.cap = "Budget allocation of algorithms across arms on real and synthetic data in experiment 2."}
load(file = "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/Thesis/experiment_postprocessing/amo4_arm_seq_ma.Rda")

rbind(amo4_LR_ma, amo4sim_LR_ma, amo4_AugUCB_ma, amo4sim_AugUCB_ma) %>%
  ggplot(aes(x = index, y = mean, group = arm, color = arm)) +
  geom_line(size = 0.3) +
  facet_grid(source~algorithm) +
  labs(x = "Index", y = "10 Rounds Moving Average",
       title = "Average Number of Pulls Allocated to Each Arm per Round") +
  theme_bw()
```

```{r, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Budget allocation of algorithms across arms on real and synthetic data in experiment 2."}
load(file = "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/Thesis/experiment_postprocessing/amo4_arm_seq_ma.Rda")
rbind(amo4sim_APT_ma, amo4_APT_ma, amo4_LR_ma, amo4sim_LR_ma, amo4_EVT_ma, amo4sim_EVT_ma, amo4_BUCB_ma, amo4sim_BUCB_ma) %>%
  ggplot(aes(x = index, y = mean, group = arm, color = arm)) +
  geom_line(size = 0.3) +
  facet_grid(source~algorithm) +
  labs(x = "Index", y = "10 Rounds Moving Average",
       title = "Share of Iterations Pulling an Arm at a Given Round") +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_discrete(name = "Arm")
```

### Experiments Conclusion

In the two previous experiments, we observe that the real time series data poses a major hurdle for the adaptive sampling algorithms. It seems that the poor performance can be mostly explained by the seasonality and sudden changes in the samples over time. The observations clearly do not correspond to i.i.d. Bernoulli samples, but have large time dependencies. However, the effect might be overly stark due to the fact that we use minutely time series data, which leads to observations even in minutes when there are few or no users on the website (during night). In contrast, the daily seasonality might be less represented in an actual online experiment observations are only drawn when there is an actual visitor on the site (and a proper conversion rate would fluctuate less than the count of visitors on a given product page).

Furthermore, some of the problems in dealing with the fluctuating distributions might be specific to the thresholding bandit problem as opposed to the cumulative regret or best-arm identification problems. In the experiments above, we observe that the order of the arms is mostly unaffected by the seasonalities and fluctuations, which implies that the best arm could still be identified. In contrast, given that all arms are individually classified against the *fixed* threshold, a weekly seasonality can declare a correct classification on the first half of the sample null and void on the second half of the sample.

# Conclusion

# Acknowledgements

We thank Alexandra Carpentier, Andrea Locatelli, Urun Dogan, and Marius Kloft for many hours of fruitful discussions and ideas. We thank Carlo Alberto Dall'Amico for bringing our attention to multi-armed bandits in the first place. Sonoma Internet GmbH (Amorelie) supported this work by giving the authors access to the discussed data set.

\newpage

# References

J.-Y. Audibert, S. Bubeck, and R. Munos. Best Arm Identification in Multi-Armed Bandits. In *Proceedings of the 23rd Conference on Learning Theory*, 2010.\par

E. Brochu, V.M. Cora, and N. de Freitas. A Tutorial on Bayesian Optimization of Expensive Cost Functions, with Application to Active User Modeling and Hierarchical Reinforcement Learning, *arXiv preprint arXiv:1012.2599*, 2010. \par

S. Bubeck and N. Cesa-Bianchi. Regret analysis of stochastic and nonstochastic multi-armed bandit problems. *Foundations and Trends in Machine Learning*, 5:1–122, 2012.\par

S. Bubeck, N. Cesa-Bianchi, and G. Lugosi. Bandits with heavy tail. *Information Theory, IEEE Transactions on*, 59(11):7711-7717, 2013.\par

A.N. Burnetas and M.N. Katehakis. Optimal adaptive policies for sequential allocation problems. *Advances in Applied Mathematics*, 17:122–142, 1996.\par

O. Capp´e, A. Garivier, O-A. Maillard, R. Munos, and G. Stoltz. Kullback-Leibler upper confidence bounds for optimal sequential allocation. *Annals of Statistics*, 41(3):1516–1541, 2013a.\par

O. Capp´e, A. Garivier, O.-A. Maillard, R. Munos, G. and Stoltz. Supple-
ment to “Kullback–Leibler upper confidence bounds for optimal sequential allocation.”, *Annals of Statistics*, <DOI:10.1214/13-AOS1119SUPP>, 2013b. \par

H. Chernoff. Sequential Design of Experiments. *The Annals of Mathematical Statistics*, 30(3):755-770, 1959. \par

J.C. Gittins. Bandit processes and dynamic allocation indices. *Journal of the Royal Statistical Society*. 41(2):148–177, 1979.\par

A. Garivier and E. Kaufmann. Optimal Best Arm Identification with Fixed Confidence. In *JMLR Workshop and Conference Proceedings (Conference on Learning Theory)*, 49:1-30, 2016.\par

E. Kaufmann and S. Kalyanakrishnan. Information Complexity in Bandit Subset Selection. In *JMLR Workshop and Conference Proceedings (Conference on Learning Theory)*, 30:228–251, 2013.\par

E. Kaufmann, O. Capp´e, and A. Garivier. On Bayesian Upper Confidence Bounds for Bandit Problems. In *Proceedings of the 15th International Conference on Artificial Intelligence and Statistics (AISTATS)*, 2012.\par

E. Kaufmann, O. Capp´e, A. Garivier. On the Complexity of Best Arm Identification in Multi-Armed Bandit Models. Journal of Machine Learning Research, 17:1-42, 2016.\par

T.L. Lai and H. Robbins. Asymptotically efficient adaptive allocation rules. *Advances in Applied Mathematics*, 6:4–22, 1985.\par

A. Locatelli, M. Gutzeit, and A. Carpentier. An optimal algorithm for the Thresholding Bandit Problem. In *Proceedings of the 33rd International Conference on Machine Learning*, JMLR: W&CP 48, 2016.\par

S. Mukherjee, K.P. Naveen, N. Sudarsanam, and B. Ravindran. Thresholding Bandits with Augmented UCB. 2017. *arXiv preprint arXiv:1704.02281*, 2015.\par

H. Robbins. Some aspects of the sequential design of experiments. *Bulletin of the American Mathematical Society*, 58:527-535, 1952.\par

D. Russo. Simple Bayesian Algorithms for Best Arm Identification. In *JMLR Workshop and Conference Proceedings (Conference on Learning Theory)*, 2016.\par

W.R. Thompson. On the likelihood that one unknown probability exceeds another in view of the evidence of two samples. *Biometrika*, 25:285–294, 1933.\par

W.R. Thompson. On the Theory of Apportionment. *American Journal of Mathematics*, 57:450–456, 1933.\par

J. Zhong, Y. Huang, and J. Liu. Asynchronous Parallel Empirical Variance Guided Algorithms for the Thresholding Bandit Problem. *arXiv preprint arXiv:1704.04567*, 2017.\par

\newpage

# Appendix

## Proof of Thresholding Bandit Lower Bound \label{sec:AppendixAPTLB}

We now give a proof for the lower bound for the thresholding bandit problem stated in \autoref{theorem:Locatelli2016Theorem1}.

*Proof of Theorem \ref{theorem:Locatelli2016Theorem1}*: Locatelli et al. (2016) derive a lower bound for the thresholding bandit problem. We restate the proof here. The idea is as follows. The problem formulation considered in the paper describes bandit problems for $R$-sub-Gaussian distributions. To show a lower bound, we show that every algorithm is expected to make a certain error even if we consider a very restricted scenario. Even then, the best algorithm will have a maximum error probability with which it cannot distinguish between the actual setting and slightly changed settings (under which the answer for the true setting leads to a misclassification).

Consider the thresholding bandit setting as in \autoref{sec:Setup}. More specifically, consider a problem in which all arms follow Normal distributions $\nu_i$ with mean $\mu_i$ and variance $\sigma^2 = 1$. "in which w"e set $\tau = 0$ and $\epsilon = 0$ w.l.o.g. Consequently, for each arm $i$ the distance from the threshold is given by $\Delta_i = |\mu_i - \tau| + \epsilon = \mu_i$. Thus, we have for every arm the distribution $\nu_i = \mathcal{N}(\mu_i,1) = \mathcal{N}(\Delta_i,1)$. Furthermore, define for each arm $i \in \{1, \dots, K\}$ the alternative distribution $\nu_i' = \mathcal{N}(-\Delta_i,1)$. It is easy to see that this is equivalent to a "flip around the threshold". Arms that originally have a mean larger than the threshold have a mean smaller than the threshold after the flip.

The intuition now is that every arm $i \in \{1,\dots,K\}$ generates a sample for every time point $t\in \{1, \dots, T\}$ from its distribution $\nu_i$ before the algorithm starts to choose arms. Thus, we have a $T \times K$ table of realized observations from the product distribution given by $\mathcal{B}^0 = \nu_1^0 \otimes \ldots \otimes \nu_K^0 = \nu_1^0 \otimes \ldots \otimes \nu_K$. This is a product distribution where the $K$ different means are all above the threshold. We would hope that an algorithm classifies the means accordingly. However, we will show in what follows that any algorithm makes a certain error by not being able to distinguish $\mathcal{B}^0$ from at least one of $K$ alternative product distributions $\mathcal{B}^i, i \in \{1, \dots, K\}$, of which each differs only slightly in one arm from the original distribution, and has the same overall complexity (as defined by $H = \sum_{i=1}^K (\Delta_i)^{-2}$). We require $K$ different distributions so that an algorithm sampling arm $i$ very often for some reason will not have a smaller lower bound by chance.

For each of the $K$ arms, we introduce an alternative model of the following kind. Before the start of the algorithm, we introduce a slight variation in the setup. The idea is similar to how one can switch arms in the pure exploration setup for multi-armed bandits (compare Audibert et al., 2010; Garivier & Kaufmann, 2016) to "confuse" any algorithm without introducing additional complexity to the specified problem. However, since we do not compare arms directly with each other in the thresholding bandit problem (switching arms would still keep all means above the threshold), we instead need to flip an arm $i$ with respect to the threshold in order to get the alternative model $\mathcal{B}_i$. Since the threshold is $\tau = 0$, we have for the flipped arm the new distribution $\nu_i'$ as defined above. Indeed, define the new product distribution $B^i = \nu_1^i \otimes \dots \otimes \nu_K^i$ where for $k \leq K, \nu_k' := \nu_i \mathbbm{1}_{k \neq i} + \nu_i' \mathbbm{1}_{k=i}$. This arm flip corresponds to multiplying all $X_{i,t}$ of arm $i$ by $-1$.

As in Locatelli et al. (2016), let for $i \leq K$, $\mathbb{P}_{\mathcal{B}^i}$ be the probability distribution describing all samples that a bandit strategy can potentially collect up to horizon $T$, i.e. according to the samples $(X_{k,s})_{k\leq K, s \leq T} \sim (\mathcal{B}^i)^{\otimes T}$. $(T_k)_{k\leq K}$ denotes the number of samples collected on arm $k$ until time $T$.

At this point, we have defined a finite set of problems of Gaussian distributions with fixed variance 1 for a given set of set of gaps between the means of the distributions and the threshold, $(\Delta_k)_k$. We now proof a lower bound on the worst case regret of any algorithm; that is, the maximum error probability that even the best algorithm makes on this problem.

### Statistical Decision

The expected loss in our thresholding bandit setting, $\mathbb{E}[\mathcal{L}(T)]$, corresponds directly to the probability of making a wrong classification of any of the $K$ means. In general, given a binary decision function $g$, define events $\Omega = \{g = 0\}$ and $\Omega^C = \{g = 1\}$. Assume that the decision $g$ distinguishes between two states of the world, $H_0$ and $H_1$. Then an overall error probability can be seperated into $P_{H_0}(g=1) + P_{H_1}(g = 0) = P_{H_0}(\Omega^C) + P_{H_1}(\Omega)$.

In our case, $H_0$ may correspond to the state in which we have $\mathcal{B}^0$, and $H_1$ corresponds to the case where any arm has been flipped, $\mathcal{B}^i$. Furthermore, we consider the events in which the algorithm classifies arm $i$ as being above the threshold: $\mathcal{A}_i = \{i \in \hat{S}_\tau\}$. The event on which all arms are being classified as above the threshold is consequently given by $\mathcal{A} = \bigcap_{i=1}^K \mathcal{A}_i$. We could thus separate the error probability $\mathbb{E}[\mathcal{L}(T)]$ into $\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$ and $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}^C)$. However, Theorem 1 does not bound the overall regret, but the maximum error probability on the $K+1$ different models. We can write this now as

\begin{align}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) & \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}^C) \big) \\
& = \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \label{LocatelliTheorem1ExpRegret}
\end{align}

### Empirical Kullback-Leibler Divergence

In order to give a bound on the previous equation, we start by considering $\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$. This probability can be written as the expected change in the likelihood of observed data when moving from distribution $P_{\mathcal{B}_0}$ to $P_{\mathcal{B}_i}$, times the probability under the original state. In other words, the probability of event $\mathcal{A}_i$ under state $\mathcal{B}^i$ is given as the probability of $\mathcal{A}_i$ under the (original) state $\mathcal{B}^0$ multiplied by a factor. This factor describes whether the new distribution increases or decreases the probability of the event, and is given by the change in the data likelihoods due to moving from $\mathcal{B}^0$ to $\mathcal{B}^i$, expressed as the expected likelihood ratio under $\mathcal{B}^0$. By definition, the latter is equal to the empirical Kullback-Leibler divergence of the two distributions.

We will define the change of distribution in terms of the empirical log-likelihood for our case. Then, we will show how to replace the log-likelihood by the empirical Kullback-Leibler divergence. Let us first introduce the empirical Kullback-Leibler divergence.

Let every $\nu_i, \nu_i'$ be dominated by the Lebesgue measure $\lambda$. Then their density functions are given by $f_i = \frac{\der \nu_i}{\der \lambda}$ and $f_i' = \frac{\der \nu_i'}{\der \lambda}$ respectively. For $T \geq t \geq 0$, define the empirical Kullback-Leibler divergence for an i.i.d. sample $X_{k,1}, \dots,X_{k,t}$ as

\begin{align*}
\hat{\KL}_{k,t} & = \frac{1}{t} \sum_{s=1}^{t} \log(\frac{\der \nu_k'}{\der \nu_k}(X_{k,s})) \\
& = \frac{1}{t} \sum_{s=1}^{t} \log \big(\frac{f_k'(X_{k,s})}{f_k(X_{k,s})} \big) \\
& = \frac{1}{t} \log \big( \prod_{s=1}^{t} \frac{f_k'(X_{k,s})}{f_k(X_{k,s})} \big) \\
& \stackrel{\text{iid}}{=} \frac{1}{t} \log \big( \frac{f_k'(X_{k,1}, \dots,X_{k,t})}{f_k(X_{k,1}, \dots,X_{k,t})} \big) \\
& = \frac{1}{t} \log \big( \text{LR}((X_{k,1}, \dots,X_{k,t}), \nu_k', \nu_k) \big),
\end{align*}

where $\text{LR}((X_{k,1}, \dots,X_{k,t}), \nu_k', \nu_k)$ is the likelihood ratio for the sample.

Since we consider Gaussian distributions with variance $\sigma^2 = 1$, we have density functions $f_i(x) \propto \exp \big(-\frac{1}{2} (x-\Delta_i)^2\big)$ and $f_i'(x) \propto \exp \big(-\frac{1}{2} (x+\Delta_i)^2\big)$. Plugging this into the above definition of the empirical Kullback-Leibler divergence, we easily see

\begin{align*}
\hat{\KL}_{k,t} & = \frac{1}{t} \sum_{s=1}^{t} \log \Big(\frac{\exp \big(-\frac{1}{2} (X_{k,s}+\Delta_k)^2\big)}{\exp \big(-\frac{1}{2} (X_{k,s}-\Delta_k)^2\big)} \Big) \\
& = \frac{1}{t} \sum_{s=1}^{t} \log \Big( \exp \big(-\frac{1}{2} (X_{k,s}+\Delta_k)^2 + \frac{1}{2} (X_{k,s}-\Delta_k)^2 \big) \Big) \\
& = - \frac{1}{t} \sum_{s=1}^{t} 2 X_{k,s} \Delta_k
\end{align*}

### Change of Distribution

We are now ready to state the change of distribution. As alluded to above, we consider how the probability changes for the event $\mathcal{A}_i$ when we move from $\mathcal{B}^0$ to $\mathcal{B}^i$. When flipping arm $i$, we only change the samples of this arm, that is, the first $T_i$ observations that the algorithm draws from arm $i$. The change of distribution is then defined as:

\begin{align*}
\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & = \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i} \text{LR}((X_{i,1}, \dots,X_{i,T_i}), \nu_i', \nu_i) ] \\
& = \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i} \exp (- T_i \hat{\KL}_{i,T_i}) ]
\end{align*}

Obviously, $\hat{\KL}_{i,T_i}$ depends on the realized samples $X_{i,1}, \dots,X_{i,T_i}$. Consequently, we need to find a bound on $\hat{\KL}_{i,T_i}$ itself to incorporate the uncertainty involved in this metric.

### Concentration of the Empirical Kullback-Leibler Divergence

For our problem, the Kullback-Leibler divergence of $\nu_k'$ from $\nu_k$ is given by $\KL_k := \KL(\nu_k', \nu_k) = 2\Delta_k^2$, and consequently

\begin{equation*}
|\hat{\KL}_{k,t} - \KL_{k,t}| = |-\frac{2}{t} \Delta_k \sum_{s=1}^{t}(X_{k,s} - \Delta_k)|
\end{equation*}

itself is a normally distributed random variable with mean zero and a variance decreasing in $t$. Thus, one can show as in Locatelli et al. (2016) that $\mathbb{P}_{\mathcal{B}^i}(\xi) \geq 3/4$ for the event

\begin{equation}
\xi = \{ \forall k \leq K, \forall t \leq T, |\hat{\KL}_{k,t} - \KL_{k,t}| \leq 4 \Delta_k \sqrt{\frac{\log(4(\log(T)+1)K)}{t}}\}. \label{LocatelliTheorem1EventXi}
\end{equation}

Thus, in order to bound the empirical Kullback-Leibler divergence in our change of measure, we can consider the intersection of $\mathcal{A}_i$ and $\xi$ as follows and plug in the bound on event $\xi$:

\begin{align*}
\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & = \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i} \exp (- T_i \hat{\KL}_{i,T_i}) ] \\
& \geq \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- T_i \hat{\KL}_{i,T_i}) ] \\
& \geq \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 2 \Delta_i^2 T_i - 4 \Delta_i \sqrt{T_i} \sqrt{\log(4(\log(T)+1)K)}) ]
\end{align*}

Two things are left to do: With $T_i$, we still have one random variable left which we cannot leave in our bound. Second, we are actually not interested in $\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$ but in $\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$.

### Bounding $T_i$ for $\max_{i \in \{1,\dots,K\}}\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$

We are actually interested in bounding $\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)$. In general it holds that $\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \frac{1}{K}\sum_{i=1}^K X_i$. Thus we can write:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) \\
& \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i \cap \xi) \\
& \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 2 \Delta_i^2 T_i - 4 \Delta_i \sqrt{T_i} \sqrt{\log(4(\log(T)+1)K)}) ]
\end{align*}

where the second line is again the intersection of $\mathcal{A}_i$ and $\xi$ as before, such that we can apply the change of distribution in the third line.

What is left to show is a bound on $T_i$. Given that we want to show that even the best algorithm makes an error on at least one of the problems (one of the arms), it makes sense to show that $T_i$ is upper bounded. This implies that no arm can be infinitely sampled; we cannot reach full confidence in any arm, and a residual error probability should remain.

First, let $a = \Delta_i \sqrt{T_i}$ and $b = 4\sqrt{\log((4\log(T)+1)K)}$. Then use $ab \leq a^2 + b^2$ to write $ab \leq \Delta_i^2 T_i + 16\log((4\log(T)+1)K)$. And so:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 3 \Delta_i^2 T_i - 16 \log(4(\log(T)+1)K)) ]
\end{align*}

Furthermore, we know that the sum of the individual arm pulls has to be equal to the overall number of pulls (equal to the budget in the fixed budget sense), $\sum_i T_i = T$, and furthermore we know that $T_i>0$. The latter holds since any reasonable algorithm has to check each arm at least once in our setup. Thus it is easy to show that $\exists i: T_i \leq \frac{T}{H \Delta_i^2} = \frac{T}{(\sum_{i=1}^{K} \Delta_i^{-2}) \Delta_i^2}$. To see this, assume the opposite: $\forall i: T_i > T \frac{\sum_{i=1}^K \Delta_i^2}{\Delta_i^2}$. This implies however $\sum_{i=1}^K T_i > \sum_{i=1}^K T \frac{\sum_{i=1}^K \Delta_i^2}{\Delta_i^2} = T \cdot 1 = T$ which is a contradiction.

So we can write $(\Delta_i \sqrt{T_i})^2 \leq \frac{T}{H}$. We use $\mathcal{A} = \cap_{i=1}^K \mathcal{A}_i$ to write:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 3 \Delta_i^2 T_i - 16 \log(4(\log(T)+1)K)) \big] \\
& = \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A} \cap \xi} \frac{1}{K} \sum_{i=1}^{K} \exp (- 3 \Delta_i^2 T_i - 16 \log(4(\log(T)+1)K)) \big]
\end{align*}

We write $\frac{1}{K} \sum_{i=1}^K \exp(-3T_i \Delta_i^2) \geq \frac{1}{K} \sum_{i=1}^K \exp(-3 T/H) = \exp(-3 T/H)$. Thus:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & = \mathbb{E}_{\mathcal{B}^0} \Big[\mathbbm{1}_{\mathcal{A} \cap \xi} \frac{1}{K} \sum_{i=1}^{K} \exp (- 3 \Delta_i^2 T_i) \exp \big(-16 \log(4(\log(T)+1)K)\big) \Big] \\
& \geq \mathbb{E}_{\mathcal{B}^0} \Big[\mathbbm{1}_{\mathcal{A} \cap \xi} \frac{1}{K} \sum_{i=1}^{K} \exp \big(- 3 \frac{T}{H}\big) \exp \big(-16 \log(4(\log(T)+1)K)\big) \Big] \\
& = \mathbb{E}_{\mathcal{B}^0} \Big[\mathbbm{1}_{\mathcal{A} \cap \xi} \exp \big(- 3 \frac{T}{H}\big) \exp \big(-16 \log(4(\log(T)+1)K)\big) \Big] \\
& = \mathbb{P}_{\mathcal{B}^0} (\mathcal{A} \cap \xi) \exp \Big(- 3 \frac{T}{H} -16 \log \big(4(\log(T)+1)K\big)\Big)
\end{align*}

### Bringing everything together

Consider again the risk which we need to bound from equation \eqref{LocatelliTheorem1ExpRegret}. We bound the maximum expected loss over all $\mathcal{B}_i$. By using again the fact that in general $\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \frac{1}{K} \sum_{i = 1}^K X_i$, we plug into our previous calculations:

\begin{align}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) & \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \\
& \geq \frac{1}{2}\mathbb{P}_{\mathcal{B}^0} (\mathcal{A} \cap \xi) \exp \Big(- 3 \frac{T}{H} -16 \log \big(4(\log(T)+1)K\big)\Big) + \frac{1}{2}(1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A})) \label{LocatelliTheorem1DefinitionOfRisk}
\end{align}

From \eqref{LocatelliTheorem1EventXi} we know that for any $i \in \{0, \dots, K\}$, $\mathbb{P}_{\mathcal{B}^i}(\xi) \geq 3/4$. Following the argument in Locatelli et al. (2016), we now consider two cases: $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \geq 1/2$ and $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \leq 1/2$. In the former case, we can plug in the two probabilities into \eqref{LocatelliTheorem1DefinitionOfRisk}. If $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \geq 1/2$ and $\mathbb{P}_{\mathcal{B}^0}(\xi) \geq 3/4$, then $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A} \cap \xi) \geq 1/4$. Thus:

\begin{align*}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) & \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \\
& \geq \frac{1}{2}\mathbb{P}_{\mathcal{B}^0} (\mathcal{A} \cap \xi) \exp (- 3 \frac{T}{H} -16 \log(4(\log(T)+1)K)) + \frac{1}{2}(1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A})) \\
& \geq \frac{1}{8} \exp (- 3 \frac{T}{H} -16 \log(4(\log(T)+1)K))
\end{align*}

where we drop the second summand of the second line so that the inequality holds.

On the other hand, when $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \leq 1/2$, then we know that $\mathbb{P}_{\mathcal{B}^0}(\mathcal{A} \cap \xi) \leq 1/2$ for any $\mathbb{P}_{\mathcal{B}^0}(\xi)$. At the same time $1-\mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) = \mathbb{P}_{\mathcal{B}^0}(\mathcal{A^C}) \geq 1/2$. Consequently, $1/4 \leq \frac{1}{2}(1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A})) \leq 1/2$, while 
$\frac{1}{2}\mathbb{P}_{\mathcal{B}^0} (\mathcal{A} \cap \xi) \exp (- 3 \frac{T}{H} -16 \log(4(\log(T)+1)K)) \geq 0$. And so the stronger lower bound in that case is, as noticed by Locatelli et al. (2016), given simply by

\begin{align*}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) & \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \\
& \geq 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) = \mathbb{P}_{\mathcal{B}^0}(\mathcal{A^C}) \\
& \geq \frac{1}{2}
\end{align*}

## Lower Bound for Thresholding Bandit Problem with KL-divergence \label{sec:corollaryLBTBPproof}

### Lower Bound for One-Armed Thresholding Bandit \label{sec:corollaryLB1TBPproof}

**Proof**: The proof follows closely the proof of Theorem 12 in Kaufmann et al. (2016) and is again an application of their transportation Lemma, Lemma 2.

Consider the one-armed bandit problem, $K = 1$ with the canonical exponential family of distributions distribution $\nu_1$ parameterized by its mean $\mu_1$. Furthermore, given we are in the thresholding bandit problem, we have a threshold $\tau$ against which the arm is classified. Assume w.l.o.g. that for $\nu_1$, $\mu_1 \geq \tau$. Thus, $\mathcal{S}_{\tau}^* = \{1\}$. After $T$ rounds, a consistent algorithm returns with error probability $p_T(\nu)$ the guess $\hat{\mathcal{S}}_{\tau}$. Thus, we have for the event $A = (\hat{\mathcal{S}}_\tau = \{1\})$ the probability $\mathbb{P}_{\nu}(A)$. On the other hand, define the alternative model $\nu'$ with mean $\mu_1' < \tau$ such that the correct classification from $\nu$ is now wrong: $\mathcal{S}_{\tau}^* {'} = (\mathcal{S}_{\tau}^*)^C = \{\emptyset\}$. Consequently, since the algorithm is still consistent on this problem, we have that $\mathbb{P}_{\nu'}(A) = p_t(\nu')$. Thus, we expect that as we let $t$ increase, $\mathbb{P}_{\nu'}(A)$ will decrease while $\mathbb{P}_{\nu}(A)$ will increase.

Indeed, given we consider only consistent algorithms, there should exist a number of samples for which the probability of the correct decision in problem $\nu$ is larger than the error probability in problem $\nu'$, that is: for every $\epsilon > 0$, there exists $t_0(\epsilon)$ such that for all $t \geq t_0(\epsilon)$ we have

\begin{equation*}
\mathbb{P}_{\nu'}(A) \leq \epsilon \leq \mathbb{P}_{\nu}(A). \label{eq:BoundForCorollary}
\end{equation*}

Now, consider again Lemma 2 in Kaufmann et al. (2016). If we let the algorithm be such that $T = t$, we can apply the Lemma to the stopping time $\sigma = t$ a.s. and the event $A$ to get

$$
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)).
$$

We can lower bound the RHS by using the bound on $\mathbb{P}_{\nu'}(A)$ given in \eqref{eq:BoundForCorollary} and the monotonicity of the binary relative entropy. We get:

\begin{align*}
\mathbb{E}_{\nu'}[N_1(t)]\KL(\nu_1', \nu_1) & = t \KL(\nu_1', \nu_1) \\
& \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)) \\
& \geq d(\epsilon, 1- p_t(\nu)) \\
& = \epsilon \log\Big(\frac{\epsilon}{1-p_t(\nu)}\Big) + (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) \\
& \geq (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) + \epsilon \log(\epsilon) \\
\end{align*}

Letting $\epsilon \rightarrow 0$, we get

$$
t \KL(\nu_1', \nu_1) \geq \log \Big(\frac{1}{p_t(\nu)}\Big) = -\log p_t(\nu).
$$

To make the bound as tight as possible, we choose the alternative model that minimizes $\KL(\nu_1', \nu_1)$ while still ensuring that $\mathcal{S}_{\tau}^* {'} = \{\emptyset\}$. In our case of the thresholding bandit with one-parameter exponential family distributions, this is given by choosing $\mu_1' \rightarrow \tau$ (not equal to $\tau$ since else the classification of $\nu_1$ is again correct). By pluggin in, we get the result:

$$
t \KL(\tau, \mu_1) \geq -\log p_t(\nu) \quad \text{for} \; \mu_1' \rightarrow \tau.
$$

### Lower Bound for $K$-Armed Thresholding Bandit \label{sec:corollaryLBKTBPproof}

**Proof**: The proof is very similar to the proof of the corollary for $K=1$. However, the alternative model needs more motivation than in the case of a single arm. But first, consider again that we are in the thresholding bandit problem with $K$ arms of one-parameter canonical exponential family distributions that are parameterized by their means $\mu_1, ..., \mu_K$ for the original problem $\nu = (\nu_1, ..., \nu_K)$. W.l.o.g, assume that all arms are above the threshold, that is, $\mu_1 \geq \tau, ..., \mu_K \geq \tau$. Then $\mathcal{S}_{\tau}^* = \{1, ... , K\}$. The correct classification returned after $t$ rounds is thus the event $A = (\hat{\mathcal{S}}_{\tau} = \{1, ..., K\})$. We can now consider Lemma 2. The RHS is maximized when the alternative model is chosen such that $d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A))$ is maximized. Given the event $A$ as defined before, this is the case when $\mathbb{P}_{\nu'}(A)$ is as small as possible. For consistent algorithms, this is the case when $\nu'$ is chosen such that all $\mu_1' < \tau, ..., \mu_K' < \tau$. Then even classifying a single arm correctly under $\nu$ would lead to a mistake in $\nu'$. Consequently, $A$ is the most unlikely classification for a consistent algorithm for the problem $\nu'$. We thus expect for any consistent algorithm that $\mathbb{P}_{\nu}(A)$ becomes large and $\mathbb{P}_{\nu'}(A)$ becomes small as the number of observations $t$ increases. Again, for every $\epsilon > 0$, there exists $t_0(\epsilon)$ such that for all $t \geq t_0(\epsilon)$ we have $\mathbb{P}_{\nu'}(A) \leq \epsilon \leq \mathbb{P}_\nu(A)$. This allows us to now use Lemma 1 as previously:

\begin{align*}
\sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\nu_i', \nu_i) & \geq \sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\tau, \mu_i) \\
& \geq d(\mathbb{P}_{\nu'}(A), \mathbb{P}_{\nu}(A)) \\
& \geq d(\epsilon, 1-p_t(\nu)) \\
& = \epsilon \log \Big(\frac{\epsilon}{1-p_t(\nu)}\Big) + (1-\epsilon) \log \Big( \frac{1-\epsilon}{p_t(\nu)}\Big) \\
& \geq (1-\epsilon) \log \Big(\frac{1-\epsilon}{p_t(\nu)}\Big) + \epsilon \log(\epsilon)
\end{align*}

We let $\epsilon \rightarrow 0$, divide by $t$ and take the limsup on both sides as in Kaufmann et al. (2016):

\begin{align*}
\lim_{t \rightarrow \infty} \sup - \frac{1}{t} \log p_t(\nu) & \leq \lim_{t \rightarrow \infty} \sup \frac{1}{t} \sum_{i=1}^K \mathbb{E}_{\nu'}[N_i(t)]\KL(\tau, \mu_i) \\
& \leq \max_{i \in \{1,...,K\}} \KL(\tau, \mu_i)
\end{align*}

Though, since for the alternative problem $\nu'$ we have for every $i \in \{1,...,K\}$ that $\mu_i = \tau$, it should hold for any consistent algorithm that $\mathbb{E}_{\nu'}[N_i(t)] = \frac{t}{K}$, that is, every arm receives an equal amount of the budget. In that case, the above equation simplifies to

$$
- \frac{1}{t} \log p_t(\nu) \leq \frac{1}{t} \sum_{i=1}^K \frac{t}{K}\KL(\tau, \mu_i) = \frac{1}{K} \sum_{i=1}^{K} \KL(\tau, \mu_i),
$$

that is, the complexity is given by the simple average of the Kullback-Leibler distances of the threshold with regard to the respective distributions. This relates well to the lower bound that Kaufmann et al. (2016) describe in their Theorem 12 for consistent algorithms using *uniform* sampling strategies; in our case, however, it is explained by the fact that under the alternative model, every consistent algorithm should adapt to a uniform sampling strategy. Thus our bound holds for any consistent strategy.

## Bayes-UCB Algorithm for the Thresholding Bandit Problem \label{sec:AppendixBUCB}

We show how the Bayes-UCB algorithm (Algorithm 1, Kaufmann et al., 2012) is adapted to the thresholding bandit problem to be used in the experiments in chapter \autoref{chap:Experiments}. The adapted version of their Algorithm 1 is given in Algorithm 4 below. In particular, notice that we not only calculate upper quantiles $q_i^u(t)$, but also lower quantiles $q_i^l(t)$, as we have to be able to compare against $\tau$ from both sides. These quantiles rely on the posterior distributions $\lambda_i^t$ for each mean $\mu_i$. In our case, we employ Beta prior distributions as conjugate prior for the Bernoulli distributions in simulations 1 to 3 and initialize them with parameters $\alpha_0$ and $\beta_0$. See the setup in Kaufmann et al. (2012), especially equation (3), to see how the posterior distribution is updated in every round.

\IncMargin{1em}
\begin{algorithm}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{\emph{$\tau$, $\epsilon$, $T$, $\Pi^0$ (initial prior on $\mu$), $c$ (parameters of the quantile)}}
\Output{$\hat{S}_\tau = \{i \in \mathbb{A},  \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
\For{$i = 1$ \KwTo $K$}{
  \emph{Compute $q_i^u(t) = Q\Big(1-\frac{1}{t(\log T)^c}, \lambda_i^{t-1}\Big)$ and $q_i^l(t) = Q\Big(\frac{1}{t(\log T)^c}, \lambda_i^{t-1}\Big)$}\;
  \emph{Compute $B_i^{BUCB}(t) = (\tau - q_i^l(t))\cdot\mathbb{I}(\hat{\mu}_i(t) \geq \tau) + (q_i^u(t) - \tau)\cdot\mathbb{I}(\hat{\mu}_i(t) < \tau)$}\;
}
  \emph{Pull arm $I_t = \arg \max_{i\leq K} B_i^{BUCB}(t)$}\;
  \emph{Observe reward $X \sim \nu_{I_t}$ and update $\Pi^t$}\;
}
\caption{Bayes-UCB algorithm from Kaufmann et al. (2012) adapted to the thresholding bandit problem.}\label{bayesucb_tbp}
\end{algorithm}\DecMargin{1em}

## Empirical Analysis of the Bayes-UCB Algorithm in Simulation 2

In order to assess the excellent performance of the Bayes-UCB algorithm, which comes at a surprise given that is is designed for the cumulative regret setting and thus should tend to explore the arms to less, we run the Bayes-UCB algorithm for simulation 2 with three different choices of the parameter. In Figure 13, we see the performance results of the three versions, whereas Figure 14 shows how the algorithm allocates its budget across the arms over time. The three versions differ in the quantile parameter $c$ (compare \autoref{sec:AppendixBUCB}) for which we choose $c \in \{1/5, 1, 5\}$ to explore more or less. In Kaufmann et al. (2012), the cumulative regret upper bound for the Bayes-UCB is shown for $c\geq 5$, though they report that the choice of $c=0$ worked much better in their experiments. In our problem, it makes also sense that the versions with smaller quantiles (smaller $c$) tend to perform better, as these correspond to more exploration of the arms. In this particular simulation, the choices of $c=1/5$ and $c=1$ perform very similar. We see that the worse performance of $c=5$ can be attributed to a initial phase of sampling primarily the arms with mean below the threshold. This comes from the fact that the Bernoulli and Beta distributions are bounded between $0$ and $1$, with $\tau$ very close to $0$. Thus, the index $B_i^{BUCB}(t)$ cannot be larger than $\tau$ for arms with $\hat{\mu}_i(t) \geq \tau$. For arms with $\hat{\mu}_i(t) < \tau$, however, the index can be as large as $B_i^{BUCB}(t) = 0.95$. This leads to the effect seen in Figure 14: Due to $c=5$, the indexes are based on very large quantiles even in the beginning, which leads to large indexes for the arms with small means. This however is not the ideal sampling behavior. The other two variants adapt much better due to their smaller quantiles (and more exploration).

```{r simulation2_results_bucb, echo = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Performance comparison of algorithms for simulation 2."}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(paste0(current_path, "loc6nt_comp_BUCB_horizon_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_BUCB_horizon_c5_7000.Rda"))
load(paste0(current_path, "loc6nt_comp_BUCB_horizon_c15_7000.Rda"))
#load(paste0(current_path, "loc6nt_comp_BUCB_horizon_7000.Rda"))

alg_names <- c(rep(c("BUCB_C0.25", "BUCB_C1", "BUCB_C5"), 
                   each = 6991))
index <- c(rep(10:7000, times = 3))
alg_res <- log(c(loc6nt_comp_BUCB_horizon_c15_7000,
                 loc6nt_comp_BUCB_horizon_7000,
                 loc6nt_comp_BUCB_horizon_c5_7000))
loc6nt_comp_res <- data.frame(index = index, names = alg_names, 
                            results = alg_res)
ggplot(loc6nt_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = names, color = names)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(#title = "Results of Synthetic Experiment 2", 
       subtitle = "Dashed line at Error = 0.1.", 
       y = "log(Error)", x = "Round Index") +
  scale_color_discrete(name = "Algorithm") +
  theme_bw() 
```

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Budget allocation across arms for the Bayes-UCB algorithm in simulation 2 at three different values for the quantile parameter."}
rbind(loc6nt_BUCB_c15_ma, loc6nt_BUCB_ma, loc6nt_BUCB_c5_ma) %>%
  rename(Arm = arm) %>%
  ggplot(aes(x = index, y = mean, group = Arm, color = Arm)) +
  geom_line(size = 0.3) +
  facet_grid(.~algorithm) +
  labs(x = "Index", y = "10 Rounds Moving Average"#,
       #title = "Average Number of Pulls Allocated to Each Arm per Round",
       #subtitle = "5000 iterations; a value of 1000 means that in 20% of 
       #the iterations the arm was pulled in that round."
       ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

## Empirical Comparison of APT, SLR, and Bayes-UCB in Simulation 2

We can compare the different algorithms used in Simulation 2 based on the overall number of pulls of each arm. In the best case, an algorithm would sample an arm for the optimal share of the budget, in every simulation. This is of course not the case, since the number of pulls of a given arm at the end of the budget, $T_i(T)$, is a random variable for adaptive sampling strategies. Given the 5000 iterations conducted in simulation 2, we can plot the empirical distribution of this random variable for different arms.

In Figure 15, we show the distribution of $T_6(T)$ for algorithms APT, Bayes-UCB, and SLR. We see that the distribution of the Bayes-UCB algorithm is much better concentrated around the optimal share of the budget, and has a bell curve shape. In contrast, the distributions of APT and SLR are much less concentrated but flat and wide, and spread across a very large interval. Both algorithms have several iterations in which they draw arm 6, the most difficult arm only a single time (during initialization). This happens for APT more than 5% of the iterations, and leads to the spike on the very left of the graph. In order to cope at least a little bit with this problem, we have also run the SLR algorithm with the D-Tracking rule as introduced in Garivier and Kaufmann (2016). The effect is nearly not noticeable in Figure 15, but the minimum number of observations is not clearly bounded away from 0. The effect on the performance was described before and is very noticeable in comparison to SLR without D-Tracking. However, even with D-Tracking the performance of Bayes-UCB is not reached. We describe SLR with D-Tracking in Algorithm 5.

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Empirical distribution of pulls of arm 6 in simulation 2 by different algorithms."}
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/synthetic_experiments/"

load(paste0(current_path, "loc6nt_pulls_of_arms_vis.Rda"))

pulls_arm_6_long %>%
  ggplot(aes(x = Pulls)) +
  geom_histogram(binwidth = 100) +
  facet_grid(.~Algorithm, scales = "free_y") +
  geom_vline(aes(xintercept = 7000/10), linetype = 2) +
  labs(x = "Pulls", y = "Count"#, 
       #title = "Distribution of Pulls of Arm 6 Across Simulations",
       #subtitle = "Binwidth = 100. Budget = 7000. 5000 Simulations.
#Vertical line indicates pulls by uniform sampling given 10 arms in experiment."
) +
  theme_bw()
```

```{r, echo = FALSE, message = FALSE, fig.align = "center", out.width = '80%', fig.cap = "Empirical distribution of pulls of arms 1 and 10 in simulation 2 by different algorithms."}

cbind.data.frame(
  Arm = rep(c("01", "10"), each = 20000),
  rbind.data.frame(pulls_arm_1_long, pulls_arm_10_long),
  stringsAsFactors = FALSE
) %>% 
  ggplot(aes(x = Pulls, group = Arm)) +
  geom_histogram(binwidth = 50) +
  facet_grid(Arm~Algorithm, scales = "free_y")+
  geom_vline(aes(xintercept = 7000/10), linetype = 2) +
  labs(x = "Pulls", y = "Count"#, 
       #title = "Distribution of Pulls of Arms 1 and 10 Across Simulations",
       #subtitle = "Binwidth = 50. Budget = 7000. 5000 Simulations.
#Vertical line indicates pulls by uniform sampling given 10 arms in experiment."
) +
  theme_bw()
```

## SLR Algorithm with D-Tracking Rule

\IncMargin{1em}
\begin{algorithm}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$, class of distributions $\nu(\mu, \theta)$, $\theta_i \: \forall i \in \mathbb{A}$}
\Output{$\hat{S}_\tau = \{i \in \mathbb{A},  \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Calculate $U_t = \{i \in \mathbb{A}, T_i(t) < \sqrt{t} - K/2\}$}\;
  \emph{Set $I_t = \arg \min_{i \in \mathbb{A}} B_i^{SLR}(t)$ as in \autoref{eqn:SLRIndex}}\;
  \If{$U_t \neq \emptyset$}{Set $I_t = \arg \min_{i \in \mathbb{A}} T_i(t)$}
  \emph{Pull arm $I_t$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{SLR algorithm with D-Tracking Rule (Garivier and Kaufmann, 2016)} \label{algo_slr_1d_dTracking}
\end{algorithm}
\DecMargin{1em}

## Means in Simulations 4 and 5

We give here the means of the distributions used in simulations 4 and 5 for completeness.

```{r, echo = FALSE, fig.align = "center", out.width = '80%'}
library(ggplot2)

set.seed(93468734)
tau_exp2 <- 1
epsilon_exp2 <- 0
mu_exp2 <- rexp(20, 1)

knitr::kable(t(sort(mu_exp2)[1:10]), digits = 4, col.names = paste0("V", 1:10), caption = "Mean of Exponential arms 1 to 10 in simulation 4.")

knitr::kable(t(sort(mu_exp2)[11:20]), digits = 4, col.names = paste0("V", 11:20), caption = "Mean of Exponential arms 11 to 20 in simulation 4.")

set.seed(3795969)
tau_pois2 <- 0.4
epsilon_pois2 <- 0
mu_pois2 <- rexp(20, 1/0.7)

knitr::kable(t(sort(mu_pois2)[1:10]), digits = 4, col.names = paste0("V", 1:10), caption = "Mean of Poisson arms 1 to 10 in simulation 5.")

knitr::kable(t(sort(mu_pois2)[11:20]), digits = 4, col.names = paste0("V", 11:20), caption = "Mean of Poisson arms 11 to 20 in simulation 10.")
```